### 2018-01-03 term paper_2018/paper.tex
upd

\documentclass[nobib]{tufte-handout}

\usepackage{ifluatex, ifxetex} % проверка, как компилируется файл

% black magic!
% solving problem with tufte-handout + xelatex
% http://tex.stackexchange.com/questions/200722/
% https://tex.stackexchange.com/questions/47576/
\ifx\ifxetex\ifluatex\else % если xelatex или lualatex
  \newcommand{\textls}[2][5]{%
    \begingroup\addfontfeatures{LetterSpace=#1}#2\endgroup
  }
  \renewcommand{\allcapsspacing}[1]{\textls[15]{#1}}
  \renewcommand{\smallcapsspacing}[1]{\textls[10]{#1}}
  \renewcommand{\allcaps}[1]{\textls[15]{\MakeTextUppercase{#1}}}
  \renewcommand{\smallcaps}[1]{\smallcapsspacing{\scshape\MakeTextLowercase{#1}}}
  \renewcommand{\textsc}[1]{\smallcapsspacing{\textsmallcaps{#1}}}
\fi

% с альтернативного источника:
%\ifluatex % Allow rendering with LuaTeX by setting up the spacing using fontspec features
%  \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
%  \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
%\fi

% lua produces strange notes
% https://tex.stackexchange.com/questions/328431

\usepackage{fontspec} % работа со шрифтами
%\usepackage{polyglossia} % учим русский как иностранный :)

%\setmainlanguage{engish}
%\setotherlanguages{english}

% заменяем --- на тире, << на кавычки и т.д.:
\defaultfontfeatures{Ligatures=TeX}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}
\newfontfamily{\cyrillicfont}{Linux Libertine O}
\newfontfamily{\cyrillicfontsf}{Linux Libertine O}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url} % вставка \url{}
\usepackage{graphicx} % вставка графиков
\usepackage{csquotes} % адаптирующиеся кавычки командой \enquote{}
\usepackage{comment} % ingore everything between \begin{comment} \end{comment}
\usepackage{answers} % separate problems and solutions
\usepackage{tikz} % pictures with tikz language
\usepackage{todonotes} % todo in documents
\usepackage{multicol} % make several columns
\usepackage{subfigure}

\usepackage{enumitem} % для создания своих нумерующих списков (хак для гиперссылок)

%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

% very useful during de-bugging!
% \usepackage[left]{showlabels}
% \showlabels{hypertarget}
% \showlabels{hyperlink}

\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Corr}{Corr}
\DeclareMathOperator{\E}{E}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\I}{\mathbb{I}} % индикатор события
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\usepackage[bibencoding = auto, backend = biber,
sorting = none]{biblatex}

%\addbibresource{probability_dna.bib}

%\title{Geometric interpretaion of main concepts and theorems in econometrics}
\title{some title}
\author{}
\date{\today}
\begin{document}

\maketitle

%\section{Literature review}

\section{Regression}

\subsection{Correlation}

\marginnote{The most widespread definition of correlation is
\[
\Corr(X,Y)  = \frac{\Cov(X, Y)}{\sqrt{\Var(X)\Var(Y)}}
\]}

The most crucial part in defining correlation geometrically is definig the dot product as it enables to compute the length of a vecotr:
\[
|\vec{a}| = \sqrt{\langle \vec{a},  \vec{a}\rangle}
\]
and the angle between any two vectors:
\[
\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}
\]
Now we define scalar product of two random vectors as covariation between them:
\[
\langle X, Y \rangle = \Cov(X, Y)
\]
The main characteristics of a random vector are its length and direction.
So, we introduce the length
\[
\sqrt{\Cov{X,X}} = \sqrt{\Var(X)} = \sigma_X
\]
and the angle between two random vectors
\[
\cos(X,Y) = \frac{\Cov(X,Y)}{\sqrt{\Var(X)\Var(Y)}} = \Corr(X,Y)
\]
Note that from the definition of the angle it follows that correlation is limited within an interval $[-1,1]$.

\begin{figure}[ht!]
\begin{center}
\includegraphics{images/corr_def.pdf}
\label{fig:corr_def}
\caption{Random vector $X$ of length $\sigma_X$ and random vector $Y$ of length $\sigma_Y$, $\cos \varphi$ is the angle between $X$ and $Y$.}
\end{center}
\end{figure}

Another important geometrical tool is projection. Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$ can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
Same holds for random vectors. The projection of a random vector $Y$ onto $\{cX| c \in \mathbb{R}\}$ is $\Corr(X,Y) \cdot Y$.

\subsection{Regression line and point of averages}

\marginnote{
If the regression contains the intercept, the following equation holds:
\begin{align*}
\hat y  &= X \hat \beta = X (X^T X)^{-1} X^T y \\
&= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon
\end{align*}
Premultiplying both sides by $X^T$, we obtain:
\begin{align*}
X^T \hat y &=  X^T X (X^T X)^{-1} X^T X \beta \\
&+ X^T X (X^T X)^{-1} X^T \varepsilon \\
&= X^T X \beta + X^T \varepsilon
\end{align*}
This is a system of equations. The first row of $X^T$ is $\mathbf{1}$ vector, so we can write out the first equation:
\[
\sum_{i=1}^n \hat y_i = \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
\]
From the first equation in the system
\[
X^T \hat y = X^T y
\]
we obtain
\[
\sum_{i=1}^{n} \hat y_i = \sum_{i=1}^{n} y
\]
And this finishes the proof:
\[
\frac{1}{n} \sum_{i=1}^{n} y = \frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
\]
}

\begin{theorem}
The point of averages lies on the estimated regression line.
\end{theorem}

\begin{proof}
For the geometrical proof it suffices to show that $\hat y$ is a linear combination of the regressors, which is true by construction,
and that $\frac{1}{n} \sum_{i=1}^{n} \hat y_i = \frac{1}{n} \sum_{i=1}^{n} y$. In order for the pictures to be more clear the proof will be presented for the case of two regressors.

The first step is regressing $y$ on $Lin(\mathbf{1}, x)$. As shown in Figure~\ref{fig:averages_lin}, we obtain $\hat y$ as a linear combination of $\mathbf{1}$ and $x$.
The next step is to regress both $y$ and $\hat y$ on $\mathbf{1}$ which results in $\bar y$ and $\bar \hat y$ correspondingly.
By the theorem of three perpendiculars, $\bar y = \bar \hat y$ which is shown in Figure~\ref{fig:averages_bars}.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:averages_lin} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
%\caption{$Lin(x, z)$}
\label{fig:averages_bars}}
\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
\end{center}
\end{figure}

\end{proof}

\subsection{Frisch–Waugh–Lovell theorem}

\marginnote{From regresison~\ref{eq:fwl_2} we get the following estimator:
\begin{align*}
\hat\beta_2 &= ((M_1 X_2)^T M_1 X_2)^{-1}(M_1 X_2)^T M_1 y \\
&= (X_2^T M_1^T M_1 X_2)^{-1}  X_2^T M_1^T M_1 y \\
&= (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
\end{align*}
As for regresison~\ref{eq:fwl_1}, let us note that due to $y = \hat y + \hat u$ $y$ can be decomposed as follows:
\[
y = Py + My = X_1 \hat \beta_1 + X_2 \hat \beta_2 + My
\]
Premultiplying both sides by $X_2^T M_1$, we obtain:
\begin{align*}
X_2^T M_1 y &= X_2^T M_1 X_1 \hat\beta_1 + X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
&=  X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
&= X_2^T M_1 X_2 \hat\beta_2
\end{align*}
On the last step we used the fact that
\begin{align*}
(X_2^T M_1 M y)^T = y^T M^T M_1^T X_2 \\
= y^T M M_1 X_2 = y^T M X_2 = 0^T
\end{align*}
Assuming $X_2^T M_1 X_2$ is invertible, we get the same estimator
\[
\hat\beta_2 = (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
\]
}

\begin{theorem}
Consider regression
\begin{equation} \label{eq:fwl_1}
y = X_1 \beta_1 + X_2 \beta_2 + u
\end{equation}
where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
Consider another regresison
\begin{equation}  \label{eq:fwl_2}
M_1 y = M_1 X_2 \beta_2 + M_1 u
\end{equation}
where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
\end{theorem}

\begin{proof}
Geometrical proof will be presented for the following model:
\begin{equation} \label{eq:fwl_proof}
y_i = \beta_1 x_i + \beta_2 z_i + u_i
\end{equation}

We start with regression `all-at-once' and will distinct its coefficients with index $(1)$. The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_1_regression_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_1_regression_lin}}
\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
\end{center}
\end{figure}

As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.

\begin{equation}\label{eq:fwl_2_y_clean}
\begin{aligned}
y &= \alpha z + \varepsilon \\
\hat\alpha &= \frac{y^T z}{z^T z} \\
\tilde{y} &= \hat\varepsilon = y - \frac{y^T z}{z^T z}z
\end{aligned}
\end{equation}

Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.

\begin{equation}\label{eq:fwl_2_x_clean}
\begin{aligned}
x &= \gamma z + \nu \\
\hat\gamma &= \frac{x^T z}{z^T z} \\
\tilde{x} &= \hat\nu = x - \frac{x^T z}{z^T z}z
\end{aligned}
\end{equation}

Geometric results of these two steps are presented in~\ref{fig:fwl_2_regression_first}.

Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_2_regression_first} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_2_regression_trans}}
\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$; \subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
\end{center}
\end{figure}

Now, let us picture all the results in one figure and mark some main points.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
\label{fig:fwl_3_3d} }
\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
%\caption{$Lin(x, z)$}
\label{fig:fwl_3_lin}}
\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$, C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
\end{center}
\end{figure}

In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$ and $\hat\beta_1^{(2)}\tilde x$ respectively,
while segments AC and BC represent $x$ and $\tilde{x}$. Having two congruent angles, triangles ABC and FHC are simillar.
Then, it follows:
\[
\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
\]
\end{proof}

%\section{Geometric properties of distributions}

%\section{Partial correlation}

\end{document}

### 2018-01-04 term paper_2018/paper.tex
correlation in simple lin reg

@@
 \DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\sVar}{sVar}
 \DeclareMathOperator{\card}{card}
@@
 \DeclareMathOperator{\Corr}{Corr}
+\DeclareMathOperator{\sCov}{sCov}
+\DeclareMathOperator{\sCorr}{sCorr}
 \DeclareMathOperator{\E}{E}
@@
 \]
-Note that from the definition of the angle it follows that correlation is limited within an interval $[-1,1]$.
+Note that from the definition of the angle it follows that correlation can range from $-1$ to $1$.
 
@@
 The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for random vectors. The projection of a random vector $Y$ onto $\{cX| c \in \mathbb{R}\}$ is $\Corr(X,Y) \cdot Y$.
+Same holds for random vectors. The projection of a random vector $Y$ onto $\{cX| c \in \mathbb{R}\}$ is $\hat Y = \Corr(X,Y) \cdot Y$.
+
+\begin{figure}[ht!]
+\begin{center}
+\includegraphics{images/corr_proj.pdf}
+\label{fig:corr_proj}
+\caption{Random vector $Y$ projected onto random vector $X$.}
+\end{center}
+\end{figure}
+
+Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of correlation coefficient.
+Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can write
+\[
+\Corr^2(X,Y) = \frac{\Var(\hat Y)}{\Var(Y)}
+\]
+as the variance of a random vecor is associated with the square of its length.
+Thus, the correlation coefficient squared shows the fraction of variance in $Y$ which can be explained
+with the most similar random variable proportional to $X$.
+
+\subsection{Sample correlation coefficient in simple linear regression}
+
+\marginnote{
+Assuming the underlying relationship between $x$ and $y$ to be
+\[
+y_i = \beta_1 + \beta_2 x_i + \varepsilon_i \quad i=1,\ldots,n
+\]
+where $\varepsilon_i$ is an error term the following holds
+\begin{align*}
+\sCorr(y, \hat y) &= \frac{\sCov(y)\sCov(\hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}} \\
+&= \frac{\sCov(y)\sCov(\hat \beta_1 + \hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_1 + \hat \beta_2 x)}} \\
+&= \frac{\sCov(y)\sCov(\hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_2 x)}} \\
+&= \frac{\hat \beta_2 \sCov(y)\sCov(x)}{|\hat \beta_2| \sqrt{\sVar(y)\sVar(x)}} \\
+&= sign(\hat \beta_2) \frac{\sCov(y)\sCov(x)}{\sqrt{\sVar(y)\sVar(x)}}
+\end{align*}
+}
+
+\begin{theorem}
+A linear regression model with one explanatory variable and constant term has the property
+\[
+\sCorr(y, \hat y) = sign(\hat \beta_2) \sCorr(y, x)
+\]
+\end{theorem}
+
+\begin{proof}
+Firstly, we consider the case when $\hat \beta_2 > 0$ so the main picuture is of the form depicted in Figure~\ref{fig:corr_pos_base}.
+It has been shown earlier that the correlation coefficient squared represents the angle betweem two random vectors.
+
+\begin{figure}[ht!]
+\begin{center}
+\includegraphics{images/corr_eq_positive.pdf}
+\label{fig:corr_pos_base}
+\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
+\end{center}
+\end{figure}
+
+However, it seems to be difficult to compare the angles in the three dimensional space.
+That is why we start with projecting both $x$ and $y$ onto the space perpendicular to the vector of all ones $\mathbf{1}$ as shown in Figure~\ref{fig:corr_pos_xyc}.
+We denote this space as $Lin^{\perp}(\mathbf{1})$. The resulting vectors are $x - \bar x \cdot \mathbf{1}$  and $y - \bar y \cdot \mathbf{1}$ respectively
+since projection of any vector $\vec{a}$ on the line given by a vector of all ones yields the vector of averages $\vec{\bar a}$.
+
+In order to get the angle between $y$ and $\hat y$ we should start with regressing $y$ on $Lin(x, \mathbf{1})$.
+Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
+The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/corr_eq_positive_xy_centered.pdf}
+\label{fig:corr_pos_xyc}}
+\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/corr_eq_positive_yhat.pdf}
+\label{fig:corr_pos_yhatc}}
+\caption{\subref{fig:corr_pos_xyc}: `Centred' $x$ and $y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$; \subref{fig:corr_pos_yhatc}: `Centred' $\hat y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$.}
+\end{center}
+\end{figure}
+
+todo: whole picture, $\beta_2<0$
+
+\end{proof}
 
@@
 \includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
 \label{fig:averages_lin} }
@@
 \includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
-%\caption{$Lin(x, z)$}
 \label{fig:averages_bars}}

### 2018-01-05 2018/images/pic2.tex
todo corr done

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}

%Angle Definitions
%-----------------

%start tikz picture, and use the tdplot_main_coords style to implement the display
%coordinate transformation provided by 3dplot
\tdplotsetmaincoords{40}{130}

\begin{tikzpicture}[scale=5,tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (xc) at (1,0);
\coordinate (yc) at (0,2);
\coordinate (un) at (0,0,4);

\tdplotsetcoord{S}{2}{30}{0}  % икс
\tdplotsetcoord{P}{4}{30}{90} % игрек
\tdplotsetcoord{Z}{3.4641}{0}{0} % 3.4641 = \sqrt{4^2 - 2^2}

%\draw[color=white, fill=gray!10] (xc) to[out=90,in=90] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
%\draw[color=white, fill=blue!10] (yc) to[out=90,in=90] (P) to[out=90,in=0] (un) -- (O) -- cycle;

\draw[color=white, fill=gray!5] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
\draw[color=white, fill=blue!5] (P) to[out=90,in=0] (un) -- (O) -- cycle;

%\draw[color=white, fill=gray!10, pattern=dots] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
%\draw[color=white, fill=blue!10, pattern=north east lines] (P) to[out=90,in=0] (un) -- (O) -- cycle;

\draw[thick,->] (O) -- (xc) node[anchor=north east]{$x - \bar{x} \cdot \mathbf{1}$};
\draw[thick,->] (O) -- (yc) node[anchor=north west]{$y - \bar{y} \cdot \mathbf{1}$};
\draw[thick,->] (O) -- (un) node[anchor=south]{$\mathbf{1}$};

\tdplotdrawarc{(O)}{0.1}{0}{90}{anchor=north}{$\phi$}

\draw[-stealth,thick] (O) -- (S) node[anchor=north east]{$x$};
\draw[-stealth,thick] (O) -- (P) node[anchor=north west]{$y$};

\draw[dashed] (S) -- (Sx);
\draw[dashed] (P) -- (Py);

\tkzMarkRightAngle[size=0.05](P,yc,O);
\tkzMarkRightAngle[size=0.05](S,xc,O);

\end{tikzpicture}

\end{document}

### 2018-01-05 2018/paper.tex
todo corr done

--- 2018-01-04 term paper_2018/paper.tex
@@
 \includegraphics{images/corr_def.pdf}
+\caption{Random vector $X$ of length $\sigma_X$ and random vector $Y$ of length $\sigma_Y$, $\cos \varphi$ is the angle between $X$ and $Y$.}
 \label{fig:corr_def}
-\caption{Random vector $X$ of length $\sigma_X$ and random vector $Y$ of length $\sigma_Y$, $\cos \varphi$ is the angle between $X$ and $Y$.}
 \end{center}
@@
 \includegraphics{images/corr_proj.pdf}
+\caption{Random vector $Y$ projected onto random vector $X$.}
 \label{fig:corr_proj}
-\caption{Random vector $Y$ projected onto random vector $X$.}
 \end{center}
@@
 Firstly, we consider the case when $\hat \beta_2 > 0$ so the main picuture is of the form depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient squared represents the angle betweem two random vectors.
+It has been shown earlier that the correlation coefficient represents the angle betweem two random vectors.
+So in order to complete the proof we need to find the appropriate angles and compare them.
 
-\begin{figure}[ht!]
-\begin{center}
-\includegraphics{images/corr_eq_positive.pdf}
+\begin{figure}[h!]
+\center{\includegraphics{images/corr_eq_positive.pdf}}
+\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
 \label{fig:corr_pos_base}
-\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
-\end{center}
 \end{figure}
@@
 Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
+Note that the projected $\hat y$ falls onto tha span of vector $x - \bar x \cdot \mathbf{1}$ as it can be decomposed into a sum $a x + b \mathbf{1}$ where $a, b \in \mathbb{R}$.
+$a x$ is projected in the same way as $x$ and $b \mathbf{1}$ yields zero when projected onto the orthogonal space.
 The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
@@
 
-todo: whole picture, $\beta_2<0$
+Since the projection of $\hat y$ lies exactly on the span of vector $x - \bar x \cdot \mathbf{1}$, we can conclude that $cos \varphi = \cos \varphi '$ and to put it another way $\sCorr(x,y) = \sCorr(y, \hat y)$.
 
+Now consider the case when $\hat \beta_2 < 0$.
+Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
+The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ not on this vector itself while the projections of $x$ and $y$ remain the same.
+Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
+Using trigonometric properties, we simplify $\cos(180^{\circ} - \varphi) = -\cos\varphi$ which in turn implies $\sCorr(x,y) = -\sCorr(y,\hat y)$.
+
+\begin{figure}[ht!]
+\begin{center}
+\includegraphics{images/corr_eq_negative.pdf}
+\label{fig:corr_negative}
+\caption{Case of $\beta_2 < 0$.}
+\end{center}
+\end{figure}
 \end{proof}

### 2018-01-06 2018/paper.tex
rss ess tss

@@
 
-\begin{figure}[ht!]
+\begin{figure}[h!]
 \begin{center}
@@
 \begin{theorem}
-A linear regression model with one explanatory variable and constant term has the property
+A linear regression model with one explanatory variable and constant term
+\[
+y = \beta_1 + \beta_2 x + \varepsilon
+\]
+has the property
 \[
@@
 \end{proof}
+
+\subsection{RSS + ESS = TSS}
+
+\marginnote{
+Consider a regresion model with $n$ observations and $k$ explanatory variables including a constant unit vector
+\[
+y = X \beta + \varepsilon
+\]
+The OLS estimator for the vector of coefficients $\beta$ is
+\[
+\hat \beta = (X^T X)^{-1} X^T y
+\]
+and the residual vector is
+\begin{align*}
+\hat e &= y - \hat y \\
+&= y - X \hat \beta \\
+&= y - X (X^T X)^{-1} X^T y
+\end{align*}
+Then we define residual sum of squares (RSS), explained sum of squares (ESS) and total sum of squares (TSS) as follows:
+\begin{align*}
+RSS &= \lVert y - \hat y \rVert^2_2 \\
+ESS &= \lVert \hat y - \bar y \rVert^2_2 \\
+TSS &= \lVert y - \bar y \rVert \\
+\end{align*}
+Disclosing parentheses and using the fact that $\hat{y}^T y = \hat{y}^T \hat{y}$
+\begin{align*}
+\hat{y}^T y &= \beta^T X^T y \\
+&=  y^T X (X^T X)^{-1} X^T y \\
+\hat{y}^T \hat{y} &= \beta^T X^T X \beta \\
+&= y^T X (X^T X)^{-1} X^T X (X^T X)^{-1} X^T y \\
+&= y^T X (X^T X)^{-1} X^T y
+\end{align*}
+we obtain
+\begin{align*}
+RSS &= y^T y -\hat{y}^T \hat{y} \\
+ESS &= \hat{y}^T \hat{y} - \hat{y}^T \bar{y} + \bar{y}^T \bar{y} \\
+TSS &= y^T y - 2 y^T \bar y +  \bar{y}^T \bar{y}
+\end{align*}
+When putting everything together all the terms cancel out which proves
+\[
+ESS + RSS = TSS
+\]
+}
+
+\begin{theorem}
+A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
+\[
+y = X \beta + \varepsilon
+\]
+has the following property
+\[
+RSS + ESS = TSS
+\]
+where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert$.
+\end{theorem}
+
+\begin{proof}
+The proof will be presented for the case of two regressor $x$ and $\mathbf{1}$ in order for the picture to be clear.
+However, the same logic applies for the case of $k$ regressors.
+
+We start with depicting the vectors $y \in \mathbb{R}^{n-2}$ and $x, \mathbf{1} \in \mathbb{R}^2$.
+Then we project $y$ onto $Lin(x, \mathbf{1})$ and obtain $\hat y$ which is shown in Figure~\ref{fig:tss_base}.
+
+From this picture we can immediately derive $\sqrt{RSS}$ as by definition this is the squared difference between $y$ and $\hat y$.
+
+\begin{figure}[h!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
+\label{fig:tss_base} }
+\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
+\label{fig:tss_rss}}
+\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
+\end{center}
+\end{figure}
+
+So as to visualize $ESS$ and $TSS$ we first need to visualize vector of averages $\bar y$.
+Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
+It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
+\[
+\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \ldots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
+\]
+
+Now we both project $y$ and $\hat y$ onto $\mathbf{1}$ and following the definition obtain $\sqrt{TSS}$ as the difference vector $y - \bar y$ and $\sqrt{ESS}$ as the vector $\hat y - \bar y$.
+
+\begin{figure}[h!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
+\label{fig:tss_tss}}
+\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
+\label{fig:tss_ess}}
+\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
+\end{center}
+\end{figure}
+
+The final step is to put everything together.
+Note that since $y - \hat y$ is perpendicular to $lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $lin(x, \mathbf{1})$.
+Then, applying the theorem of three perpendiculars we conclude that vecotr foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
+Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the hypothenuse $\sqrt{TSS}$:
+\[
+\sqrt{RSS} + \sqrt{ESS} = \sqrt{TSS}
+\]
+
+\begin{figure}[h!]
+\center{\includegraphics{images/tss_sum_final.pdf}}
+\caption{$\sqrt{RSS} + \sqrt{ESS} = \sqrt{TSS}$}
+\label{fig:tss_final}
+\end{figure}
+
+\end{proof}
+
+\subsection{Determination coefficient}
 

### 2018-01-07 2018/images/corr_const_br_up.tex
corr and r^2

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}

%Angle Definitions
%-----------------

%start tikz picture, and use the tdplot_main_coords style to implement the display
%coordinate transformation provided by 3dplot
\tdplotsetmaincoords{40}{120}

\begin{tikzpicture}[tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (xc) at (2.1,-2.1);
\coordinate (xcp) at (0.84,-0.84);
\coordinate (yc) at (-0.9,3);
\coordinate (un) at (0,0,6);
\coordinate (unt) at (0,0,2.9);

\tdplotsetcoord{S}{10}{35}{0}  % икс
\tdplotsetcoord{R}{4}{35}{0} % разложение
\tdplotsetcoord{P}{7}{30}{90} % игрек
\tdplotsetcoord{T}{6.5}{20}{0} % икс с тильдой

\draw[->] (O) -- (un) node[anchor=east]{$\mathbf{1}$};
\draw[->] (O) -- (S) node[anchor=south east]{$x$};
\draw[->, gray] (O) -- (P) node[anchor=north west, gray]{$y$};
\draw[->] (O) -- (T) node[anchor=south east]{$\tilde x$};

\draw (0,0.5) ellipse (4cm and 1cm) node[anchor=west, right=2.7cm]{\tiny$Lin^{\perp}(\mathbf{1})$};

\draw[->, gray] (O) -- (yc) node[anchor=north east]{\small $y^c$};
\draw[->] (O) -- (xc) node[anchor=north east]{\small $x^c$};
\draw[dashed, gray] (P) -- (yc);
\draw[dashed] (S) -- (xc);

\draw[->, dashed, blue] (unt) -- (T);
\draw[->, dashed, red] (R) -- (T);
\draw[->, blue] (O) -- (R) node[anchor=north east]{\footnotesize$\beta x$};
\draw[->, red] (O) -- (unt) node[anchor=west]{\footnotesize$\alpha \mathbf{1}$};

\draw[dashed, blue] (R) -- (xcp);
\draw[->,dashed,blue] (O) -- (xcp) node[anchor=north]{\small$\tilde x^c$};
\tkzMarkRightAngle[size=0.15,color=blue](R,xcp,O);
\tkzDrawPoint[color=red,fill=red,size=1.8](O);

\tkzMarkRightAngle[size=0.15,color=gray](P,yc,O);
\tkzMarkRightAngle[size=0.15](S,xc,O);
\tdplotdrawarc{(O)}{0.15}{-45}{110}{anchor=north}{\tiny$\varphi$}

\end{tikzpicture}
\end{document}

### 2018-01-07 2018/images/corr_def.tex
corr and r^2

--- 2018-01-07 2018/images/corr_const_br_up.tex
@@
 \usetkzobj{all}
+\usetikzlibrary{calc}
 \usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
@@
 \begin{document}
+\begin{tikzpicture}
 
-%Angle Definitions
-%-----------------
+\tkzDefPoint(0,0){O}
+\tkzDefPoint(4,0){X}
+\tkzDefPoint(2,3){Y}
 
-%start tikz picture, and use the tdplot_main_coords style to implement the display
-%coordinate transformation provided by 3dplot
-\tdplotsetmaincoords{40}{120}
+\draw[->] (O) -- (X) node[anchor=north east]{$x$};
+\draw[->] (O) -- (Y) node[anchor=east]{$y$};
 
-\begin{tikzpicture}[tdplot_main_coords]
-
-\coordinate (O) at (0,0,0);
-\coordinate (xc) at (2.1,-2.1);
-\coordinate (xcp) at (0.84,-0.84);
-\coordinate (yc) at (-0.9,3);
-\coordinate (un) at (0,0,6);
-\coordinate (unt) at (0,0,2.9);
-
-\tdplotsetcoord{S}{10}{35}{0}  % икс
-\tdplotsetcoord{R}{4}{35}{0} % разложение
-\tdplotsetcoord{P}{7}{30}{90} % игрек
-\tdplotsetcoord{T}{6.5}{20}{0} % икс с тильдой
-
-\draw[->] (O) -- (un) node[anchor=east]{$\mathbf{1}$};
-\draw[->] (O) -- (S) node[anchor=south east]{$x$};
-\draw[->, gray] (O) -- (P) node[anchor=north west, gray]{$y$};
-\draw[->] (O) -- (T) node[anchor=south east]{$\tilde x$};
-
-\draw (0,0.5) ellipse (4cm and 1cm) node[anchor=west, right=2.7cm]{\tiny$Lin^{\perp}(\mathbf{1})$};
-
-\draw[->, gray] (O) -- (yc) node[anchor=north east]{\small $y^c$};
-\draw[->] (O) -- (xc) node[anchor=north east]{\small $x^c$};
-\draw[dashed, gray] (P) -- (yc);
-\draw[dashed] (S) -- (xc);
-
-\draw[->, dashed, blue] (unt) -- (T);
-\draw[->, dashed, red] (R) -- (T);
-\draw[->, blue] (O) -- (R) node[anchor=north east]{\footnotesize$\beta x$};
-\draw[->, red] (O) -- (unt) node[anchor=west]{\footnotesize$\alpha \mathbf{1}$};
-
-\draw[dashed, blue] (R) -- (xcp);
-\draw[->,dashed,blue] (O) -- (xcp) node[anchor=north]{\small$\tilde x^c$};
-\tkzMarkRightAngle[size=0.15,color=blue](R,xcp,O);
-\tkzDrawPoint[color=red,fill=red,size=1.8](O);
-
-\tkzMarkRightAngle[size=0.15,color=gray](P,yc,O);
-\tkzMarkRightAngle[size=0.15](S,xc,O);
-\tdplotdrawarc{(O)}{0.15}{-45}{110}{anchor=north}{\tiny$\varphi$}
+\tkzMarkAngle[size=0.4](X,O,Y)
+\tkzLabelAngle[pos=0.6](X,O,Y){$\varphi$}
 

### 2018-01-07 2018/paper.tex
corr and r^2

@@
 
-\subsection{Correlation}
+\subsection{Sample correlation}
 
-\marginnote{The most widespread definition of correlation is
-\[
-\Corr(X,Y)  = \frac{\Cov(X, Y)}{\sqrt{\Var(X)\Var(Y)}}
-\]}
+\marginnote{
+\begin{align*}
+\sCorr(x,y) &= \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} \\
+&= \frac{\frac{1}{n-1}\sum_{i=1}^n (x_i - \bar x)(y_i - \bar y)}{\sqrt{\frac{1}{n-1} \sum_{i=1}^n (y_i - \bar y)\frac{1}{n-1} \sum_{i=1}^n (\hat y_i - \bar{\hat y})}}
+\end{align*}
+}
 
@@
 \]
-Now we define scalar product of two random vectors as covariation between them:
+Now we define scalar product of two vectors
+$x =
+\begin{pmatrix}
+x_1 \\
+\vdots \\
+x_n
+\end{pmatrix}$
+and
+$y =
+\begin{pmatrix}
+y_1 \\
+\vdots \\
+y_n
+\end{pmatrix}$
+as a sample covariation between them:
 \[
-\langle X, Y \rangle = \Cov(X, Y)
+\langle x, y \rangle = \sCov(x, y)
 \]
-The main characteristics of a random vector are its length and direction.
+The main characteristics of a vector are its length and direction.
 So, we introduce the length
 \[
-\sqrt{\Cov{X,X}} = \sqrt{\Var(X)} = \sigma_X
+\sqrt{\sCov(x,x)} = \sqrt{\sVar(x)} = \sigma_x
 \]
@@
 \[
-\cos(X,Y) = \frac{\Cov(X,Y)}{\sqrt{\Var(X)\Var(Y)}} = \Corr(X,Y)
+\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
 \]
-Note that from the definition of the angle it follows that correlation can range from $-1$ to $1$.
+Note that from the definition of the angle it follows that the sample correlation can range from $-1$ to $1$.
 
@@
 \includegraphics{images/corr_def.pdf}
-\caption{Random vector $X$ of length $\sigma_X$ and random vector $Y$ of length $\sigma_Y$, $\cos \varphi$ is the angle between $X$ and $Y$.}
+\caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
 \label{fig:corr_def}
@@
 The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for random vectors. The projection of a random vector $Y$ onto $\{cX| c \in \mathbb{R}\}$ is $\hat Y = \Corr(X,Y) \cdot Y$.
+Same holds for the vectors sampled from some distribution.
+The projection of such a vector $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
 
@@
 \includegraphics{images/corr_proj.pdf}
-\caption{Random vector $Y$ projected onto random vector $X$.}
+\caption{Vector $y$ projected onto vector $x$.}
 \label{fig:corr_proj}
@@
 
-Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of correlation coefficient.
-Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can write
+Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of sample correlation coefficient.
+Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can conclude that
 \[
-\Corr^2(X,Y) = \frac{\Var(\hat Y)}{\Var(Y)}
+\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)}
 \]
-as the variance of a random vecor is associated with the square of its length.
-Thus, the correlation coefficient squared shows the fraction of variance in $Y$ which can be explained
-with the most similar random variable proportional to $X$.
+as the variance of a vector is associated with the square of its length.
+Thus, the sample correlation coefficient squared shows the fraction of variance in $y$ which can be explained
+with the most similar vector proportional to $x$.
+
+\subsection{Sample correlation when a constatnt vector added}
+
+\marginnote{
+\begin{align*}
+\sCorr(x + \alpha \mathbf{1}, y) &= \frac{\sCov(x + \alpha \mathbf{1}, y)}{\sVar(x + \alpha \mathbf{1}) \sVar(y)} \\
+&= \frac{\sCov(x,y)\sCov(\alpha \mathbf{1},y)}{\sVar(x)\sVar(y)} \\
+&= \frac{\sCov(x,y)}{\sVar(x)\sVar(y)} \\
+&= \sCorr(x,y)
+\end{align*}
+}
+
+\begin{theorem}
+Adding a vector of constants does not affect the sample correlation coefficient:
+\[
+\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
+\]
+where $\alpha \in \mathbb{R}$.
+\end{theorem}
+
+\begin{proof}
+Firstly, we project vectors $x$ and $y$ onto $Lin^{\perp}(\mathbf{1})$ in order to get $x^c = x - \bar x$ and $y^c = y - \bar y$
+(`c' stands for `centred'). It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
+\[
+\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \ldots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
+\]
+Thus, projecting onto an orthogonal subspace is equivalent to substracting the projected vector, i.e., the vector of averages, from the original one.
+
+Also note that the angle $\varphi$ between original and centred vectors remains the same.
+The results of this step is shown in Figure~\ref{fig:corr_xyc}.
+
+\begin{figure}
+  \centering
+  \includegraphics{images/corr_const_cent.pdf}
+  \caption{Centred vectors $x^c$ and $y^c$}
+  \label{fig:corr_xyc}
+\end{figure}
+
+Then we need to derive a new vector $\tilde x$ with constants added to each component.
+Geometrically adding a vector of costants means adding a vector of all ones scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
+Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and $\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
+After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
+By the properties of projection it is of no difference whether to project the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$ and $\beta x$ — the result is the same.
+So, while $\beta x$ is projected onto a span of $x^c$, the projection of $\alpha \mathbf{1}$ onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated in Figure~\ref{fig:corr_tildex}.
+Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
+
+\begin{figure}
+  \centering
+  \includegraphics{images/corr_const_br_up.pdf}
+  \caption{Decomposition and projection of $\tilde x$}
+  \label{fig:corr_tildex}
+\end{figure}
+
+Finally, putting everything together we finish the proof:
+\[
+\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
+\]
+
+\begin{figure}
+  \centering
+  \includegraphics{images/corr_const_final.pdf}
+  \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
+  \label{fig:corr_final}
+\end{figure}
+
+\end{proof}
 
@@
 Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
-The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ not on this vector itself while the projections of $x$ and $y$ remain the same.
+The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ and not on this vector itself while the projections of $x$ and $y$ remain the same.
 Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
@@
 \]
-where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert$.
+where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert^2_2$.
 \end{theorem}
@@
 Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
-It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
-\[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \ldots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
-\]
 
@@
 The final step is to put everything together.
-Note that since $y - \hat y$ is perpendicular to $lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $lin(x, \mathbf{1})$.
-Then, applying the theorem of three perpendiculars we conclude that vecotr foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
+Note that since $y - \hat y$ is perpendicular to $Lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $Lin(x, \mathbf{1})$.
+Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
 Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the hypothenuse $\sqrt{TSS}$:
@@
 \center{\includegraphics{images/tss_sum_final.pdf}}
-\caption{$\sqrt{RSS} + \sqrt{ESS} = \sqrt{TSS}$}
+\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
 \label{fig:tss_final}
@@
 
+\marginnote{
+\begin{align*}
+\sCorr^2(y,\hat y) &= \left(\frac{\sCov(y, \hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}}\right)^2 \\
+&= \frac{\sCov(y, \hat y) \sCov(y, \hat y)}{\sVar(y)\sVar(\hat y)} \\
+&= \frac{\sCov(\hat y + e, \hat y) \sCov(\hat y + e, \hat y)}{\sVar(y)\sVar(\hat y)} \\
+&= \frac{(\sCov(\hat y, \hat y) + \sCov(e, \hat y))(\sCov(\hat y, \hat y) + \sCov(e, \hat y))}{\sVar(y)\sVar(\hat y)} \\
+&= \frac{\sVar(\hat y) \sVar(\hat y)}{\sVar(y)\sVar(\hat y)} \\
+&= \frac{\sVar(\hat y)}{\sVar(y)} \\
+&= \frac{ESS}{TSS} \\
+&= R^2
+\end{align*}
+}
+
+\begin{theorem}
+A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
+\[
+y = X \beta + \varepsilon
+\]
+has the following property
+\[
+R^2 = \sCorr^2(y, \hat y)
+\]
+\end{theorem}
+
+\begin{proof}
+Proving this theorem geometrically means showing that the determination coefficient can be interpreted as some squared angle
+which happens to be eqaul to the squared angle betwen $y$ and $\hat y$.
+
+Consider Figure~\ref{fig:tss_final} from the previous proof.
+It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
+Having defined the determination coefficient as
+\[
+R^2 = \frac{ESS}{TSS}
+\]
+we conclude that its geometric interpretaion is
+\[
+\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
+\]
+as shown in Figure~\ref{fig:r_sq_angle}.
+
+\begin{figure}
+  \center{\includegraphics{images/r_sq_ss.pdf}}
+  \caption{Determination coefficient as squared $\cos \varphi$}
+  \label{fig:r_sq_angle}
+\end{figure}
+
+Recall that the sample correlation coefficient two vectors was defined earlier as the angle between these two vectors.
+Thus, we conclude that $\sCorr(y, \hat y)$ is the angle between $y$ and $\hat y$ which is also eqaul to $\cos \varphi$.
+Finally, squaring both sides, we obtain
+\[
+R^2 = \sCorr^2(y, \hat y)
+\]
+\end{proof}
+
 \subsection{Regression line and point of averages}
@@
 \begin{align*}
-\hat y  &= X \hat \beta = X (X^T X)^{-1} X^T y \\
+\hat y  &= X \hat \beta \\
+&= X (X^T X)^{-1} X^T y \\
 &= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon

### 2018-01-07 2018/paper.tex
minor upd

@@
 Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
-Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the hypothenuse $\sqrt{TSS}$:
+Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the
+hypotenuse $\sqrt{TSS}$:
 \[
-\sqrt{RSS} + \sqrt{ESS} = \sqrt{TSS}
+(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2
 \]

### 2018-01-08 2018/paper.tex
partial correlation

@@
 \DeclareMathOperator{\sVar}{sVar}
+\DeclareMathOperator{\pVar}{pVar}
 \DeclareMathOperator{\card}{card}
@@
 \DeclareMathOperator{\sCov}{sCov}
+\DeclareMathOperator{\pCov}{pCov}
 \DeclareMathOperator{\sCorr}{sCorr}
+\DeclareMathOperator{\pCorr}{pCorr}
 \DeclareMathOperator{\E}{E}
@@
 \newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
 
@@
 \begin{center}
-\includegraphics{images/corr_def.pdf}
+\includegraphics[width=0.4\linewidth]{images/corr_def.pdf}
 \caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
@@
 \begin{center}
-\includegraphics{images/corr_proj.pdf}
+\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
 \caption{Vector $y$ projected onto vector $x$.}
@@
 \[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \ldots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
+\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \vdots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
 \]
@@
 
-\begin{figure}
+\begin{figure}[h!]
   \centering
-  \includegraphics{images/corr_const_cent.pdf}
+  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
   \caption{Centred vectors $x^c$ and $y^c$}
@@
 Then we need to derive a new vector $\tilde x$ with constants added to each component.
-Geometrically adding a vector of costants means adding a vector of all ones scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
-Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and $\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
+Geometrically adding a vector of costants means adding a vector of all ones
+scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
+Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
+$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
 After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
-By the properties of projection it is of no difference whether to project the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$ and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto a span of $x^c$, the projection of $\alpha \mathbf{1}$ onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated in Figure~\ref{fig:corr_tildex}.
+By the properties of projection it is of no difference whether to project
+the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
+and $\beta x$ — the result is the same.
+So, while $\beta x$ is projected onto a span of $x^c$, the projection of $\alpha \mathbf{1}$
+onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
+in Figure~\ref{fig:corr_tildex}.
 Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
@@
 \begin{figure}[h!]
-\center{\includegraphics{images/corr_eq_positive.pdf}}
+\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
 \caption{Vectors $x$, $y$ and $\mathbf{1}$.}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/corr_eq_positive_xy_centered.pdf}
+\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
 \label{fig:corr_pos_xyc}}
-\hspace{4ex}
+%\hspace{4ex}
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/corr_eq_positive_yhat.pdf}
+\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
 \label{fig:corr_pos_yhatc}}
@@
 \begin{theorem}
-The point of averages lies on the estimated regression line.
+In a linear regression model with one explanatory variable and constant term
+\[
+y = \beta_1 + \beta_2 x + \varepsilon
+\]
+the point of averages lies on the estimated regression line.
 \end{theorem}
@@
 \label{fig:fwl_2_regression_trans}}
-\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$; \subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
+\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
+\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
 \end{center}
@@
 \label{fig:fwl_3_lin}}
-\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$, C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
+\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
+C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
 F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
@@
 
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$ and $\hat\beta_1^{(2)}\tilde x$ respectively,
-while segments AC and BC represent $x$ and $\tilde{x}$. Having two congruent angles, triangles ABC and FHC are simillar.
+In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
+and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
+Having two congruent angles, triangles ABC and FHC are simillar.
 Then, it follows:
@@
 
-%\section{Partial correlation}
+\section{Partial correlation}
+
+\subsection{Definition of partial correlation}
+
+\marginnote{
+Partial correlation is the measure of degree of dependence between two random
+variables while controlling for the effect of other random variables:
+\[
+\pCorr(X,Y; Z) = \frac{\pCov(X,Y;Z)}{\sqrt{\pVar(X;Z)\pVar(Y;Z)}}
+\]
+where $\pVar(X;Z) = \Var(X - \alpha Z)$, $\alpha$ is such a constant that $\Cov(X - \alpha Z, Z) = 0$,
+and $\pCov(X,Y; Z) = \Cov(X - \alpha Z, Y - \beta Z)$, $\alpha$, $\beta$ are such
+constants that $\Cov(X - \alpha Z, Z) = \Cov(Y - \beta Z, Z) = 0$
+}
+
+Partial correlation can be defined into two ways.
+We will provide both definitions and show their equivalence.
+
+\begin{definition}
+Partial correlation between random variables $X$ and $Y$ holding random variable $Z$
+fixed is the correlation coefficient between the residuals in regression of $X$ onto
+$Z$ and the residuals in regression of $Y$ onto $Z$.
+\end{definition}
+
+Firstly, we project random variable $X$ onto $Z$, which yields $\E(X)$.
+The residuals in this regression are $X - \E(X)$ — a vector in $Lin^{\perp}(Z)$.
+We will call this variable `centered' and label it as $\widetilde X$.
+Repeating this step for $Y$ yields `centerd' variable $\tilde Y = Y - \E(Y) \in Lin^{\perp}(Z)$.
+The angle between $\widetilde X$ and $\widetilde Y$ ($\varphi$ in Figure~\ref{fig:pcorr_def1})
+is the correaletion coefficient between these `centred' random variables and
+the partial correlaiton between the original ones.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
+  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
+  \label{fig:pcorr_def1}
+\end{figure}
+
+\begin{definition}
+Partial correlaiton between random variables $X$ and $Y$ holding random variable $Z$
+fixed is the geometric mean between the coefficient $\hat \beta_{XY}$ in regression
+\[
+\hat X = \hat \beta_{XY} Y + \hat \beta_{XZ} Z
+\]
+and the coefficinent $\hat \beta_{YX}$ in regression
+\[
+\hat Y = \hat \beta_{YX} X + \hat \beta_{YZ} Z
+\]
+which has the same sign as the coefficients $\hat \beta_{XY}$ and $\hat \beta_{YX}$.
+\end{definition}
+
+Following the definition, we need to start with regressing variable $X$ onto
+$Y$ and $Z$. Then, the vector we obtained $\hat X$ can be broken up into the sum of
+$\hat \beta_{XY} Y$ and $\hat \beta_{XZ} Z$.
+Projecting $\hat \beta_{XY} Y$ onto $Lin^{\perp}(Z)$ we results in a vector
+$\alpha \widetilde Y$ where $\widetilde Y = Y - \E(Y)$ is the projection of $Y$ onto $Lin^{\perp}(Z)$.
+
+By the properties of similar triangles
+\[
+\frac{\beta_{XY} Y}{Y} = \frac{\alpha \widetilde Y}{\widetilde Y} \Leftrightarrow \beta_{XY} = \alpha
+\]
+
+In the same way we perform a regression of $Y$ onto $X$ and $Z$ and repeat the
+same steps as for $X$. Finally, we get the whole picture:
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
+\label{fig:part_corr_alt}}
+%\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
+\label{fig:part_corr_alt_lin}}
+\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
+\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
+\end{center}
+\end{figure}
+
+Having plotted $Lin^{\perp}(Z)$ now we can express $\cos \varphi$ in terms of $\beta_{XY}$
+and $\beta_{YX}$
+
+\begin{equation}\label{eq:part_cor_cos}
+\begin{split}
+\cos \varphi &= \frac{\vert \beta_{XY} \widetilde Y \vert}{\vert \widetilde X \vert} = \vert \beta_{XY} \vert \\
+\cos \varphi &= \frac{\vert \beta_{YX} \widetilde X \vert}{\vert \widetilde Y \vert} = \vert \beta_{YX} \vert \\
+\cos^2 \varphi &= \vert \beta_{XY} \beta_{YX} \vert \stackrel{sign(\beta_{XY}) = sign(\beta_{YX})}{=} \beta_{XY} \beta_{YX}
+\end{split}
+\end{equation}
+
+Recall that the angle $\varphi$ can be interpreted as the partial correlation
+between $X$ and $Y$ holding $Z$ fixed, so it follows form equaitions~\eqref{eq:part_cor_cos}
+
+\[
+\pCorr^2(X,Y; Z) = \cos^2 \varphi = \beta_{XY} \beta_{YX}
+\]
 

### 2018-01-11 2018/paper.tex
partial corr theorem

@@
 and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto a span of $x^c$, the projection of $\alpha \mathbf{1}$
+So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
 onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
@@
 \begin{proof}
-Firstly, we consider the case when $\hat \beta_2 > 0$ so the main picuture is of the form depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient represents the angle betweem two random vectors.
+Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
+picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
+It has been shown earlier that the correlation coefficient represents the angle
+betweem two random vectors.
 So in order to complete the proof we need to find the appropriate angles and compare them.
@@
 
+\subsection{Partial correlation as correlaiton between residuals}
+
+\marginnote{
+???
+}
+
+\begin{theorem}
+Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
+correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
+\[
+X = \alpha_1 Y + \alpha_2 Z + u
+\]
+and the residuals $\hat v$ in the model
+\[
+Y = \beta_1 Z + \beta_2 Z + v
+\]
+\end{theorem}
+
+\begin{proof}
+The first step is to find the residuals in the mentioned regresisions.
+For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
+which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
+and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
+in Figure~\ref{fig:pcorr_t_x}.
+
+Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
+\label{fig:pcorr_t_x}}
+%\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
+\label{fig:pcorr_t_y}}
+\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
+\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
+\end{center}
+\end{figure}
+
+After all we put these figure together and the goal is to find the angle between
+the red and blue lines which is the same as the angle between the dashed
+red and blue lines. However, before that we need to apply translating to them to
+get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
+complementary angles, angles in triangles and parallelograms
+we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
+is eqaul to $\varphi$.
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
+\label{fig:pcorr_t_xy}}
+%\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
+\label{fig:pcorr_t_lin}}
+\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
+\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
+\end{center}
+\end{figure}
+
+\end{proof}
+
 \end{document}

### 2018-01-14 2018/paper.tex
t,f-tests

@@
 \DeclareMathOperator{\E}{E}
+\DeclareMathOperator{\ctg}{ctg}
 \renewcommand{\P}{\mathbb{P}}
@@
 \end{figure}
-
 \end{proof}
 
+\section{Probability distributions}
+
+\subsection{Normal}
+
+\subsection{Chi-squared}
+
+\subsection{Student's}
+
+\subsection{t-test}
+
+In a simple linear regression model
+\[
+y = \beta_1 + \beta_2 x + \varepsilon
+\]
+the adjusted t-value $\frac{t}{\sqrt{n-2}}$ when $H_0: \beta_2 = 0$ is tested
+can be expreesed in terms of the angle between $y$ and $\hat y$ $\varphi$ and
+is equal to $\ctg \varphi$.
+
+Recall that the t-statistic is defined in the following way:
+\[
+t = \frac{\hat \beta - \beta}{s.e.(\hat\beta)}
+\]
+Adjusting this formula for the null hypothesis $H_0: \beta_2 = 0$, we obtain
+\begin{equation}\label{eq:tstat}
+t = \frac{\hat \beta_2}{s.e.(\hat\beta_2)}
+\end{equation}
+Then, we need to express $s.e.(\hat\beta_2)$ in terms of vectors which can be
+plotted. From standard OLS procedure it follows that
+\begin{equation}\label{eq:varbeta2}
+\Var(\hat \beta_2) = \frac{\sigma}{\sum_{i=1}^n (x_i - \bar x)^2}
+\end{equation}
+Since actual $\sigma$ is unknown the estimator will be used instead:
+\begin{equation}\label{eq:sigmaest}
+\hat \sigma^2 = \frac{RSS}{n-2}
+\end{equation}
+Substituting \eqref{eq:varbeta2} and \eqref{eq:sigmaest} into \eqref{eq:tstat}
+divided by $\sqrt{n-2}$, we obtain
+\begin{align*}
+\frac{t}{\sqrt{n-2}} &= \frac{\hat \beta_2}{\sqrt{n-2}\widehat{s.e.}(\hat\beta_2)} \\
+&= \frac{\hat \beta_2}{\sqrt{n-2}\frac{\hat \sigma}{\sqrt{\sum_{i=1}^n (x_i - \bar x)^2}}} \\
+&= \frac{\hat \beta_2 \sqrt{\sum_{i=1}^n (x_i - \bar x)^2}}{\sqrt{n-2}\frac{\sqrt{\sum_{i=1}^n (y_i - \hat y_i)^2}}{\sqrt{n-2}}} \\
+&= \frac{\hat \beta_2 \vert x^c \vert}{\sqrt{RSS}}
+\end{align*}
+where $ \vert x^c \vert = \sqrt{\sum_{i=1}^n (x_i - \bar x)^2}$ is the length of the
+centred vector $x$.
+
+Now the result can be demonstrated visually.
+Again we will project $x$ and $y$ vectors onto the $Lin^{\perp}(\mathbf{1})$ so as to
+get their centred versions $x^c$ and $y^c$.
+Then, we perform regression of $y$ onto $Lin(x, \mathbf{1})$ which results in $\hat y$.
+Following that, we project $\hat y = \hat \beta_1 + \hat \beta_2 x$ onto $Lin^{\perp}(\mathbf{1})$
+which yields $\hat \beta_2 x^c$.
+After all, we translate $\sqrt{RSS}$ onto $Lin^{\perp}(\mathbf{1})$.
+These steps are demonstrated in Figure~\ref{fig:ttest_3d}.
+
+Looking at Figure~\ref{fig:ttest_lin} which depicts the $Lin^{\perp}(\mathbf{1})$,
+we derive
+\[
+\ctg \varphi = \frac{\hat \beta_2 \vert x^c \vert}{\sqrt{RSS}} = t
+\]
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/ttest.pdf}
+\label{fig:ttest_3d}}
+%\hspace{4ex}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{images/ttest_lin.pdf}
+\label{fig:ttest_lin}}
+\caption{\subref{fig:pcorr_t_x}: Regression of $y$ onto $Lin(x, \mathbf{1})$ and appropriate projections;
+\subref{fig:pcorr_t_y}: $Lin^{\perp}(\mathbf{1})$.}
+\end{center}
+\end{figure}
+
+\subsection{F-distribution}
+
+\subsection{F-test}
+
+The significance of several coefficients at once can be tested with the F-test.
+The F-statistic has the following form
+\[
+F = \frac{(RSS_{R} - RSS_{UR})/q}{RSS_{UR}/(n-k_{UR})}
+\]
+where indices $R$ and $UR$ stand for the restricted and unrestricted models
+respectively, $n$ — number of observations, $k$ — number of regressors,
+$q$ — number of equtions used in the null hypothesis.
+
+Due to plotting limitations, we consider the unrestricted model to be
+\[
+y = \beta_1 + \beta_2 x + u
+\]
+and the restricted model to be
+\[
+y = \alpha_1 + v
+\]
+Note that there was a choice in the restricted models.
+
+We perform both regressions in order to get the ressiduals and plot them
+in Figure~\ref{fig:ftest}.
+Adjusted to the degrees of freedom, the ratio can be expressed in terms of the
+angle between two vectors, $\varphi$, as demonstrated in Figure~\ref{fig:ftest}
+\[
+F = \frac{(RSS_{R} - RSS_{UR})/q}{RSS_{UR}/(n-k_{UR})} = \ctg^2 \varphi
+\]
+
+\begin{figure}
+\centering
+\includegraphics[width=0.55\linewidth]{images/ftest.pdf}
+\caption{F-statistic as the cotangens squared of $\varphi$}
+\label{fig:ftest}
+\end{figure}
+
 \end{document}

### 2018-01-31 2018/images/f_dist_def.tex
geometry of r.v.s

--- 2018-01-05 2018/images/pic2.tex
@@
 
-\begin{tikzpicture}[scale=5,tdplot_main_coords]
+\begin{tikzpicture}[tdplot_main_coords]
 
 \coordinate (O) at (0,0,0);
-\coordinate (xc) at (1,0);
-\coordinate (yc) at (0,2);
 \coordinate (un) at (0,0,4);
+\coordinate (nun) at (0,0,-2.2);
+\coordinate (hun) at (0,0,3.2);
+\coordinate (z) at (0,3,5);
+\coordinate (hz) at (-1.05,2.05,0);
 
-\tdplotsetcoord{S}{2}{30}{0}  % икс
-\tdplotsetcoord{P}{4}{30}{90} % игрек
-\tdplotsetcoord{Z}{3.4641}{0}{0} % 3.4641 = \sqrt{4^2 - 2^2}
+\draw[->] (O) -- (un) node[anchor=east]{$L_1$};
+\draw[dashed] (O) -- (nun);
+\draw (0,0) ellipse (4.5cm and 1.5cm) node[anchor=east, left=3.2cm]{$L_2$};
 
-%\draw[color=white, fill=gray!10] (xc) to[out=90,in=90] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
-%\draw[color=white, fill=blue!10] (yc) to[out=90,in=90] (P) to[out=90,in=0] (un) -- (O) -- cycle;
-
-\draw[color=white, fill=gray!5] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
-\draw[color=white, fill=blue!5] (P) to[out=90,in=0] (un) -- (O) -- cycle;
-
-%\draw[color=white, fill=gray!10, pattern=dots] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
-%\draw[color=white, fill=blue!10, pattern=north east lines] (P) to[out=90,in=0] (un) -- (O) -- cycle;
-
-\draw[thick,->] (O) -- (xc) node[anchor=north east]{$x - \bar{x} \cdot \mathbf{1}$};
-\draw[thick,->] (O) -- (yc) node[anchor=north west]{$y - \bar{y} \cdot \mathbf{1}$};
-\draw[thick,->] (O) -- (un) node[anchor=south]{$\mathbf{1}$};
-
-\tdplotdrawarc{(O)}{0.1}{0}{90}{anchor=north}{$\phi$}
-
-\draw[-stealth,thick] (O) -- (S) node[anchor=north east]{$x$};
-\draw[-stealth,thick] (O) -- (P) node[anchor=north west]{$y$};
-
-\draw[dashed] (S) -- (Sx);
-\draw[dashed] (P) -- (Py);
-
-\tkzMarkRightAngle[size=0.05](P,yc,O);
-\tkzMarkRightAngle[size=0.05](S,xc,O);
+\draw[->] (O) -- (z) node[anchor=south]{$z$};
+\draw[->] (O) -- (hz) node[anchor=north]{$H_2 z$};
+\draw[->] (O) -- (hun) node[anchor=east]{$H_1 z$};
+\draw[dashed] (hz) -- (z);
+\draw[dashed] (z) -- (hun);
+\tkzMarkRightAngle[size=0.15](z,hz,O);
+\tkzMarkRightAngle[size=0.15](O,hun,z);
 
 \end{tikzpicture}
-
 \end{document}

### 2018-01-31 2018/paper.tex
geometry of r.v.s

@@
 
+\section{Geometric interpretation of random variables}
+
+The fundamental part to start with is defining the geometric properties of random
+variables using some concepts of linear algebra.
+
+Consider the vector space which consists of all random variables with finite
+mean and variance.
+We will regard each point in this space (or vector that correponds to that point
+in terms of linear algebra) a random variable.
+We define the scalar product of two random variables $X$ and $Y$ to be
+\[
+\langle X, Y \rangle = \Cov(X,Y).
+\]
+It is of no difficulty to check that the definition satisfies the properties
+of scalar product assuming that $X$ and $Y$ are the same random variables
+if $\P(X=Y)=1$.
+
+Having defined the inner product, we are now able to introduce the squared
+length of a random variable $X$ which is
+\[
+\lVert X \rVert^2 = \langle X, X \rangle = \Cov(X,X) = \Var(X),
+\]
+so the length is simply the square root of this expression, i.e., the~standard
+deviation of $X$ ($\sigma_X$).
+
+Recall that for any non-random vectors $\vec{a}$ and $\vec{b}$ the angle
+between them is calculated with the formula
+\[
+\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}.
+\]
+The same applies for the random variables and it is already clear that
+two random variables are uncorrelated if and only if their scalar product
+equals to $0$. Additionally, it means that these two random variables
+are orthogonal in the vector space.
+
+The analogue for $cos(\vec{a}, \vec{b})$ in the vector space of all the random
+variables is the correlation between two of them:
+\[
+\Corr(X,Y) = \frac{\Cov(X,Y)}{\sqrt{\Var(X)\Var(Y)}} = \frac{\langle X, Y \rangle}{\sqrt{\lVert X \rVert^2 \lVert Y \rVert^2}}.
+\]
+From the equivalence of $\Corr(X,Y)$ to the $\cos(\vec{a}, \vec{b})$ it
+automatically follows that the correlation coefficient can range from $-1$ to $1$.
+
+A useful property of the geoemtry of random variables is that all the
+geometric theorems still hold. For instance, the Pythogorean theorem can
+be formulated as follows: if the ranadom variables $X$ and $Y$ are uncorrelated
+(which implies that they are orthogonal), then the variance of their sum equals
+the sum of their variances:
+\[
+\Var(X + Y) = \sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} = \Var(X) + \Var(Y).
+\]
+Translated to the non-random language, assumption of uncorrelatedness correspnds
+to the right triangle setting, the variance of the sum of two random variables
+stands for the hypotenuse squared and the sum of the variances is the sum of
+the legs squared.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.4\linewidth]{images/rv_pythagorean.pdf}
+\caption{The Pythagorean theorem for random variables $X$ and $Y$.}
+\label{fig:rv_pyth}
+\end{center}
+\end{figure}
+
+Another important geometric tool is projection.
+Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$
+can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
+The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
+Same holds for the random variables.
+The projection of such a random variable $Y$ onto $\{cX| c \in \mathbb{R}\}$ is
+$\hat Y = \Corr(X,Y) \cdot Y$.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.4\linewidth]{images/rv_proj.pdf}
+\caption{The projection of a random variable $Y$ onto the line spanned by
+a random varibale $X$.}
+\label{fig:rv_proj}
+\end{center}
+\end{figure}
+
+Note that the squared lengths of the leg adjacent to $\varphi$ and the
+hypotenuse are $\Var(\hat Y)$ and $\Var(Y)$.
+So, the Figure~\ref{fig:rv_proj} gives a useful expression for the correlation
+coefficient squared:
+\[
+\Corr^2(X,Y) = \frac{\Var(\hat Y)}{\Var(Y)}.
+\]
+
+\subsection{The law of iterated expectations}
+
+\marginnote{
+Here is the proof for the case when $X$ and $Y$ are both discrete. Let $\E(Y|X) = g(X)$.
+\begin{align*}
+\E(g(X)) &= \sum_x g(x) \P(X=x) \\
+&= \sum_x \left( \sum_y y \P(Y=y|X=x) \right) \P(X=x) \\
+&= \sum_x \sum_y y  \P(X=x) \P(Y=y|X=x)  \\
+&= \sum_y y \sum_x \P(X=x, Y=y) \\
+&= \sum_y y \P(Y=y) \\
+&= \E(Y)
+\end{align*}
+The proof in case of continous random variables is absolutely analogous.
+}
+
+\begin{theorem}
+For any random variable $X$ and $Y$,
+\[
+\E(\E(X|Y)) = \E(Y).
+\]
+\end{theorem}
+
+\begin{proof}
+
+Consider the vector space of all the random variables.  The random variable
+which can be described as functions $h(X)$ of $X$ form a subspace of that vector
+space, represented as a plane $\alpha$ in Figure~\ref{fig:adams}.
+Another subspace is a subspace of constants, denoted as a vector $\mathbf{1} \in \alpha$.
+
+In order to obtain $\E(Y|X)$, first, we need to project $Y$ onto the subspace
+corresponding to $X$. As a result of this step, we get $\E(Y|X)$ — the function
+of $X$ that predicts $Y$ the best. Next, projecting $\E(Y|X)$ onto the space
+of all constants, we end up with $E(Y)$.
+
+Notice that the vector $Y - \E(Y|X)$ (which is also called the residual) is
+perpendicular to the plane $\alpha$. Moreover, the vecotr $\E(Y|X) - E(Y)$ is
+perpendicular to the vector of constants $\mathbf{1}$. Thus, we can apply
+the theorem of three perpendiculars and conclude that the vector $Y - \E(Y)$ is
+also perpendicular to the vector of constants $\mathbf{1}$.
+
+So, we showed that the expectation of the random variable $Y$ can be obtained either
+in two steps or by its direct projection onto the subpace of constans.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.6\linewidth]{images/adams.pdf}
+\caption{The law of iterated expectations. Equivalence of the two-step projecttion
+and direct projection of $Y$ onto $\mathbf{1}$.}
+\label{fig:adams}
+\end{center}
+\end{figure}
+
+\end{proof}
+
+\subsection{MSE decomposition}
+
+\marginnote{
+\begin{align*}
+MSE(\hat \theta) &= \E((\hat \theta - \theta)^2) \\
+&= \E \left[ \left( \hat \theta - \E(\hat \theta) +  \E(\hat \theta) - \theta \right) \right] \\
+&= \E \left[ \left( \hat \theta - \E(\hat \theta) \right)^2 + 2 \left( (\hat \theta - \E(\hat \theta)) (\E(\hat \theta) - \theta ) \right) \right. \\
+&+ \left. \left( \E(\hat \theta) - \theta  \right)^2 \right] \\
+&= \E \left[\left( \hat \theta - \E(\hat \theta) \right)^2 \right] \\
+&+ 2 \E \left[\left( (\hat \theta - \E(\hat \theta)) (\E(\hat \theta) - \theta ) \right) \right] \\
+&+ \E \left[\left( \E(\hat \theta) - \theta  \right)^2 \right] \\
+&= \E \left[\left( \hat \theta - \E(\hat \theta) \right)^2 \right] \\
+&+ 2 (\E(\hat \theta - \E(\hat \theta))) \E(\hat \theta - \E(\hat \theta)) \\
+&+ \E \left[\left( \E(\hat \theta) - \theta  \right)^2 \right] \\
+&= \E \left[\left( \hat \theta - \E(\hat \theta) \right)^2 \right] + \E \left[\left( \E(\hat \theta) - \theta  \right)^2 \right] \\
+&= \Var(\hat \theta) + \E \left[\left( \E(\hat \theta) - \theta  \right)^2 \right]
+\end{align*}
+}
+
+\begin{theorem}
+The mean squared error of an estimator $\hat \theta$ with respect to an unknown
+parameter $\theta$ defined as $MSE(\hat \theta) = \E((\hat \theta - \theta)^2)$
+can be decomposed into the sum of the variance of the estimator and its squared bias:
+\[
+MSE(\hat \theta) = \Var(\hat \theta) + \E \left[\left( \E(\hat \theta) - \theta  \right)^2 \right]
+\]
+\end{theorem}
+
+\begin{proof}
+We start with a random variable $\theta$ and its estimate $\hat \theta$ in the
+vector space. We know that an unbiased estimator's projection would be exactly
+the vector representing $\theta$. However, in general it does not have to and
+Figure~JJJJ illustrates this case: the projection of the estimator falls onto
+the line spanned by the vector $\theta$.
+
+Connecting vectors $\theta$ and $\hat \theta$, we obtain the right triangle which
+legs are $\hat \theta - \E(\hat\theta)$, $\E(\hat\theta) - \theta$ and the
+hypotenuse $\hat \theta - \theta$.
+Applying the Pythagorean theorem, we finish the proof:
+\begin{align*}
+\lVert \hat \theta - \theta \rVert^2 &= \lVert \hat \theta - \E(\hat \theta) \rVert^2  + \lVert \E(\hat \theta) - \theta \rVert^2 \\
+\E((\hat \theta - \theta)^2) &= \E((\hat \theta - \E(\hat \theta))^2) + \E((\E(\hat \theta) - \theta)^2) \\
+MSE(\hat \theta) &= \Var(\hat \theta) + \E((\E(\hat \theta) - \theta)^2)
+\end{align*}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.6\linewidth]{images/mse_decomposed.pdf}
+\caption{Decomposition of mean squred error into the variance and the bias squared.}
+\label{fig:mse_decomposed}
+\end{center}
+\end{figure}
+
+\end{proof}
+
 \section{Regression}
 
-\subsection{Sample correlation}
+The concepts discussed in the following section could also be presented in
+random variables instead of sample ones. As the geometry of sample variables
+is almost of no difference comparing to the random ones, the logic of
+all the theorems is also the same.
+
+\subsection{Geometry of sample variables}
 
@@
 
-The most crucial part in defining correlation geometrically is definig the dot product as it enables to compute the length of a vecotr:
-\[
-|\vec{a}| = \sqrt{\langle \vec{a},  \vec{a}\rangle}
-\]
-and the angle between any two vectors:
-\[
-\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}
-\]
-Now we define scalar product of two vectors
+In the same manner which was performed in Section 1, we define
+the scalar product of two sample variables
 $x =
@@
 \[
-\langle x, y \rangle = \sCov(x, y)
+\langle x, y \rangle = \sCov(x, y).
 \]
 The main characteristics of a vector are its length and direction.
-So, we introduce the length
+Again, we introduce the length
 \[
@@
 \]
-and the angle between two random vectors
-\[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
+and the angle between two sample variables
+\[
+\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y).
 \]
@@
 
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_def.pdf}
-\caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
-\label{fig:corr_def}
-\end{center}
-\end{figure}
-
-Another important geometrical tool is projection.
-Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$ can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
-The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for the vectors sampled from some distribution.
-The projection of such a vector $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
+Completely analogus to the case of random variables,
+the projection of such a sample variable $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
 
@@
 \[
-\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)}
+\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)},
 \]
@@
 \subsection{Chi-squared}
+
+Consider a random vector $z \in \mathbb{R}^n$ which components are independent
+and follow standard normal distribution, $z_i \sim \mathcal{N}(0,1)$.
+Consider also a fixed $k$-dimensional subspace in $\mathbb{R}^n$ $L$.
+Let the projection of vector $z$ onto the subspace $L$ be $\hat z$ and
+its length squared $Q$
+\[
+Q = \lVert \hat z \rVert^2 = \langle \hat z, \hat z \rangle = \hat z^T \hat z
+\]
+Then $Q$ follows the chi-squared distribution with $k$ degrees of freedom.
 

### 2018-02-02 2018/chapters/1_random_variables.tex
restructured

--- 2018-01-31 2018/paper.tex
@@
-\documentclass[nobib]{tufte-handout}
-
-\usepackage{ifluatex, ifxetex} % проверка, как компилируется файл
-
-% black magic!
-% solving problem with tufte-handout + xelatex
-% http://tex.stackexchange.com/questions/200722/
-% https://tex.stackexchange.com/questions/47576/
-\ifx\ifxetex\ifluatex\else % если xelatex или lualatex
-  \newcommand{\textls}[2][5]{%
-    \begingroup\addfontfeatures{LetterSpace=#1}#2\endgroup
-  }
-  \renewcommand{\allcapsspacing}[1]{\textls[15]{#1}}
-  \renewcommand{\smallcapsspacing}[1]{\textls[10]{#1}}
-  \renewcommand{\allcaps}[1]{\textls[15]{\MakeTextUppercase{#1}}}
-  \renewcommand{\smallcaps}[1]{\smallcapsspacing{\scshape\MakeTextLowercase{#1}}}
-  \renewcommand{\textsc}[1]{\smallcapsspacing{\textsmallcaps{#1}}}
-\fi
-
-% с альтернативного источника:
-%\ifluatex % Allow rendering with LuaTeX by setting up the spacing using fontspec features
-%  \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
-%  \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
-%\fi
-
-% lua produces strange notes
-% https://tex.stackexchange.com/questions/328431
-
-\usepackage{fontspec} % работа со шрифтами
-%\usepackage{polyglossia} % учим русский как иностранный :)
-
-%\setmainlanguage{engish}
-%\setotherlanguages{english}
-
-% заменяем --- на тире, << на кавычки и т.д.:
-\defaultfontfeatures{Ligatures=TeX}
-
-% download "Linux Libertine" fonts:
-% http://www.linuxlibertine.org/index.php?id=91&L=1
-\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
-% why do we need \newfontfamily:
-% http://tex.stackexchange.com/questions/91507/
-\newfontfamily{\cyrillicfonttt}{Linux Libertine O}
-\newfontfamily{\cyrillicfont}{Linux Libertine O}
-\newfontfamily{\cyrillicfontsf}{Linux Libertine O}
-
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{url} % вставка \url{}
-\usepackage{graphicx} % вставка графиков
-\usepackage{csquotes} % адаптирующиеся кавычки командой \enquote{}
-\usepackage{comment} % ingore everything between \begin{comment} \end{comment}
-\usepackage{answers} % separate problems and solutions
-\usepackage{tikz} % pictures with tikz language
-\usepackage{todonotes} % todo in documents
-\usepackage{multicol} % make several columns
-\usepackage{subfigure}
-
-\usepackage{enumitem} % для создания своих нумерующих списков (хак для гиперссылок)
-
-%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-
-% very useful during de-bugging!
-% \usepackage[left]{showlabels}
-% \showlabels{hypertarget}
-% \showlabels{hyperlink}
-
-\DeclareMathOperator{\Var}{Var}
-\DeclareMathOperator{\sVar}{sVar}
-\DeclareMathOperator{\pVar}{pVar}
-\DeclareMathOperator{\card}{card}
-\DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\Corr}{Corr}
-\DeclareMathOperator{\sCov}{sCov}
-\DeclareMathOperator{\pCov}{pCov}
-\DeclareMathOperator{\sCorr}{sCorr}
-\DeclareMathOperator{\pCorr}{pCorr}
-\DeclareMathOperator{\E}{E}
-\DeclareMathOperator{\ctg}{ctg}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\I}{\mathbb{I}} % индикатор события
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-
-\usepackage[bibencoding = auto, backend = biber,
-sorting = none]{biblatex}
-
-%\addbibresource{probability_dna.bib}
-
-%\title{Geometric interpretaion of main concepts and theorems in econometrics}
-\title{some title}
-\author{}
-\date{\today}
-\begin{document}
-
-\maketitle
-
-%\section{Literature review}
-
 \section{Geometric interpretation of random variables}
@@
 
-Recall that for any non-random vectors $\vec{a}$ and $\vec{b}$ the angle
+Recall that for any non-random vectors $a$ and $b$ the angle
 between them is calculated with the formula
 \[
-\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}.
+\cos(a, b) = \frac{\langle a,  a\rangle}{|a| |b|}.
 \]
@@
 
-The analogue for $cos(\vec{a}, \vec{b})$ in the vector space of all the random
+The analogue for $cos(a, b)$ in the vector space of all the random
 variables is the correlation between two of them:
@@
 \]
-From the equivalence of $\Corr(X,Y)$ to the $\cos(\vec{a}, \vec{b})$ it
+From the equivalence of $\Corr(X,Y)$ to the $\cos(a, b)$ it
 automatically follows that the correlation coefficient can range from $-1$ to $1$.
@@
 \begin{center}
-\includegraphics[width=0.4\linewidth]{images/rv_pythagorean.pdf}
+\includegraphics[width=0.4\linewidth]{figures/01_pythagorean_theorem.pdf}
 \caption{The Pythagorean theorem for random variables $X$ and $Y$.}
@@
 Another important geometric tool is projection.
-Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$
-can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
-The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
+Recall that for any two vectors the scalar product $\langle a, b \rangle$
+can be interpreted as the length of projected $b$ multuplied by the length of $a$.
+The projection itself is $\cos(a, b) b$.
 Same holds for the random variables.
@@
 \begin{center}
-\includegraphics[width=0.4\linewidth]{images/rv_proj.pdf}
+\includegraphics[width=0.4\linewidth]{figures/01_basic_projection.pdf}
 \caption{The projection of a random variable $Y$ onto the line spanned by
@@
 \begin{center}
-\includegraphics[width=0.6\linewidth]{images/adams.pdf}
+\includegraphics[width=0.6\linewidth]{figures/01_law_of_iterated_expectations.pdf}
 \caption{The law of iterated expectations. Equivalence of the two-step projecttion
@@
 the vector representing $\theta$. However, in general it does not have to and
-Figure~JJJJ illustrates this case: the projection of the estimator falls onto
+Figure~\ref{fig:mse_decomposed} illustrates this case: the projection of the estimator falls onto
 the line spanned by the vector $\theta$.
@@
 \begin{center}
-\includegraphics[width=0.6\linewidth]{images/mse_decomposed.pdf}
+\includegraphics[width=0.6\linewidth]{figures/01_mse_decomposition.pdf}
 \caption{Decomposition of mean squred error into the variance and the bias squared.}
@@
 \end{proof}
-
-\section{Regression}
-
-The concepts discussed in the following section could also be presented in
-random variables instead of sample ones. As the geometry of sample variables
-is almost of no difference comparing to the random ones, the logic of
-all the theorems is also the same.
-
-\subsection{Geometry of sample variables}
-
-\marginnote{
-\begin{align*}
-\sCorr(x,y) &= \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} \\
-&= \frac{\frac{1}{n-1}\sum_{i=1}^n (x_i - \bar x)(y_i - \bar y)}{\sqrt{\frac{1}{n-1} \sum_{i=1}^n (y_i - \bar y)\frac{1}{n-1} \sum_{i=1}^n (\hat y_i - \bar{\hat y})}}
-\end{align*}
-}
-
-In the same manner which was performed in Section 1, we define
-the scalar product of two sample variables
-$x =
-\begin{pmatrix}
-x_1 \\
-\vdots \\
-x_n
-\end{pmatrix}$
-and
-$y =
-\begin{pmatrix}
-y_1 \\
-\vdots \\
-y_n
-\end{pmatrix}$
-as a sample covariation between them:
-\[
-\langle x, y \rangle = \sCov(x, y).
-\]
-The main characteristics of a vector are its length and direction.
-Again, we introduce the length
-\[
-\sqrt{\sCov(x,x)} = \sqrt{\sVar(x)} = \sigma_x
-\]
-and the angle between two sample variables
-\[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y).
-\]
-Note that from the definition of the angle it follows that the sample correlation can range from $-1$ to $1$.
-
-Completely analogus to the case of random variables,
-the projection of such a sample variable $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
-\caption{Vector $y$ projected onto vector $x$.}
-\label{fig:corr_proj}
-\end{center}
-\end{figure}
-
-Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of sample correlation coefficient.
-Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can conclude that
-\[
-\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)},
-\]
-as the variance of a vector is associated with the square of its length.
-Thus, the sample correlation coefficient squared shows the fraction of variance in $y$ which can be explained
-with the most similar vector proportional to $x$.
-
-\subsection{Sample correlation when a constatnt vector added}
-
-\marginnote{
-\begin{align*}
-\sCorr(x + \alpha \mathbf{1}, y) &= \frac{\sCov(x + \alpha \mathbf{1}, y)}{\sVar(x + \alpha \mathbf{1}) \sVar(y)} \\
-&= \frac{\sCov(x,y)\sCov(\alpha \mathbf{1},y)}{\sVar(x)\sVar(y)} \\
-&= \frac{\sCov(x,y)}{\sVar(x)\sVar(y)} \\
-&= \sCorr(x,y)
-\end{align*}
-}
-
-\begin{theorem}
-Adding a vector of constants does not affect the sample correlation coefficient:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-where $\alpha \in \mathbb{R}$.
-\end{theorem}
-
-\begin{proof}
-Firstly, we project vectors $x$ and $y$ onto $Lin^{\perp}(\mathbf{1})$ in order to get $x^c = x - \bar x$ and $y^c = y - \bar y$
-(`c' stands for `centred'). It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
-\[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \vdots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
-\]
-Thus, projecting onto an orthogonal subspace is equivalent to substracting the projected vector, i.e., the vector of averages, from the original one.
-
-Also note that the angle $\varphi$ between original and centred vectors remains the same.
-The results of this step is shown in Figure~\ref{fig:corr_xyc}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
-  \caption{Centred vectors $x^c$ and $y^c$}
-  \label{fig:corr_xyc}
-\end{figure}
-
-Then we need to derive a new vector $\tilde x$ with constants added to each component.
-Geometrically adding a vector of costants means adding a vector of all ones
-scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
-Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
-$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
-After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
-By the properties of projection it is of no difference whether to project
-the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
-and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
-onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
-in Figure~\ref{fig:corr_tildex}.
-Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_br_up.pdf}
-  \caption{Decomposition and projection of $\tilde x$}
-  \label{fig:corr_tildex}
-\end{figure}
-
-Finally, putting everything together we finish the proof:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_final.pdf}
-  \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
-  \label{fig:corr_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Sample correlation coefficient in simple linear regression}
-
-\marginnote{
-Assuming the underlying relationship between $x$ and $y$ to be
-\[
-y_i = \beta_1 + \beta_2 x_i + \varepsilon_i \quad i=1,\ldots,n
-\]
-where $\varepsilon_i$ is an error term the following holds
-\begin{align*}
-\sCorr(y, \hat y) &= \frac{\sCov(y)\sCov(\hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_1 + \hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_1 + \hat \beta_2 x)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_2 x)}} \\
-&= \frac{\hat \beta_2 \sCov(y)\sCov(x)}{|\hat \beta_2| \sqrt{\sVar(y)\sVar(x)}} \\
-&= sign(\hat \beta_2) \frac{\sCov(y)\sCov(x)}{\sqrt{\sVar(y)\sVar(x)}}
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-has the property
-\[
-\sCorr(y, \hat y) = sign(\hat \beta_2) \sCorr(y, x)
-\]
-\end{theorem}
-
-\begin{proof}
-Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
-picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient represents the angle
-betweem two random vectors.
-So in order to complete the proof we need to find the appropriate angles and compare them.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
-\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
-\label{fig:corr_pos_base}
-\end{figure}
-
-However, it seems to be difficult to compare the angles in the three dimensional space.
-That is why we start with projecting both $x$ and $y$ onto the space perpendicular to the vector of all ones $\mathbf{1}$ as shown in Figure~\ref{fig:corr_pos_xyc}.
-We denote this space as $Lin^{\perp}(\mathbf{1})$. The resulting vectors are $x - \bar x \cdot \mathbf{1}$  and $y - \bar y \cdot \mathbf{1}$ respectively
-since projection of any vector $\vec{a}$ on the line given by a vector of all ones yields the vector of averages $\vec{\bar a}$.
-
-In order to get the angle between $y$ and $\hat y$ we should start with regressing $y$ on $Lin(x, \mathbf{1})$.
-Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
-Note that the projected $\hat y$ falls onto tha span of vector $x - \bar x \cdot \mathbf{1}$ as it can be decomposed into a sum $a x + b \mathbf{1}$ where $a, b \in \mathbb{R}$.
-$a x$ is projected in the same way as $x$ and $b \mathbf{1}$ yields zero when projected onto the orthogonal space.
-The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
-\label{fig:corr_pos_xyc}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
-\label{fig:corr_pos_yhatc}}
-\caption{\subref{fig:corr_pos_xyc}: `Centred' $x$ and $y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$; \subref{fig:corr_pos_yhatc}: `Centred' $\hat y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$.}
-\end{center}
-\end{figure}
-
-Since the projection of $\hat y$ lies exactly on the span of vector $x - \bar x \cdot \mathbf{1}$, we can conclude that $cos \varphi = \cos \varphi '$ and to put it another way $\sCorr(x,y) = \sCorr(y, \hat y)$.
-
-Now consider the case when $\hat \beta_2 < 0$.
-Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
-The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ and not on this vector itself while the projections of $x$ and $y$ remain the same.
-Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
-Using trigonometric properties, we simplify $\cos(180^{\circ} - \varphi) = -\cos\varphi$ which in turn implies $\sCorr(x,y) = -\sCorr(y,\hat y)$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics{images/corr_eq_negative.pdf}
-\label{fig:corr_negative}
-\caption{Case of $\beta_2 < 0$.}
-\end{center}
-\end{figure}
-\end{proof}
-
-\subsection{RSS + ESS = TSS}
-
-\marginnote{
-Consider a regresion model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-The OLS estimator for the vector of coefficients $\beta$ is
-\[
-\hat \beta = (X^T X)^{-1} X^T y
-\]
-and the residual vector is
-\begin{align*}
-\hat e &= y - \hat y \\
-&= y - X \hat \beta \\
-&= y - X (X^T X)^{-1} X^T y
-\end{align*}
-Then we define residual sum of squares (RSS), explained sum of squares (ESS) and total sum of squares (TSS) as follows:
-\begin{align*}
-RSS &= \lVert y - \hat y \rVert^2_2 \\
-ESS &= \lVert \hat y - \bar y \rVert^2_2 \\
-TSS &= \lVert y - \bar y \rVert \\
-\end{align*}
-Disclosing parentheses and using the fact that $\hat{y}^T y = \hat{y}^T \hat{y}$
-\begin{align*}
-\hat{y}^T y &= \beta^T X^T y \\
-&=  y^T X (X^T X)^{-1} X^T y \\
-\hat{y}^T \hat{y} &= \beta^T X^T X \beta \\
-&= y^T X (X^T X)^{-1} X^T X (X^T X)^{-1} X^T y \\
-&= y^T X (X^T X)^{-1} X^T y
-\end{align*}
-we obtain
-\begin{align*}
-RSS &= y^T y -\hat{y}^T \hat{y} \\
-ESS &= \hat{y}^T \hat{y} - \hat{y}^T \bar{y} + \bar{y}^T \bar{y} \\
-TSS &= y^T y - 2 y^T \bar y +  \bar{y}^T \bar{y}
-\end{align*}
-When putting everything together all the terms cancel out which proves
-\[
-ESS + RSS = TSS
-\]
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-RSS + ESS = TSS
-\]
-where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert^2_2$.
-\end{theorem}
-
-\begin{proof}
-The proof will be presented for the case of two regressor $x$ and $\mathbf{1}$ in order for the picture to be clear.
-However, the same logic applies for the case of $k$ regressors.
-
-We start with depicting the vectors $y \in \mathbb{R}^{n-2}$ and $x, \mathbf{1} \in \mathbb{R}^2$.
-Then we project $y$ onto $Lin(x, \mathbf{1})$ and obtain $\hat y$ which is shown in Figure~\ref{fig:tss_base}.
-
-From this picture we can immediately derive $\sqrt{RSS}$ as by definition this is the squared difference between $y$ and $\hat y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
-\label{fig:tss_base} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
-\label{fig:tss_rss}}
-\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
-\end{center}
-\end{figure}
-
-So as to visualize $ESS$ and $TSS$ we first need to visualize vector of averages $\bar y$.
-Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
-
-Now we both project $y$ and $\hat y$ onto $\mathbf{1}$ and following the definition obtain $\sqrt{TSS}$ as the difference vector $y - \bar y$ and $\sqrt{ESS}$ as the vector $\hat y - \bar y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
-\label{fig:tss_tss}}
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
-\label{fig:tss_ess}}
-\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
-\end{center}
-\end{figure}
-
-The final step is to put everything together.
-Note that since $y - \hat y$ is perpendicular to $Lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $Lin(x, \mathbf{1})$.
-Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
-Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the
-hypotenuse $\sqrt{TSS}$:
-\[
-(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2
-\]
-
-\begin{figure}[h!]
-\center{\includegraphics{images/tss_sum_final.pdf}}
-\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
-\label{fig:tss_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Determination coefficient}
-
-\marginnote{
-\begin{align*}
-\sCorr^2(y,\hat y) &= \left(\frac{\sCov(y, \hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}}\right)^2 \\
-&= \frac{\sCov(y, \hat y) \sCov(y, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sCov(\hat y + e, \hat y) \sCov(\hat y + e, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{(\sCov(\hat y, \hat y) + \sCov(e, \hat y))(\sCov(\hat y, \hat y) + \sCov(e, \hat y))}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y) \sVar(\hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y)}{\sVar(y)} \\
-&= \frac{ESS}{TSS} \\
-&= R^2
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{theorem}
-
-\begin{proof}
-Proving this theorem geometrically means showing that the determination coefficient can be interpreted as some squared angle
-which happens to be eqaul to the squared angle betwen $y$ and $\hat y$.
-
-Consider Figure~\ref{fig:tss_final} from the previous proof.
-It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
-Having defined the determination coefficient as
-\[
-R^2 = \frac{ESS}{TSS}
-\]
-we conclude that its geometric interpretaion is
-\[
-\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
-\]
-as shown in Figure~\ref{fig:r_sq_angle}.
-
-\begin{figure}
-  \center{\includegraphics{images/r_sq_ss.pdf}}
-  \caption{Determination coefficient as squared $\cos \varphi$}
-  \label{fig:r_sq_angle}
-\end{figure}
-
-Recall that the sample correlation coefficient two vectors was defined earlier as the angle between these two vectors.
-Thus, we conclude that $\sCorr(y, \hat y)$ is the angle between $y$ and $\hat y$ which is also eqaul to $\cos \varphi$.
-Finally, squaring both sides, we obtain
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{proof}
-
-\subsection{Regression line and point of averages}
-
-\marginnote{
-If the regression contains the intercept, the following equation holds:
-\begin{align*}
-\hat y  &= X \hat \beta \\
-&= X (X^T X)^{-1} X^T y \\
-&= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon
-\end{align*}
-Premultiplying both sides by $X^T$, we obtain:
-\begin{align*}
-X^T \hat y &=  X^T X (X^T X)^{-1} X^T X \beta \\
-&+ X^T X (X^T X)^{-1} X^T \varepsilon \\
-&= X^T X \beta + X^T \varepsilon
-\end{align*}
-This is a system of equations. The first row of $X^T$ is $\mathbf{1}$ vector, so we can write out the first equation:
-\[
-\sum_{i=1}^n \hat y_i = \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-From the first equation in the system
-\[
-X^T \hat y = X^T y
-\]
-we obtain
-\[
-\sum_{i=1}^{n} \hat y_i = \sum_{i=1}^{n} y
-\]
-And this finishes the proof:
-\[
-\frac{1}{n} \sum_{i=1}^{n} y = \frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-}
-
-\begin{theorem}
-In a linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-the point of averages lies on the estimated regression line.
-\end{theorem}
-
-\begin{proof}
-For the geometrical proof it suffices to show that $\hat y$ is a linear combination of the regressors, which is true by construction,
-and that $\frac{1}{n} \sum_{i=1}^{n} \hat y_i = \frac{1}{n} \sum_{i=1}^{n} y$. In order for the pictures to be more clear the proof will be presented for the case of two regressors.
-
-The first step is regressing $y$ on $Lin(\mathbf{1}, x)$. As shown in Figure~\ref{fig:averages_lin}, we obtain $\hat y$ as a linear combination of $\mathbf{1}$ and $x$.
-The next step is to regress both $y$ and $\hat y$ on $\mathbf{1}$ which results in $\bar y$ and $\bar \hat y$ correspondingly.
-By the theorem of three perpendiculars, $\bar y = \bar \hat y$ which is shown in Figure~\ref{fig:averages_bars}.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
-\label{fig:averages_lin} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
-\label{fig:averages_bars}}
-\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
-\end{center}
-\end{figure}
-
-\end{proof}
-
-\subsection{Frisch–Waugh–Lovell theorem}
-
-\marginnote{From regresison~\ref{eq:fwl_2} we get the following estimator:
-\begin{align*}
-\hat\beta_2 &= ((M_1 X_2)^T M_1 X_2)^{-1}(M_1 X_2)^T M_1 y \\
-&= (X_2^T M_1^T M_1 X_2)^{-1}  X_2^T M_1^T M_1 y \\
-&= (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\end{align*}
-As for regresison~\ref{eq:fwl_1}, let us note that due to $y = \hat y + \hat u$ $y$ can be decomposed as follows:
-\[
-y = Py + My = X_1 \hat \beta_1 + X_2 \hat \beta_2 + My
-\]
-Premultiplying both sides by $X_2^T M_1$, we obtain:
-\begin{align*}
-X_2^T M_1 y &= X_2^T M_1 X_1 \hat\beta_1 + X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&=  X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&= X_2^T M_1 X_2 \hat\beta_2
-\end{align*}
-On the last step we used the fact that
-\begin{align*}
-(X_2^T M_1 M y)^T = y^T M^T M_1^T X_2 \\
-= y^T M M_1 X_2 = y^T M X_2 = 0^T
-\end{align*}
-Assuming $X_2^T M_1 X_2$ is invertible, we get the same estimator
-\[
-\hat\beta_2 = (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\]
-}
-
-\begin{theorem}
-Consider regression
-\begin{equation} \label{eq:fwl_1}
-y = X_1 \beta_1 + X_2 \beta_2 + u
-\end{equation}
-where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
-$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
-Consider another regresison
-\begin{equation}  \label{eq:fwl_2}
-M_1 y = M_1 X_2 \beta_2 + M_1 u
-\end{equation}
-where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
-Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
-\end{theorem}
-
-\begin{proof}
-Geometrical proof will be presented for the following model:
-\begin{equation} \label{eq:fwl_proof}
-y_i = \beta_1 x_i + \beta_2 z_i + u_i
-\end{equation}
-
-We start with regression `all-at-once' and will distinct its coefficients with index $(1)$.
-The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
-which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_1_regression_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_1_regression_lin}}
-\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
-\end{center}
-\end{figure}
-
-As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.
-
-\begin{equation}\label{eq:fwl_2_y_clean}
-\begin{aligned}
-y &= \alpha z + \varepsilon \\
-\hat\alpha &= \frac{y^T z}{z^T z} \\
-\tilde{y} &= \hat\varepsilon = y - \frac{y^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.
-
-\begin{equation}\label{eq:fwl_2_x_clean}
-\begin{aligned}
-x &= \gamma z + \nu \\
-\hat\gamma &= \frac{x^T z}{z^T z} \\
-\tilde{x} &= \hat\nu = x - \frac{x^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Geometric results of these two steps are presented in~\ref{fig:fwl_2_regression_first}.
-
-Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
-So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_2_regression_first} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_2_regression_trans}}
-\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
-\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
-\end{center}
-\end{figure}
-
-Now, let us picture all the results in one figure and mark some main points.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_3_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_3_lin}}
-\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
-C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
-F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
-\end{center}
-\end{figure}
-
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
-and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
-Having two congruent angles, triangles ABC and FHC are simillar.
-Then, it follows:
-\[
-\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
-\]
-\end{proof}
-
-%\section{Geometric properties of distributions}
-
-\section{Partial correlation}
-
-\subsection{Definition of partial correlation}
-
-\marginnote{
-Partial correlation is the measure of degree of dependence between two random
-variables while controlling for the effect of other random variables:
-\[
-\pCorr(X,Y; Z) = \frac{\pCov(X,Y;Z)}{\sqrt{\pVar(X;Z)\pVar(Y;Z)}}
-\]
-where $\pVar(X;Z) = \Var(X - \alpha Z)$, $\alpha$ is such a constant that $\Cov(X - \alpha Z, Z) = 0$,
-and $\pCov(X,Y; Z) = \Cov(X - \alpha Z, Y - \beta Z)$, $\alpha$, $\beta$ are such
-constants that $\Cov(X - \alpha Z, Z) = \Cov(Y - \beta Z, Z) = 0$
-}
-
-Partial correlation can be defined into two ways.
-We will provide both definitions and show their equivalence.
-
-\begin{definition}
-Partial correlation between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the correlation coefficient between the residuals in regression of $X$ onto
-$Z$ and the residuals in regression of $Y$ onto $Z$.
-\end{definition}
-
-Firstly, we project random variable $X$ onto $Z$, which yields $\E(X)$.
-The residuals in this regression are $X - \E(X)$ — a vector in $Lin^{\perp}(Z)$.
-We will call this variable `centered' and label it as $\widetilde X$.
-Repeating this step for $Y$ yields `centerd' variable $\tilde Y = Y - \E(Y) \in Lin^{\perp}(Z)$.
-The angle between $\widetilde X$ and $\widetilde Y$ ($\varphi$ in Figure~\ref{fig:pcorr_def1})
-is the correaletion coefficient between these `centred' random variables and
-the partial correlaiton between the original ones.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
-  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
-  \label{fig:pcorr_def1}
-\end{figure}
-
-\begin{definition}
-Partial correlaiton between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the geometric mean between the coefficient $\hat \beta_{XY}$ in regression
-\[
-\hat X = \hat \beta_{XY} Y + \hat \beta_{XZ} Z
-\]
-and the coefficinent $\hat \beta_{YX}$ in regression
-\[
-\hat Y = \hat \beta_{YX} X + \hat \beta_{YZ} Z
-\]
-which has the same sign as the coefficients $\hat \beta_{XY}$ and $\hat \beta_{YX}$.
-\end{definition}
-
-Following the definition, we need to start with regressing variable $X$ onto
-$Y$ and $Z$. Then, the vector we obtained $\hat X$ can be broken up into the sum of
-$\hat \beta_{XY} Y$ and $\hat \beta_{XZ} Z$.
-Projecting $\hat \beta_{XY} Y$ onto $Lin^{\perp}(Z)$ we results in a vector
-$\alpha \widetilde Y$ where $\widetilde Y = Y - \E(Y)$ is the projection of $Y$ onto $Lin^{\perp}(Z)$.
-
-By the properties of similar triangles
-\[
-\frac{\beta_{XY} Y}{Y} = \frac{\alpha \widetilde Y}{\widetilde Y} \Leftrightarrow \beta_{XY} = \alpha
-\]
-
-In the same way we perform a regression of $Y$ onto $X$ and $Z$ and repeat the
-same steps as for $X$. Finally, we get the whole picture:
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
-\label{fig:part_corr_alt}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
-\label{fig:part_corr_alt_lin}}
-\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
-\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-Having plotted $Lin^{\perp}(Z)$ now we can express $\cos \varphi$ in terms of $\beta_{XY}$
-and $\beta_{YX}$
-
-\begin{equation}\label{eq:part_cor_cos}
-\begin{split}
-\cos \varphi &= \frac{\vert \beta_{XY} \widetilde Y \vert}{\vert \widetilde X \vert} = \vert \beta_{XY} \vert \\
-\cos \varphi &= \frac{\vert \beta_{YX} \widetilde X \vert}{\vert \widetilde Y \vert} = \vert \beta_{YX} \vert \\
-\cos^2 \varphi &= \vert \beta_{XY} \beta_{YX} \vert \stackrel{sign(\beta_{XY}) = sign(\beta_{YX})}{=} \beta_{XY} \beta_{YX}
-\end{split}
-\end{equation}
-
-Recall that the angle $\varphi$ can be interpreted as the partial correlation
-between $X$ and $Y$ holding $Z$ fixed, so it follows form equaitions~\eqref{eq:part_cor_cos}
-
-\[
-\pCorr^2(X,Y; Z) = \cos^2 \varphi = \beta_{XY} \beta_{YX}
-\]
-
-\subsection{Partial correlation as correlaiton between residuals}
-
-\marginnote{
-???
-}
-
-\begin{theorem}
-Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
-correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
-\[
-X = \alpha_1 Y + \alpha_2 Z + u
-\]
-and the residuals $\hat v$ in the model
-\[
-Y = \beta_1 Z + \beta_2 Z + v
-\]
-\end{theorem}
-
-\begin{proof}
-The first step is to find the residuals in the mentioned regresisions.
-For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
-which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
-and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
-in Figure~\ref{fig:pcorr_t_x}.
-
-Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
-\label{fig:pcorr_t_x}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
-\label{fig:pcorr_t_y}}
-\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
-\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
-\end{center}
-\end{figure}
-
-After all we put these figure together and the goal is to find the angle between
-the red and blue lines which is the same as the angle between the dashed
-red and blue lines. However, before that we need to apply translating to them to
-get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
-complementary angles, angles in triangles and parallelograms
-we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
-is eqaul to $\varphi$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
-\label{fig:pcorr_t_xy}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
-\label{fig:pcorr_t_lin}}
-\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
-\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-\end{proof}
-
-\section{Probability distributions}
-
-\subsection{Normal}
-
-\subsection{Chi-squared}
-
-Consider a random vector $z \in \mathbb{R}^n$ which components are independent
-and follow standard normal distribution, $z_i \sim \mathcal{N}(0,1)$.
-Consider also a fixed $k$-dimensional subspace in $\mathbb{R}^n$ $L$.
-Let the projection of vector $z$ onto the subspace $L$ be $\hat z$ and
-its length squared $Q$
-\[
-Q = \lVert \hat z \rVert^2 = \langle \hat z, \hat z \rangle = \hat z^T \hat z
-\]
-Then $Q$ follows the chi-squared distribution with $k$ degrees of freedom.
-
-\subsection{Student's}
-
-\subsection{t-test}
-
-In a simple linear regression model
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-the adjusted t-value $\frac{t}{\sqrt{n-2}}$ when $H_0: \beta_2 = 0$ is tested
-can be expreesed in terms of the angle between $y$ and $\hat y$ $\varphi$ and
-is equal to $\ctg \varphi$.
-
-Recall that the t-statistic is defined in the following way:
-\[
-t = \frac{\hat \beta - \beta}{s.e.(\hat\beta)}
-\]
-Adjusting this formula for the null hypothesis $H_0: \beta_2 = 0$, we obtain
-\begin{equation}\label{eq:tstat}
-t = \frac{\hat \beta_2}{s.e.(\hat\beta_2)}
-\end{equation}
-Then, we need to express $s.e.(\hat\beta_2)$ in terms of vectors which can be
-plotted. From standard OLS procedure it follows that
-\begin{equation}\label{eq:varbeta2}
-\Var(\hat \beta_2) = \frac{\sigma}{\sum_{i=1}^n (x_i - \bar x)^2}
-\end{equation}
-Since actual $\sigma$ is unknown the estimator will be used instead:
-\begin{equation}\label{eq:sigmaest}
-\hat \sigma^2 = \frac{RSS}{n-2}
-\end{equation}
-Substituting \eqref{eq:varbeta2} and \eqref{eq:sigmaest} into \eqref{eq:tstat}
-divided by $\sqrt{n-2}$, we obtain
-\begin{align*}
-\frac{t}{\sqrt{n-2}} &= \frac{\hat \beta_2}{\sqrt{n-2}\widehat{s.e.}(\hat\beta_2)} \\
-&= \frac{\hat \beta_2}{\sqrt{n-2}\frac{\hat \sigma}{\sqrt{\sum_{i=1}^n (x_i - \bar x)^2}}} \\
-&= \frac{\hat \beta_2 \sqrt{\sum_{i=1}^n (x_i - \bar x)^2}}{\sqrt{n-2}\frac{\sqrt{\sum_{i=1}^n (y_i - \hat y_i)^2}}{\sqrt{n-2}}} \\
-&= \frac{\hat \beta_2 \vert x^c \vert}{\sqrt{RSS}}
-\end{align*}
-where $ \vert x^c \vert = \sqrt{\sum_{i=1}^n (x_i - \bar x)^2}$ is the length of the
-centred vector $x$.
-
-Now the result can be demonstrated visually.
-Again we will project $x$ and $y$ vectors onto the $Lin^{\perp}(\mathbf{1})$ so as to
-get their centred versions $x^c$ and $y^c$.
-Then, we perform regression of $y$ onto $Lin(x, \mathbf{1})$ which results in $\hat y$.
-Following that, we project $\hat y = \hat \beta_1 + \hat \beta_2 x$ onto $Lin^{\perp}(\mathbf{1})$
-which yields $\hat \beta_2 x^c$.
-After all, we translate $\sqrt{RSS}$ onto $Lin^{\perp}(\mathbf{1})$.
-These steps are demonstrated in Figure~\ref{fig:ttest_3d}.
-
-Looking at Figure~\ref{fig:ttest_lin} which depicts the $Lin^{\perp}(\mathbf{1})$,
-we derive
-\[
-\ctg \varphi = \frac{\hat \beta_2 \vert x^c \vert}{\sqrt{RSS}} = t
-\]
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/ttest.pdf}
-\label{fig:ttest_3d}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/ttest_lin.pdf}
-\label{fig:ttest_lin}}
-\caption{\subref{fig:pcorr_t_x}: Regression of $y$ onto $Lin(x, \mathbf{1})$ and appropriate projections;
-\subref{fig:pcorr_t_y}: $Lin^{\perp}(\mathbf{1})$.}
-\end{center}
-\end{figure}
-
-\subsection{F-distribution}
-
-\subsection{F-test}
-
-The significance of several coefficients at once can be tested with the F-test.
-The F-statistic has the following form
-\[
-F = \frac{(RSS_{R} - RSS_{UR})/q}{RSS_{UR}/(n-k_{UR})}
-\]
-where indices $R$ and $UR$ stand for the restricted and unrestricted models
-respectively, $n$ — number of observations, $k$ — number of regressors,
-$q$ — number of equtions used in the null hypothesis.
-
-Due to plotting limitations, we consider the unrestricted model to be
-\[
-y = \beta_1 + \beta_2 x + u
-\]
-and the restricted model to be
-\[
-y = \alpha_1 + v
-\]
-Note that there was a choice in the restricted models.
-
-We perform both regressions in order to get the ressiduals and plot them
-in Figure~\ref{fig:ftest}.
-Adjusted to the degrees of freedom, the ratio can be expressed in terms of the
-angle between two vectors, $\varphi$, as demonstrated in Figure~\ref{fig:ftest}
-\[
-F = \frac{(RSS_{R} - RSS_{UR})/q}{RSS_{UR}/(n-k_{UR})} = \ctg^2 \varphi
-\]
-
-\begin{figure}
-\centering
-\includegraphics[width=0.55\linewidth]{images/ftest.pdf}
-\caption{F-statistic as the cotangens squared of $\varphi$}
-\label{fig:ftest}
-\end{figure}
-
-\end{document}

### 2018-02-02 2018/chapters/2_regression.tex
restructured

--- 2018-01-11 2018/paper.tex
@@
-\documentclass[nobib]{tufte-handout}
-
-\usepackage{ifluatex, ifxetex} % проверка, как компилируется файл
-
-% black magic!
-% solving problem with tufte-handout + xelatex
-% http://tex.stackexchange.com/questions/200722/
-% https://tex.stackexchange.com/questions/47576/
-\ifx\ifxetex\ifluatex\else % если xelatex или lualatex
-  \newcommand{\textls}[2][5]{%
-    \begingroup\addfontfeatures{LetterSpace=#1}#2\endgroup
-  }
-  \renewcommand{\allcapsspacing}[1]{\textls[15]{#1}}
-  \renewcommand{\smallcapsspacing}[1]{\textls[10]{#1}}
-  \renewcommand{\allcaps}[1]{\textls[15]{\MakeTextUppercase{#1}}}
-  \renewcommand{\smallcaps}[1]{\smallcapsspacing{\scshape\MakeTextLowercase{#1}}}
-  \renewcommand{\textsc}[1]{\smallcapsspacing{\textsmallcaps{#1}}}
-\fi
-
-% с альтернативного источника:
-%\ifluatex % Allow rendering with LuaTeX by setting up the spacing using fontspec features
-%  \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
-%  \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
-%\fi
-
-% lua produces strange notes
-% https://tex.stackexchange.com/questions/328431
-
-\usepackage{fontspec} % работа со шрифтами
-%\usepackage{polyglossia} % учим русский как иностранный :)
-
-%\setmainlanguage{engish}
-%\setotherlanguages{english}
-
-% заменяем --- на тире, << на кавычки и т.д.:
-\defaultfontfeatures{Ligatures=TeX}
-
-% download "Linux Libertine" fonts:
-% http://www.linuxlibertine.org/index.php?id=91&L=1
-\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
-% why do we need \newfontfamily:
-% http://tex.stackexchange.com/questions/91507/
-\newfontfamily{\cyrillicfonttt}{Linux Libertine O}
-\newfontfamily{\cyrillicfont}{Linux Libertine O}
-\newfontfamily{\cyrillicfontsf}{Linux Libertine O}
-
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{url} % вставка \url{}
-\usepackage{graphicx} % вставка графиков
-\usepackage{csquotes} % адаптирующиеся кавычки командой \enquote{}
-\usepackage{comment} % ingore everything between \begin{comment} \end{comment}
-\usepackage{answers} % separate problems and solutions
-\usepackage{tikz} % pictures with tikz language
-\usepackage{todonotes} % todo in documents
-\usepackage{multicol} % make several columns
-\usepackage{subfigure}
-
-\usepackage{enumitem} % для создания своих нумерующих списков (хак для гиперссылок)
-
-%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-
-% very useful during de-bugging!
-% \usepackage[left]{showlabels}
-% \showlabels{hypertarget}
-% \showlabels{hyperlink}
-
-\DeclareMathOperator{\Var}{Var}
-\DeclareMathOperator{\sVar}{sVar}
-\DeclareMathOperator{\pVar}{pVar}
-\DeclareMathOperator{\card}{card}
-\DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\Corr}{Corr}
-\DeclareMathOperator{\sCov}{sCov}
-\DeclareMathOperator{\pCov}{pCov}
-\DeclareMathOperator{\sCorr}{sCorr}
-\DeclareMathOperator{\pCorr}{pCorr}
-\DeclareMathOperator{\E}{E}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\I}{\mathbb{I}} % индикатор события
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-
-\usepackage[bibencoding = auto, backend = biber,
-sorting = none]{biblatex}
-
-%\addbibresource{probability_dna.bib}
-
-%\title{Geometric interpretaion of main concepts and theorems in econometrics}
-\title{some title}
-\author{}
-\date{\today}
-\begin{document}
-
-\maketitle
-
-%\section{Literature review}
-
 \section{Regression}
 
-\subsection{Sample correlation}
+The concepts discussed in the following section could also be presented in
+random variables instead of sample ones. As the geometry of sample variables
+is almost of no difference comparing to the random ones, the logic of
+all the theorems is also the same.
+
+\subsection{Geometry of sample variables}
 
@@
 \]
-and the angle between two random vectors
+and the angle between two sample variables
 \[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
+\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y).
 \]
@@
 \begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_basic_projection.pdf}
 \caption{Vector $y$ projected onto vector $x$.}
@@
   \centering
-  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
+  \includegraphics[width=0.45\linewidth]{figures/02_correlation_constant_centered_variables.pdf}
   \caption{Centred vectors $x^c$ and $y^c$}
@@
   \centering
-  \includegraphics{images/corr_const_br_up.pdf}
+  \includegraphics{figures/02_correlation_constant_decomposition.pdf}
   \caption{Decomposition and projection of $\tilde x$}
@@
   \centering
-  \includegraphics{images/corr_const_final.pdf}
+  \includegraphics{figures/02_correlation_constant_proof.pdf}
   \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
@@
 \begin{figure}[h!]
-\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
+\center{\includegraphics[width=0.35\linewidth]{figures/02_simple_regression_coefficient_basic.pdf}}
 \caption{Vectors $x$, $y$ and $\mathbf{1}$.}
@@
 \subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
+\includegraphics[width=0.6\linewidth]{figures/02_simple_regression_coefficient_centred_variables.pdf}
 \label{fig:corr_pos_xyc}}
@@
 \subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
+\includegraphics[width=0.6\linewidth]{figures/02_simple_regression_coefficient_yhat_projected.pdf}
 \label{fig:corr_pos_yhatc}}
@@
 \begin{center}
-\includegraphics{images/corr_eq_negative.pdf}
+\includegraphics{figures/02_simple_regression_coefficient_negative.pdf}
 \label{fig:corr_negative}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
+\includegraphics[width=0.4\linewidth]{figures/tss_sum_base.pdf}
 \label{fig:tss_base} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
+\includegraphics[width=0.4\linewidth]{figures/tss_sum_rss.pdf}
 \label{fig:tss_rss}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
+\includegraphics[width=0.4\linewidth]{figures/tss_sum_tss.pdf}
 \label{fig:tss_tss}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
+\includegraphics[width=0.4\linewidth]{figures/tss_sum_ess.pdf}
 \label{fig:tss_ess}}
@@
 \begin{figure}[h!]
-\center{\includegraphics{images/tss_sum_final.pdf}}
+\center{\includegraphics{figures/tss_sum_final.pdf}}
 \caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
@@
 \begin{figure}
-  \center{\includegraphics{images/r_sq_ss.pdf}}
+  \center{\includegraphics{figures/r_sq_ss.pdf}}
   \caption{Determination coefficient as squared $\cos \varphi$}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
+\includegraphics[width=0.4\linewidth]{figures/averages_1.pdf}
 \label{fig:averages_lin} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
+\includegraphics[width=0.4\linewidth]{figures/averages_2.pdf}
 \label{fig:averages_bars}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression.pdf}
 %\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression_lin.pdf}
 %\caption{$Lin(x, z)$}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression.pdf}
 %\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression_translation.pdf}
 %\caption{$Lin(x, z)$}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_3_3d.pdf}
 %\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
+\includegraphics[width=0.4\linewidth]{figures/fwl_3_lin.pdf}
 %\caption{$Lin(x, z)$}
@@
 \end{proof}
-
-%\section{Geometric properties of distributions}
-
-\section{Partial correlation}
-
-\subsection{Definition of partial correlation}
-
-\marginnote{
-Partial correlation is the measure of degree of dependence between two random
-variables while controlling for the effect of other random variables:
-\[
-\pCorr(X,Y; Z) = \frac{\pCov(X,Y;Z)}{\sqrt{\pVar(X;Z)\pVar(Y;Z)}}
-\]
-where $\pVar(X;Z) = \Var(X - \alpha Z)$, $\alpha$ is such a constant that $\Cov(X - \alpha Z, Z) = 0$,
-and $\pCov(X,Y; Z) = \Cov(X - \alpha Z, Y - \beta Z)$, $\alpha$, $\beta$ are such
-constants that $\Cov(X - \alpha Z, Z) = \Cov(Y - \beta Z, Z) = 0$
-}
-
-Partial correlation can be defined into two ways.
-We will provide both definitions and show their equivalence.
-
-\begin{definition}
-Partial correlation between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the correlation coefficient between the residuals in regression of $X$ onto
-$Z$ and the residuals in regression of $Y$ onto $Z$.
-\end{definition}
-
-Firstly, we project random variable $X$ onto $Z$, which yields $\E(X)$.
-The residuals in this regression are $X - \E(X)$ — a vector in $Lin^{\perp}(Z)$.
-We will call this variable `centered' and label it as $\widetilde X$.
-Repeating this step for $Y$ yields `centerd' variable $\tilde Y = Y - \E(Y) \in Lin^{\perp}(Z)$.
-The angle between $\widetilde X$ and $\widetilde Y$ ($\varphi$ in Figure~\ref{fig:pcorr_def1})
-is the correaletion coefficient between these `centred' random variables and
-the partial correlaiton between the original ones.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
-  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
-  \label{fig:pcorr_def1}
-\end{figure}
-
-\begin{definition}
-Partial correlaiton between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the geometric mean between the coefficient $\hat \beta_{XY}$ in regression
-\[
-\hat X = \hat \beta_{XY} Y + \hat \beta_{XZ} Z
-\]
-and the coefficinent $\hat \beta_{YX}$ in regression
-\[
-\hat Y = \hat \beta_{YX} X + \hat \beta_{YZ} Z
-\]
-which has the same sign as the coefficients $\hat \beta_{XY}$ and $\hat \beta_{YX}$.
-\end{definition}
-
-Following the definition, we need to start with regressing variable $X$ onto
-$Y$ and $Z$. Then, the vector we obtained $\hat X$ can be broken up into the sum of
-$\hat \beta_{XY} Y$ and $\hat \beta_{XZ} Z$.
-Projecting $\hat \beta_{XY} Y$ onto $Lin^{\perp}(Z)$ we results in a vector
-$\alpha \widetilde Y$ where $\widetilde Y = Y - \E(Y)$ is the projection of $Y$ onto $Lin^{\perp}(Z)$.
-
-By the properties of similar triangles
-\[
-\frac{\beta_{XY} Y}{Y} = \frac{\alpha \widetilde Y}{\widetilde Y} \Leftrightarrow \beta_{XY} = \alpha
-\]
-
-In the same way we perform a regression of $Y$ onto $X$ and $Z$ and repeat the
-same steps as for $X$. Finally, we get the whole picture:
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
-\label{fig:part_corr_alt}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
-\label{fig:part_corr_alt_lin}}
-\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
-\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-Having plotted $Lin^{\perp}(Z)$ now we can express $\cos \varphi$ in terms of $\beta_{XY}$
-and $\beta_{YX}$
-
-\begin{equation}\label{eq:part_cor_cos}
-\begin{split}
-\cos \varphi &= \frac{\vert \beta_{XY} \widetilde Y \vert}{\vert \widetilde X \vert} = \vert \beta_{XY} \vert \\
-\cos \varphi &= \frac{\vert \beta_{YX} \widetilde X \vert}{\vert \widetilde Y \vert} = \vert \beta_{YX} \vert \\
-\cos^2 \varphi &= \vert \beta_{XY} \beta_{YX} \vert \stackrel{sign(\beta_{XY}) = sign(\beta_{YX})}{=} \beta_{XY} \beta_{YX}
-\end{split}
-\end{equation}
-
-Recall that the angle $\varphi$ can be interpreted as the partial correlation
-between $X$ and $Y$ holding $Z$ fixed, so it follows form equaitions~\eqref{eq:part_cor_cos}
-
-\[
-\pCorr^2(X,Y; Z) = \cos^2 \varphi = \beta_{XY} \beta_{YX}
-\]
-
-\subsection{Partial correlation as correlaiton between residuals}
-
-\marginnote{
-???
-}
-
-\begin{theorem}
-Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
-correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
-\[
-X = \alpha_1 Y + \alpha_2 Z + u
-\]
-and the residuals $\hat v$ in the model
-\[
-Y = \beta_1 Z + \beta_2 Z + v
-\]
-\end{theorem}
-
-\begin{proof}
-The first step is to find the residuals in the mentioned regresisions.
-For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
-which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
-and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
-in Figure~\ref{fig:pcorr_t_x}.
-
-Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
-\label{fig:pcorr_t_x}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
-\label{fig:pcorr_t_y}}
-\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
-\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
-\end{center}
-\end{figure}
-
-After all we put these figure together and the goal is to find the angle between
-the red and blue lines which is the same as the angle between the dashed
-red and blue lines. However, before that we need to apply translating to them to
-get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
-complementary angles, angles in triangles and parallelograms
-we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
-is eqaul to $\varphi$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
-\label{fig:pcorr_t_xy}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
-\label{fig:pcorr_t_lin}}
-\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
-\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-\end{proof}
-
-\end{document}

### 2018-02-02 2018/chapters/3_partial_correlation.tex
restructured

--- 2018-01-11 2018/paper.tex
@@
-\documentclass[nobib]{tufte-handout}
-
-\usepackage{ifluatex, ifxetex} % проверка, как компилируется файл
-
-% black magic!
-% solving problem with tufte-handout + xelatex
-% http://tex.stackexchange.com/questions/200722/
-% https://tex.stackexchange.com/questions/47576/
-\ifx\ifxetex\ifluatex\else % если xelatex или lualatex
-  \newcommand{\textls}[2][5]{%
-    \begingroup\addfontfeatures{LetterSpace=#1}#2\endgroup
-  }
-  \renewcommand{\allcapsspacing}[1]{\textls[15]{#1}}
-  \renewcommand{\smallcapsspacing}[1]{\textls[10]{#1}}
-  \renewcommand{\allcaps}[1]{\textls[15]{\MakeTextUppercase{#1}}}
-  \renewcommand{\smallcaps}[1]{\smallcapsspacing{\scshape\MakeTextLowercase{#1}}}
-  \renewcommand{\textsc}[1]{\smallcapsspacing{\textsmallcaps{#1}}}
-\fi
-
-% с альтернативного источника:
-%\ifluatex % Allow rendering with LuaTeX by setting up the spacing using fontspec features
-%  \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
-%  \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
-%\fi
-
-% lua produces strange notes
-% https://tex.stackexchange.com/questions/328431
-
-\usepackage{fontspec} % работа со шрифтами
-%\usepackage{polyglossia} % учим русский как иностранный :)
-
-%\setmainlanguage{engish}
-%\setotherlanguages{english}
-
-% заменяем --- на тире, << на кавычки и т.д.:
-\defaultfontfeatures{Ligatures=TeX}
-
-% download "Linux Libertine" fonts:
-% http://www.linuxlibertine.org/index.php?id=91&L=1
-\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
-% why do we need \newfontfamily:
-% http://tex.stackexchange.com/questions/91507/
-\newfontfamily{\cyrillicfonttt}{Linux Libertine O}
-\newfontfamily{\cyrillicfont}{Linux Libertine O}
-\newfontfamily{\cyrillicfontsf}{Linux Libertine O}
-
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{url} % вставка \url{}
-\usepackage{graphicx} % вставка графиков
-\usepackage{csquotes} % адаптирующиеся кавычки командой \enquote{}
-\usepackage{comment} % ingore everything between \begin{comment} \end{comment}
-\usepackage{answers} % separate problems and solutions
-\usepackage{tikz} % pictures with tikz language
-\usepackage{todonotes} % todo in documents
-\usepackage{multicol} % make several columns
-\usepackage{subfigure}
-
-\usepackage{enumitem} % для создания своих нумерующих списков (хак для гиперссылок)
-
-%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-
-% very useful during de-bugging!
-% \usepackage[left]{showlabels}
-% \showlabels{hypertarget}
-% \showlabels{hyperlink}
-
-\DeclareMathOperator{\Var}{Var}
-\DeclareMathOperator{\sVar}{sVar}
-\DeclareMathOperator{\pVar}{pVar}
-\DeclareMathOperator{\card}{card}
-\DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\Corr}{Corr}
-\DeclareMathOperator{\sCov}{sCov}
-\DeclareMathOperator{\pCov}{pCov}
-\DeclareMathOperator{\sCorr}{sCorr}
-\DeclareMathOperator{\pCorr}{pCorr}
-\DeclareMathOperator{\E}{E}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\I}{\mathbb{I}} % индикатор события
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-
-\usepackage[bibencoding = auto, backend = biber,
-sorting = none]{biblatex}
-
-%\addbibresource{probability_dna.bib}
-
-%\title{Geometric interpretaion of main concepts and theorems in econometrics}
-\title{some title}
-\author{}
-\date{\today}
-\begin{document}
-
-\maketitle
-
-%\section{Literature review}
-
-\section{Regression}
-
-\subsection{Sample correlation}
-
-\marginnote{
-\begin{align*}
-\sCorr(x,y) &= \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} \\
-&= \frac{\frac{1}{n-1}\sum_{i=1}^n (x_i - \bar x)(y_i - \bar y)}{\sqrt{\frac{1}{n-1} \sum_{i=1}^n (y_i - \bar y)\frac{1}{n-1} \sum_{i=1}^n (\hat y_i - \bar{\hat y})}}
-\end{align*}
-}
-
-The most crucial part in defining correlation geometrically is definig the dot product as it enables to compute the length of a vecotr:
-\[
-|\vec{a}| = \sqrt{\langle \vec{a},  \vec{a}\rangle}
-\]
-and the angle between any two vectors:
-\[
-\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}
-\]
-Now we define scalar product of two vectors
-$x =
-\begin{pmatrix}
-x_1 \\
-\vdots \\
-x_n
-\end{pmatrix}$
-and
-$y =
-\begin{pmatrix}
-y_1 \\
-\vdots \\
-y_n
-\end{pmatrix}$
-as a sample covariation between them:
-\[
-\langle x, y \rangle = \sCov(x, y)
-\]
-The main characteristics of a vector are its length and direction.
-So, we introduce the length
-\[
-\sqrt{\sCov(x,x)} = \sqrt{\sVar(x)} = \sigma_x
-\]
-and the angle between two random vectors
-\[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
-\]
-Note that from the definition of the angle it follows that the sample correlation can range from $-1$ to $1$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_def.pdf}
-\caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
-\label{fig:corr_def}
-\end{center}
-\end{figure}
-
-Another important geometrical tool is projection.
-Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$ can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
-The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for the vectors sampled from some distribution.
-The projection of such a vector $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
-\caption{Vector $y$ projected onto vector $x$.}
-\label{fig:corr_proj}
-\end{center}
-\end{figure}
-
-Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of sample correlation coefficient.
-Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can conclude that
-\[
-\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)}
-\]
-as the variance of a vector is associated with the square of its length.
-Thus, the sample correlation coefficient squared shows the fraction of variance in $y$ which can be explained
-with the most similar vector proportional to $x$.
-
-\subsection{Sample correlation when a constatnt vector added}
-
-\marginnote{
-\begin{align*}
-\sCorr(x + \alpha \mathbf{1}, y) &= \frac{\sCov(x + \alpha \mathbf{1}, y)}{\sVar(x + \alpha \mathbf{1}) \sVar(y)} \\
-&= \frac{\sCov(x,y)\sCov(\alpha \mathbf{1},y)}{\sVar(x)\sVar(y)} \\
-&= \frac{\sCov(x,y)}{\sVar(x)\sVar(y)} \\
-&= \sCorr(x,y)
-\end{align*}
-}
-
-\begin{theorem}
-Adding a vector of constants does not affect the sample correlation coefficient:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-where $\alpha \in \mathbb{R}$.
-\end{theorem}
-
-\begin{proof}
-Firstly, we project vectors $x$ and $y$ onto $Lin^{\perp}(\mathbf{1})$ in order to get $x^c = x - \bar x$ and $y^c = y - \bar y$
-(`c' stands for `centred'). It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
-\[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \vdots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
-\]
-Thus, projecting onto an orthogonal subspace is equivalent to substracting the projected vector, i.e., the vector of averages, from the original one.
-
-Also note that the angle $\varphi$ between original and centred vectors remains the same.
-The results of this step is shown in Figure~\ref{fig:corr_xyc}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
-  \caption{Centred vectors $x^c$ and $y^c$}
-  \label{fig:corr_xyc}
-\end{figure}
-
-Then we need to derive a new vector $\tilde x$ with constants added to each component.
-Geometrically adding a vector of costants means adding a vector of all ones
-scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
-Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
-$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
-After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
-By the properties of projection it is of no difference whether to project
-the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
-and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
-onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
-in Figure~\ref{fig:corr_tildex}.
-Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_br_up.pdf}
-  \caption{Decomposition and projection of $\tilde x$}
-  \label{fig:corr_tildex}
-\end{figure}
-
-Finally, putting everything together we finish the proof:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_final.pdf}
-  \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
-  \label{fig:corr_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Sample correlation coefficient in simple linear regression}
-
-\marginnote{
-Assuming the underlying relationship between $x$ and $y$ to be
-\[
-y_i = \beta_1 + \beta_2 x_i + \varepsilon_i \quad i=1,\ldots,n
-\]
-where $\varepsilon_i$ is an error term the following holds
-\begin{align*}
-\sCorr(y, \hat y) &= \frac{\sCov(y)\sCov(\hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_1 + \hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_1 + \hat \beta_2 x)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_2 x)}} \\
-&= \frac{\hat \beta_2 \sCov(y)\sCov(x)}{|\hat \beta_2| \sqrt{\sVar(y)\sVar(x)}} \\
-&= sign(\hat \beta_2) \frac{\sCov(y)\sCov(x)}{\sqrt{\sVar(y)\sVar(x)}}
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-has the property
-\[
-\sCorr(y, \hat y) = sign(\hat \beta_2) \sCorr(y, x)
-\]
-\end{theorem}
-
-\begin{proof}
-Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
-picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient represents the angle
-betweem two random vectors.
-So in order to complete the proof we need to find the appropriate angles and compare them.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
-\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
-\label{fig:corr_pos_base}
-\end{figure}
-
-However, it seems to be difficult to compare the angles in the three dimensional space.
-That is why we start with projecting both $x$ and $y$ onto the space perpendicular to the vector of all ones $\mathbf{1}$ as shown in Figure~\ref{fig:corr_pos_xyc}.
-We denote this space as $Lin^{\perp}(\mathbf{1})$. The resulting vectors are $x - \bar x \cdot \mathbf{1}$  and $y - \bar y \cdot \mathbf{1}$ respectively
-since projection of any vector $\vec{a}$ on the line given by a vector of all ones yields the vector of averages $\vec{\bar a}$.
-
-In order to get the angle between $y$ and $\hat y$ we should start with regressing $y$ on $Lin(x, \mathbf{1})$.
-Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
-Note that the projected $\hat y$ falls onto tha span of vector $x - \bar x \cdot \mathbf{1}$ as it can be decomposed into a sum $a x + b \mathbf{1}$ where $a, b \in \mathbb{R}$.
-$a x$ is projected in the same way as $x$ and $b \mathbf{1}$ yields zero when projected onto the orthogonal space.
-The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
-\label{fig:corr_pos_xyc}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
-\label{fig:corr_pos_yhatc}}
-\caption{\subref{fig:corr_pos_xyc}: `Centred' $x$ and $y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$; \subref{fig:corr_pos_yhatc}: `Centred' $\hat y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$.}
-\end{center}
-\end{figure}
-
-Since the projection of $\hat y$ lies exactly on the span of vector $x - \bar x \cdot \mathbf{1}$, we can conclude that $cos \varphi = \cos \varphi '$ and to put it another way $\sCorr(x,y) = \sCorr(y, \hat y)$.
-
-Now consider the case when $\hat \beta_2 < 0$.
-Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
-The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ and not on this vector itself while the projections of $x$ and $y$ remain the same.
-Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
-Using trigonometric properties, we simplify $\cos(180^{\circ} - \varphi) = -\cos\varphi$ which in turn implies $\sCorr(x,y) = -\sCorr(y,\hat y)$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics{images/corr_eq_negative.pdf}
-\label{fig:corr_negative}
-\caption{Case of $\beta_2 < 0$.}
-\end{center}
-\end{figure}
-\end{proof}
-
-\subsection{RSS + ESS = TSS}
-
-\marginnote{
-Consider a regresion model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-The OLS estimator for the vector of coefficients $\beta$ is
-\[
-\hat \beta = (X^T X)^{-1} X^T y
-\]
-and the residual vector is
-\begin{align*}
-\hat e &= y - \hat y \\
-&= y - X \hat \beta \\
-&= y - X (X^T X)^{-1} X^T y
-\end{align*}
-Then we define residual sum of squares (RSS), explained sum of squares (ESS) and total sum of squares (TSS) as follows:
-\begin{align*}
-RSS &= \lVert y - \hat y \rVert^2_2 \\
-ESS &= \lVert \hat y - \bar y \rVert^2_2 \\
-TSS &= \lVert y - \bar y \rVert \\
-\end{align*}
-Disclosing parentheses and using the fact that $\hat{y}^T y = \hat{y}^T \hat{y}$
-\begin{align*}
-\hat{y}^T y &= \beta^T X^T y \\
-&=  y^T X (X^T X)^{-1} X^T y \\
-\hat{y}^T \hat{y} &= \beta^T X^T X \beta \\
-&= y^T X (X^T X)^{-1} X^T X (X^T X)^{-1} X^T y \\
-&= y^T X (X^T X)^{-1} X^T y
-\end{align*}
-we obtain
-\begin{align*}
-RSS &= y^T y -\hat{y}^T \hat{y} \\
-ESS &= \hat{y}^T \hat{y} - \hat{y}^T \bar{y} + \bar{y}^T \bar{y} \\
-TSS &= y^T y - 2 y^T \bar y +  \bar{y}^T \bar{y}
-\end{align*}
-When putting everything together all the terms cancel out which proves
-\[
-ESS + RSS = TSS
-\]
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-RSS + ESS = TSS
-\]
-where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert^2_2$.
-\end{theorem}
-
-\begin{proof}
-The proof will be presented for the case of two regressor $x$ and $\mathbf{1}$ in order for the picture to be clear.
-However, the same logic applies for the case of $k$ regressors.
-
-We start with depicting the vectors $y \in \mathbb{R}^{n-2}$ and $x, \mathbf{1} \in \mathbb{R}^2$.
-Then we project $y$ onto $Lin(x, \mathbf{1})$ and obtain $\hat y$ which is shown in Figure~\ref{fig:tss_base}.
-
-From this picture we can immediately derive $\sqrt{RSS}$ as by definition this is the squared difference between $y$ and $\hat y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
-\label{fig:tss_base} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
-\label{fig:tss_rss}}
-\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
-\end{center}
-\end{figure}
-
-So as to visualize $ESS$ and $TSS$ we first need to visualize vector of averages $\bar y$.
-Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
-
-Now we both project $y$ and $\hat y$ onto $\mathbf{1}$ and following the definition obtain $\sqrt{TSS}$ as the difference vector $y - \bar y$ and $\sqrt{ESS}$ as the vector $\hat y - \bar y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
-\label{fig:tss_tss}}
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
-\label{fig:tss_ess}}
-\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
-\end{center}
-\end{figure}
-
-The final step is to put everything together.
-Note that since $y - \hat y$ is perpendicular to $Lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $Lin(x, \mathbf{1})$.
-Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
-Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the
-hypotenuse $\sqrt{TSS}$:
-\[
-(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2
-\]
-
-\begin{figure}[h!]
-\center{\includegraphics{images/tss_sum_final.pdf}}
-\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
-\label{fig:tss_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Determination coefficient}
-
-\marginnote{
-\begin{align*}
-\sCorr^2(y,\hat y) &= \left(\frac{\sCov(y, \hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}}\right)^2 \\
-&= \frac{\sCov(y, \hat y) \sCov(y, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sCov(\hat y + e, \hat y) \sCov(\hat y + e, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{(\sCov(\hat y, \hat y) + \sCov(e, \hat y))(\sCov(\hat y, \hat y) + \sCov(e, \hat y))}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y) \sVar(\hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y)}{\sVar(y)} \\
-&= \frac{ESS}{TSS} \\
-&= R^2
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{theorem}
-
-\begin{proof}
-Proving this theorem geometrically means showing that the determination coefficient can be interpreted as some squared angle
-which happens to be eqaul to the squared angle betwen $y$ and $\hat y$.
-
-Consider Figure~\ref{fig:tss_final} from the previous proof.
-It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
-Having defined the determination coefficient as
-\[
-R^2 = \frac{ESS}{TSS}
-\]
-we conclude that its geometric interpretaion is
-\[
-\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
-\]
-as shown in Figure~\ref{fig:r_sq_angle}.
-
-\begin{figure}
-  \center{\includegraphics{images/r_sq_ss.pdf}}
-  \caption{Determination coefficient as squared $\cos \varphi$}
-  \label{fig:r_sq_angle}
-\end{figure}
-
-Recall that the sample correlation coefficient two vectors was defined earlier as the angle between these two vectors.
-Thus, we conclude that $\sCorr(y, \hat y)$ is the angle between $y$ and $\hat y$ which is also eqaul to $\cos \varphi$.
-Finally, squaring both sides, we obtain
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{proof}
-
-\subsection{Regression line and point of averages}
-
-\marginnote{
-If the regression contains the intercept, the following equation holds:
-\begin{align*}
-\hat y  &= X \hat \beta \\
-&= X (X^T X)^{-1} X^T y \\
-&= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon
-\end{align*}
-Premultiplying both sides by $X^T$, we obtain:
-\begin{align*}
-X^T \hat y &=  X^T X (X^T X)^{-1} X^T X \beta \\
-&+ X^T X (X^T X)^{-1} X^T \varepsilon \\
-&= X^T X \beta + X^T \varepsilon
-\end{align*}
-This is a system of equations. The first row of $X^T$ is $\mathbf{1}$ vector, so we can write out the first equation:
-\[
-\sum_{i=1}^n \hat y_i = \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-From the first equation in the system
-\[
-X^T \hat y = X^T y
-\]
-we obtain
-\[
-\sum_{i=1}^{n} \hat y_i = \sum_{i=1}^{n} y
-\]
-And this finishes the proof:
-\[
-\frac{1}{n} \sum_{i=1}^{n} y = \frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-}
-
-\begin{theorem}
-In a linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-the point of averages lies on the estimated regression line.
-\end{theorem}
-
-\begin{proof}
-For the geometrical proof it suffices to show that $\hat y$ is a linear combination of the regressors, which is true by construction,
-and that $\frac{1}{n} \sum_{i=1}^{n} \hat y_i = \frac{1}{n} \sum_{i=1}^{n} y$. In order for the pictures to be more clear the proof will be presented for the case of two regressors.
-
-The first step is regressing $y$ on $Lin(\mathbf{1}, x)$. As shown in Figure~\ref{fig:averages_lin}, we obtain $\hat y$ as a linear combination of $\mathbf{1}$ and $x$.
-The next step is to regress both $y$ and $\hat y$ on $\mathbf{1}$ which results in $\bar y$ and $\bar \hat y$ correspondingly.
-By the theorem of three perpendiculars, $\bar y = \bar \hat y$ which is shown in Figure~\ref{fig:averages_bars}.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
-\label{fig:averages_lin} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
-\label{fig:averages_bars}}
-\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
-\end{center}
-\end{figure}
-
-\end{proof}
-
-\subsection{Frisch–Waugh–Lovell theorem}
-
-\marginnote{From regresison~\ref{eq:fwl_2} we get the following estimator:
-\begin{align*}
-\hat\beta_2 &= ((M_1 X_2)^T M_1 X_2)^{-1}(M_1 X_2)^T M_1 y \\
-&= (X_2^T M_1^T M_1 X_2)^{-1}  X_2^T M_1^T M_1 y \\
-&= (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\end{align*}
-As for regresison~\ref{eq:fwl_1}, let us note that due to $y = \hat y + \hat u$ $y$ can be decomposed as follows:
-\[
-y = Py + My = X_1 \hat \beta_1 + X_2 \hat \beta_2 + My
-\]
-Premultiplying both sides by $X_2^T M_1$, we obtain:
-\begin{align*}
-X_2^T M_1 y &= X_2^T M_1 X_1 \hat\beta_1 + X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&=  X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&= X_2^T M_1 X_2 \hat\beta_2
-\end{align*}
-On the last step we used the fact that
-\begin{align*}
-(X_2^T M_1 M y)^T = y^T M^T M_1^T X_2 \\
-= y^T M M_1 X_2 = y^T M X_2 = 0^T
-\end{align*}
-Assuming $X_2^T M_1 X_2$ is invertible, we get the same estimator
-\[
-\hat\beta_2 = (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\]
-}
-
-\begin{theorem}
-Consider regression
-\begin{equation} \label{eq:fwl_1}
-y = X_1 \beta_1 + X_2 \beta_2 + u
-\end{equation}
-where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
-$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
-Consider another regresison
-\begin{equation}  \label{eq:fwl_2}
-M_1 y = M_1 X_2 \beta_2 + M_1 u
-\end{equation}
-where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
-Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
-\end{theorem}
-
-\begin{proof}
-Geometrical proof will be presented for the following model:
-\begin{equation} \label{eq:fwl_proof}
-y_i = \beta_1 x_i + \beta_2 z_i + u_i
-\end{equation}
-
-We start with regression `all-at-once' and will distinct its coefficients with index $(1)$.
-The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
-which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_1_regression_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_1_regression_lin}}
-\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
-\end{center}
-\end{figure}
-
-As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.
-
-\begin{equation}\label{eq:fwl_2_y_clean}
-\begin{aligned}
-y &= \alpha z + \varepsilon \\
-\hat\alpha &= \frac{y^T z}{z^T z} \\
-\tilde{y} &= \hat\varepsilon = y - \frac{y^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.
-
-\begin{equation}\label{eq:fwl_2_x_clean}
-\begin{aligned}
-x &= \gamma z + \nu \\
-\hat\gamma &= \frac{x^T z}{z^T z} \\
-\tilde{x} &= \hat\nu = x - \frac{x^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Geometric results of these two steps are presented in~\ref{fig:fwl_2_regression_first}.
-
-Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
-So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_2_regression_first} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_2_regression_trans}}
-\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
-\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
-\end{center}
-\end{figure}
-
-Now, let us picture all the results in one figure and mark some main points.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_3_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_3_lin}}
-\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
-C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
-F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
-\end{center}
-\end{figure}
-
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
-and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
-Having two congruent angles, triangles ABC and FHC are simillar.
-Then, it follows:
-\[
-\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
-\]
-\end{proof}
-
-%\section{Geometric properties of distributions}
-
 \section{Partial correlation}
@@
   \centering
-  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
+  \includegraphics[width=0.55\linewidth]{figures/partial_corr.pdf}
   \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
@@
 \subfigure[]{
-\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
+\includegraphics[width=0.55\linewidth]{figures/part_corr_2.pdf}
 \label{fig:part_corr_alt}}
@@
 \subfigure[]{
-\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
+\includegraphics[width=0.35\linewidth]{figures/part_corr_2_lin.pdf}
 \label{fig:part_corr_alt_lin}}
@@
 \begin{theorem}
-Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
+Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the negative
 correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
@@
 \[
-Y = \beta_1 Z + \beta_2 Z + v
+Y = \beta_1 X + \beta_2 Z + v
 \]
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
+\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_x.pdf}
 \label{fig:pcorr_t_x}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
+\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_y.pdf}
 \label{fig:pcorr_t_y}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
+\includegraphics[width=0.45\linewidth]{figures/pcorr_t_xy.pdf}
 \label{fig:pcorr_t_xy}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
+\includegraphics[width=0.45\linewidth]{figures/pcorr_t_lin.pdf}
 \label{fig:pcorr_t_lin}}
@@
 \end{figure}
-
 \end{proof}
-
-\end{document}

### 2018-02-02 2018/chapters/4_distributions.tex
restructured

--- 2018-01-14 2018/paper.tex
@@
-\documentclass[nobib]{tufte-handout}
-
-\usepackage{ifluatex, ifxetex} % проверка, как компилируется файл
-
-% black magic!
-% solving problem with tufte-handout + xelatex
-% http://tex.stackexchange.com/questions/200722/
-% https://tex.stackexchange.com/questions/47576/
-\ifx\ifxetex\ifluatex\else % если xelatex или lualatex
-  \newcommand{\textls}[2][5]{%
-    \begingroup\addfontfeatures{LetterSpace=#1}#2\endgroup
-  }
-  \renewcommand{\allcapsspacing}[1]{\textls[15]{#1}}
-  \renewcommand{\smallcapsspacing}[1]{\textls[10]{#1}}
-  \renewcommand{\allcaps}[1]{\textls[15]{\MakeTextUppercase{#1}}}
-  \renewcommand{\smallcaps}[1]{\smallcapsspacing{\scshape\MakeTextLowercase{#1}}}
-  \renewcommand{\textsc}[1]{\smallcapsspacing{\textsmallcaps{#1}}}
-\fi
-
-% с альтернативного источника:
-%\ifluatex % Allow rendering with LuaTeX by setting up the spacing using fontspec features
-%  \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
-%  \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
-%\fi
-
-% lua produces strange notes
-% https://tex.stackexchange.com/questions/328431
-
-\usepackage{fontspec} % работа со шрифтами
-%\usepackage{polyglossia} % учим русский как иностранный :)
-
-%\setmainlanguage{engish}
-%\setotherlanguages{english}
-
-% заменяем --- на тире, << на кавычки и т.д.:
-\defaultfontfeatures{Ligatures=TeX}
-
-% download "Linux Libertine" fonts:
-% http://www.linuxlibertine.org/index.php?id=91&L=1
-\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
-% why do we need \newfontfamily:
-% http://tex.stackexchange.com/questions/91507/
-\newfontfamily{\cyrillicfonttt}{Linux Libertine O}
-\newfontfamily{\cyrillicfont}{Linux Libertine O}
-\newfontfamily{\cyrillicfontsf}{Linux Libertine O}
-
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{url} % вставка \url{}
-\usepackage{graphicx} % вставка графиков
-\usepackage{csquotes} % адаптирующиеся кавычки командой \enquote{}
-\usepackage{comment} % ingore everything between \begin{comment} \end{comment}
-\usepackage{answers} % separate problems and solutions
-\usepackage{tikz} % pictures with tikz language
-\usepackage{todonotes} % todo in documents
-\usepackage{multicol} % make several columns
-\usepackage{subfigure}
-
-\usepackage{enumitem} % для создания своих нумерующих списков (хак для гиперссылок)
-
-%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-
-% very useful during de-bugging!
-% \usepackage[left]{showlabels}
-% \showlabels{hypertarget}
-% \showlabels{hyperlink}
-
-\DeclareMathOperator{\Var}{Var}
-\DeclareMathOperator{\sVar}{sVar}
-\DeclareMathOperator{\pVar}{pVar}
-\DeclareMathOperator{\card}{card}
-\DeclareMathOperator{\Cov}{Cov}
-\DeclareMathOperator{\Corr}{Corr}
-\DeclareMathOperator{\sCov}{sCov}
-\DeclareMathOperator{\pCov}{pCov}
-\DeclareMathOperator{\sCorr}{sCorr}
-\DeclareMathOperator{\pCorr}{pCorr}
-\DeclareMathOperator{\E}{E}
-\DeclareMathOperator{\ctg}{ctg}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\I}{\mathbb{I}} % индикатор события
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-
-\usepackage[bibencoding = auto, backend = biber,
-sorting = none]{biblatex}
-
-%\addbibresource{probability_dna.bib}
-
-%\title{Geometric interpretaion of main concepts and theorems in econometrics}
-\title{some title}
-\author{}
-\date{\today}
-\begin{document}
-
-\maketitle
-
-%\section{Literature review}
-
-\section{Regression}
-
-\subsection{Sample correlation}
-
-\marginnote{
-\begin{align*}
-\sCorr(x,y) &= \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} \\
-&= \frac{\frac{1}{n-1}\sum_{i=1}^n (x_i - \bar x)(y_i - \bar y)}{\sqrt{\frac{1}{n-1} \sum_{i=1}^n (y_i - \bar y)\frac{1}{n-1} \sum_{i=1}^n (\hat y_i - \bar{\hat y})}}
-\end{align*}
-}
-
-The most crucial part in defining correlation geometrically is definig the dot product as it enables to compute the length of a vecotr:
-\[
-|\vec{a}| = \sqrt{\langle \vec{a},  \vec{a}\rangle}
-\]
-and the angle between any two vectors:
-\[
-\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}
-\]
-Now we define scalar product of two vectors
-$x =
-\begin{pmatrix}
-x_1 \\
-\vdots \\
-x_n
-\end{pmatrix}$
-and
-$y =
-\begin{pmatrix}
-y_1 \\
-\vdots \\
-y_n
-\end{pmatrix}$
-as a sample covariation between them:
-\[
-\langle x, y \rangle = \sCov(x, y)
-\]
-The main characteristics of a vector are its length and direction.
-So, we introduce the length
-\[
-\sqrt{\sCov(x,x)} = \sqrt{\sVar(x)} = \sigma_x
-\]
-and the angle between two random vectors
-\[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
-\]
-Note that from the definition of the angle it follows that the sample correlation can range from $-1$ to $1$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_def.pdf}
-\caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
-\label{fig:corr_def}
-\end{center}
-\end{figure}
-
-Another important geometrical tool is projection.
-Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$ can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
-The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for the vectors sampled from some distribution.
-The projection of such a vector $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
-\caption{Vector $y$ projected onto vector $x$.}
-\label{fig:corr_proj}
-\end{center}
-\end{figure}
-
-Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of sample correlation coefficient.
-Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can conclude that
-\[
-\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)}
-\]
-as the variance of a vector is associated with the square of its length.
-Thus, the sample correlation coefficient squared shows the fraction of variance in $y$ which can be explained
-with the most similar vector proportional to $x$.
-
-\subsection{Sample correlation when a constatnt vector added}
-
-\marginnote{
-\begin{align*}
-\sCorr(x + \alpha \mathbf{1}, y) &= \frac{\sCov(x + \alpha \mathbf{1}, y)}{\sVar(x + \alpha \mathbf{1}) \sVar(y)} \\
-&= \frac{\sCov(x,y)\sCov(\alpha \mathbf{1},y)}{\sVar(x)\sVar(y)} \\
-&= \frac{\sCov(x,y)}{\sVar(x)\sVar(y)} \\
-&= \sCorr(x,y)
-\end{align*}
-}
-
-\begin{theorem}
-Adding a vector of constants does not affect the sample correlation coefficient:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-where $\alpha \in \mathbb{R}$.
-\end{theorem}
-
-\begin{proof}
-Firstly, we project vectors $x$ and $y$ onto $Lin^{\perp}(\mathbf{1})$ in order to get $x^c = x - \bar x$ and $y^c = y - \bar y$
-(`c' stands for `centred'). It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
-\[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \vdots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
-\]
-Thus, projecting onto an orthogonal subspace is equivalent to substracting the projected vector, i.e., the vector of averages, from the original one.
-
-Also note that the angle $\varphi$ between original and centred vectors remains the same.
-The results of this step is shown in Figure~\ref{fig:corr_xyc}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
-  \caption{Centred vectors $x^c$ and $y^c$}
-  \label{fig:corr_xyc}
-\end{figure}
-
-Then we need to derive a new vector $\tilde x$ with constants added to each component.
-Geometrically adding a vector of costants means adding a vector of all ones
-scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
-Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
-$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
-After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
-By the properties of projection it is of no difference whether to project
-the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
-and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
-onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
-in Figure~\ref{fig:corr_tildex}.
-Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_br_up.pdf}
-  \caption{Decomposition and projection of $\tilde x$}
-  \label{fig:corr_tildex}
-\end{figure}
-
-Finally, putting everything together we finish the proof:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_final.pdf}
-  \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
-  \label{fig:corr_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Sample correlation coefficient in simple linear regression}
-
-\marginnote{
-Assuming the underlying relationship between $x$ and $y$ to be
-\[
-y_i = \beta_1 + \beta_2 x_i + \varepsilon_i \quad i=1,\ldots,n
-\]
-where $\varepsilon_i$ is an error term the following holds
-\begin{align*}
-\sCorr(y, \hat y) &= \frac{\sCov(y)\sCov(\hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_1 + \hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_1 + \hat \beta_2 x)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_2 x)}} \\
-&= \frac{\hat \beta_2 \sCov(y)\sCov(x)}{|\hat \beta_2| \sqrt{\sVar(y)\sVar(x)}} \\
-&= sign(\hat \beta_2) \frac{\sCov(y)\sCov(x)}{\sqrt{\sVar(y)\sVar(x)}}
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-has the property
-\[
-\sCorr(y, \hat y) = sign(\hat \beta_2) \sCorr(y, x)
-\]
-\end{theorem}
-
-\begin{proof}
-Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
-picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient represents the angle
-betweem two random vectors.
-So in order to complete the proof we need to find the appropriate angles and compare them.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
-\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
-\label{fig:corr_pos_base}
-\end{figure}
-
-However, it seems to be difficult to compare the angles in the three dimensional space.
-That is why we start with projecting both $x$ and $y$ onto the space perpendicular to the vector of all ones $\mathbf{1}$ as shown in Figure~\ref{fig:corr_pos_xyc}.
-We denote this space as $Lin^{\perp}(\mathbf{1})$. The resulting vectors are $x - \bar x \cdot \mathbf{1}$  and $y - \bar y \cdot \mathbf{1}$ respectively
-since projection of any vector $\vec{a}$ on the line given by a vector of all ones yields the vector of averages $\vec{\bar a}$.
-
-In order to get the angle between $y$ and $\hat y$ we should start with regressing $y$ on $Lin(x, \mathbf{1})$.
-Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
-Note that the projected $\hat y$ falls onto tha span of vector $x - \bar x \cdot \mathbf{1}$ as it can be decomposed into a sum $a x + b \mathbf{1}$ where $a, b \in \mathbb{R}$.
-$a x$ is projected in the same way as $x$ and $b \mathbf{1}$ yields zero when projected onto the orthogonal space.
-The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
-\label{fig:corr_pos_xyc}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
-\label{fig:corr_pos_yhatc}}
-\caption{\subref{fig:corr_pos_xyc}: `Centred' $x$ and $y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$; \subref{fig:corr_pos_yhatc}: `Centred' $\hat y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$.}
-\end{center}
-\end{figure}
-
-Since the projection of $\hat y$ lies exactly on the span of vector $x - \bar x \cdot \mathbf{1}$, we can conclude that $cos \varphi = \cos \varphi '$ and to put it another way $\sCorr(x,y) = \sCorr(y, \hat y)$.
-
-Now consider the case when $\hat \beta_2 < 0$.
-Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
-The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ and not on this vector itself while the projections of $x$ and $y$ remain the same.
-Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
-Using trigonometric properties, we simplify $\cos(180^{\circ} - \varphi) = -\cos\varphi$ which in turn implies $\sCorr(x,y) = -\sCorr(y,\hat y)$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics{images/corr_eq_negative.pdf}
-\label{fig:corr_negative}
-\caption{Case of $\beta_2 < 0$.}
-\end{center}
-\end{figure}
-\end{proof}
-
-\subsection{RSS + ESS = TSS}
-
-\marginnote{
-Consider a regresion model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-The OLS estimator for the vector of coefficients $\beta$ is
-\[
-\hat \beta = (X^T X)^{-1} X^T y
-\]
-and the residual vector is
-\begin{align*}
-\hat e &= y - \hat y \\
-&= y - X \hat \beta \\
-&= y - X (X^T X)^{-1} X^T y
-\end{align*}
-Then we define residual sum of squares (RSS), explained sum of squares (ESS) and total sum of squares (TSS) as follows:
-\begin{align*}
-RSS &= \lVert y - \hat y \rVert^2_2 \\
-ESS &= \lVert \hat y - \bar y \rVert^2_2 \\
-TSS &= \lVert y - \bar y \rVert \\
-\end{align*}
-Disclosing parentheses and using the fact that $\hat{y}^T y = \hat{y}^T \hat{y}$
-\begin{align*}
-\hat{y}^T y &= \beta^T X^T y \\
-&=  y^T X (X^T X)^{-1} X^T y \\
-\hat{y}^T \hat{y} &= \beta^T X^T X \beta \\
-&= y^T X (X^T X)^{-1} X^T X (X^T X)^{-1} X^T y \\
-&= y^T X (X^T X)^{-1} X^T y
-\end{align*}
-we obtain
-\begin{align*}
-RSS &= y^T y -\hat{y}^T \hat{y} \\
-ESS &= \hat{y}^T \hat{y} - \hat{y}^T \bar{y} + \bar{y}^T \bar{y} \\
-TSS &= y^T y - 2 y^T \bar y +  \bar{y}^T \bar{y}
-\end{align*}
-When putting everything together all the terms cancel out which proves
-\[
-ESS + RSS = TSS
-\]
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-RSS + ESS = TSS
-\]
-where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert^2_2$.
-\end{theorem}
-
-\begin{proof}
-The proof will be presented for the case of two regressor $x$ and $\mathbf{1}$ in order for the picture to be clear.
-However, the same logic applies for the case of $k$ regressors.
-
-We start with depicting the vectors $y \in \mathbb{R}^{n-2}$ and $x, \mathbf{1} \in \mathbb{R}^2$.
-Then we project $y$ onto $Lin(x, \mathbf{1})$ and obtain $\hat y$ which is shown in Figure~\ref{fig:tss_base}.
-
-From this picture we can immediately derive $\sqrt{RSS}$ as by definition this is the squared difference between $y$ and $\hat y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
-\label{fig:tss_base} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
-\label{fig:tss_rss}}
-\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
-\end{center}
-\end{figure}
-
-So as to visualize $ESS$ and $TSS$ we first need to visualize vector of averages $\bar y$.
-Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
-
-Now we both project $y$ and $\hat y$ onto $\mathbf{1}$ and following the definition obtain $\sqrt{TSS}$ as the difference vector $y - \bar y$ and $\sqrt{ESS}$ as the vector $\hat y - \bar y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
-\label{fig:tss_tss}}
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
-\label{fig:tss_ess}}
-\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
-\end{center}
-\end{figure}
-
-The final step is to put everything together.
-Note that since $y - \hat y$ is perpendicular to $Lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $Lin(x, \mathbf{1})$.
-Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
-Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the
-hypotenuse $\sqrt{TSS}$:
-\[
-(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2
-\]
-
-\begin{figure}[h!]
-\center{\includegraphics{images/tss_sum_final.pdf}}
-\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
-\label{fig:tss_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Determination coefficient}
-
-\marginnote{
-\begin{align*}
-\sCorr^2(y,\hat y) &= \left(\frac{\sCov(y, \hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}}\right)^2 \\
-&= \frac{\sCov(y, \hat y) \sCov(y, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sCov(\hat y + e, \hat y) \sCov(\hat y + e, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{(\sCov(\hat y, \hat y) + \sCov(e, \hat y))(\sCov(\hat y, \hat y) + \sCov(e, \hat y))}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y) \sVar(\hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y)}{\sVar(y)} \\
-&= \frac{ESS}{TSS} \\
-&= R^2
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{theorem}
-
-\begin{proof}
-Proving this theorem geometrically means showing that the determination coefficient can be interpreted as some squared angle
-which happens to be eqaul to the squared angle betwen $y$ and $\hat y$.
-
-Consider Figure~\ref{fig:tss_final} from the previous proof.
-It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
-Having defined the determination coefficient as
-\[
-R^2 = \frac{ESS}{TSS}
-\]
-we conclude that its geometric interpretaion is
-\[
-\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
-\]
-as shown in Figure~\ref{fig:r_sq_angle}.
-
-\begin{figure}
-  \center{\includegraphics{images/r_sq_ss.pdf}}
-  \caption{Determination coefficient as squared $\cos \varphi$}
-  \label{fig:r_sq_angle}
-\end{figure}
-
-Recall that the sample correlation coefficient two vectors was defined earlier as the angle between these two vectors.
-Thus, we conclude that $\sCorr(y, \hat y)$ is the angle between $y$ and $\hat y$ which is also eqaul to $\cos \varphi$.
-Finally, squaring both sides, we obtain
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{proof}
-
-\subsection{Regression line and point of averages}
-
-\marginnote{
-If the regression contains the intercept, the following equation holds:
-\begin{align*}
-\hat y  &= X \hat \beta \\
-&= X (X^T X)^{-1} X^T y \\
-&= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon
-\end{align*}
-Premultiplying both sides by $X^T$, we obtain:
-\begin{align*}
-X^T \hat y &=  X^T X (X^T X)^{-1} X^T X \beta \\
-&+ X^T X (X^T X)^{-1} X^T \varepsilon \\
-&= X^T X \beta + X^T \varepsilon
-\end{align*}
-This is a system of equations. The first row of $X^T$ is $\mathbf{1}$ vector, so we can write out the first equation:
-\[
-\sum_{i=1}^n \hat y_i = \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-From the first equation in the system
-\[
-X^T \hat y = X^T y
-\]
-we obtain
-\[
-\sum_{i=1}^{n} \hat y_i = \sum_{i=1}^{n} y
-\]
-And this finishes the proof:
-\[
-\frac{1}{n} \sum_{i=1}^{n} y = \frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-}
-
-\begin{theorem}
-In a linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-the point of averages lies on the estimated regression line.
-\end{theorem}
-
-\begin{proof}
-For the geometrical proof it suffices to show that $\hat y$ is a linear combination of the regressors, which is true by construction,
-and that $\frac{1}{n} \sum_{i=1}^{n} \hat y_i = \frac{1}{n} \sum_{i=1}^{n} y$. In order for the pictures to be more clear the proof will be presented for the case of two regressors.
-
-The first step is regressing $y$ on $Lin(\mathbf{1}, x)$. As shown in Figure~\ref{fig:averages_lin}, we obtain $\hat y$ as a linear combination of $\mathbf{1}$ and $x$.
-The next step is to regress both $y$ and $\hat y$ on $\mathbf{1}$ which results in $\bar y$ and $\bar \hat y$ correspondingly.
-By the theorem of three perpendiculars, $\bar y = \bar \hat y$ which is shown in Figure~\ref{fig:averages_bars}.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
-\label{fig:averages_lin} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
-\label{fig:averages_bars}}
-\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
-\end{center}
-\end{figure}
-
-\end{proof}
-
-\subsection{Frisch–Waugh–Lovell theorem}
-
-\marginnote{From regresison~\ref{eq:fwl_2} we get the following estimator:
-\begin{align*}
-\hat\beta_2 &= ((M_1 X_2)^T M_1 X_2)^{-1}(M_1 X_2)^T M_1 y \\
-&= (X_2^T M_1^T M_1 X_2)^{-1}  X_2^T M_1^T M_1 y \\
-&= (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\end{align*}
-As for regresison~\ref{eq:fwl_1}, let us note that due to $y = \hat y + \hat u$ $y$ can be decomposed as follows:
-\[
-y = Py + My = X_1 \hat \beta_1 + X_2 \hat \beta_2 + My
-\]
-Premultiplying both sides by $X_2^T M_1$, we obtain:
-\begin{align*}
-X_2^T M_1 y &= X_2^T M_1 X_1 \hat\beta_1 + X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&=  X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&= X_2^T M_1 X_2 \hat\beta_2
-\end{align*}
-On the last step we used the fact that
-\begin{align*}
-(X_2^T M_1 M y)^T = y^T M^T M_1^T X_2 \\
-= y^T M M_1 X_2 = y^T M X_2 = 0^T
-\end{align*}
-Assuming $X_2^T M_1 X_2$ is invertible, we get the same estimator
-\[
-\hat\beta_2 = (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\]
-}
-
-\begin{theorem}
-Consider regression
-\begin{equation} \label{eq:fwl_1}
-y = X_1 \beta_1 + X_2 \beta_2 + u
-\end{equation}
-where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
-$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
-Consider another regresison
-\begin{equation}  \label{eq:fwl_2}
-M_1 y = M_1 X_2 \beta_2 + M_1 u
-\end{equation}
-where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
-Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
-\end{theorem}
-
-\begin{proof}
-Geometrical proof will be presented for the following model:
-\begin{equation} \label{eq:fwl_proof}
-y_i = \beta_1 x_i + \beta_2 z_i + u_i
-\end{equation}
-
-We start with regression `all-at-once' and will distinct its coefficients with index $(1)$.
-The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
-which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_1_regression_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_1_regression_lin}}
-\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
-\end{center}
-\end{figure}
-
-As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.
-
-\begin{equation}\label{eq:fwl_2_y_clean}
-\begin{aligned}
-y &= \alpha z + \varepsilon \\
-\hat\alpha &= \frac{y^T z}{z^T z} \\
-\tilde{y} &= \hat\varepsilon = y - \frac{y^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.
-
-\begin{equation}\label{eq:fwl_2_x_clean}
-\begin{aligned}
-x &= \gamma z + \nu \\
-\hat\gamma &= \frac{x^T z}{z^T z} \\
-\tilde{x} &= \hat\nu = x - \frac{x^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Geometric results of these two steps are presented in~\ref{fig:fwl_2_regression_first}.
-
-Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
-So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_2_regression_first} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_2_regression_trans}}
-\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
-\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
-\end{center}
-\end{figure}
-
-Now, let us picture all the results in one figure and mark some main points.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_3_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_3_lin}}
-\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
-C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
-F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
-\end{center}
-\end{figure}
-
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
-and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
-Having two congruent angles, triangles ABC and FHC are simillar.
-Then, it follows:
-\[
-\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
-\]
-\end{proof}
-
-%\section{Geometric properties of distributions}
-
-\section{Partial correlation}
-
-\subsection{Definition of partial correlation}
-
-\marginnote{
-Partial correlation is the measure of degree of dependence between two random
-variables while controlling for the effect of other random variables:
-\[
-\pCorr(X,Y; Z) = \frac{\pCov(X,Y;Z)}{\sqrt{\pVar(X;Z)\pVar(Y;Z)}}
-\]
-where $\pVar(X;Z) = \Var(X - \alpha Z)$, $\alpha$ is such a constant that $\Cov(X - \alpha Z, Z) = 0$,
-and $\pCov(X,Y; Z) = \Cov(X - \alpha Z, Y - \beta Z)$, $\alpha$, $\beta$ are such
-constants that $\Cov(X - \alpha Z, Z) = \Cov(Y - \beta Z, Z) = 0$
-}
-
-Partial correlation can be defined into two ways.
-We will provide both definitions and show their equivalence.
-
-\begin{definition}
-Partial correlation between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the correlation coefficient between the residuals in regression of $X$ onto
-$Z$ and the residuals in regression of $Y$ onto $Z$.
-\end{definition}
-
-Firstly, we project random variable $X$ onto $Z$, which yields $\E(X)$.
-The residuals in this regression are $X - \E(X)$ — a vector in $Lin^{\perp}(Z)$.
-We will call this variable `centered' and label it as $\widetilde X$.
-Repeating this step for $Y$ yields `centerd' variable $\tilde Y = Y - \E(Y) \in Lin^{\perp}(Z)$.
-The angle between $\widetilde X$ and $\widetilde Y$ ($\varphi$ in Figure~\ref{fig:pcorr_def1})
-is the correaletion coefficient between these `centred' random variables and
-the partial correlaiton between the original ones.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
-  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
-  \label{fig:pcorr_def1}
-\end{figure}
-
-\begin{definition}
-Partial correlaiton between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the geometric mean between the coefficient $\hat \beta_{XY}$ in regression
-\[
-\hat X = \hat \beta_{XY} Y + \hat \beta_{XZ} Z
-\]
-and the coefficinent $\hat \beta_{YX}$ in regression
-\[
-\hat Y = \hat \beta_{YX} X + \hat \beta_{YZ} Z
-\]
-which has the same sign as the coefficients $\hat \beta_{XY}$ and $\hat \beta_{YX}$.
-\end{definition}
-
-Following the definition, we need to start with regressing variable $X$ onto
-$Y$ and $Z$. Then, the vector we obtained $\hat X$ can be broken up into the sum of
-$\hat \beta_{XY} Y$ and $\hat \beta_{XZ} Z$.
-Projecting $\hat \beta_{XY} Y$ onto $Lin^{\perp}(Z)$ we results in a vector
-$\alpha \widetilde Y$ where $\widetilde Y = Y - \E(Y)$ is the projection of $Y$ onto $Lin^{\perp}(Z)$.
-
-By the properties of similar triangles
-\[
-\frac{\beta_{XY} Y}{Y} = \frac{\alpha \widetilde Y}{\widetilde Y} \Leftrightarrow \beta_{XY} = \alpha
-\]
-
-In the same way we perform a regression of $Y$ onto $X$ and $Z$ and repeat the
-same steps as for $X$. Finally, we get the whole picture:
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
-\label{fig:part_corr_alt}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
-\label{fig:part_corr_alt_lin}}
-\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
-\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-Having plotted $Lin^{\perp}(Z)$ now we can express $\cos \varphi$ in terms of $\beta_{XY}$
-and $\beta_{YX}$
-
-\begin{equation}\label{eq:part_cor_cos}
-\begin{split}
-\cos \varphi &= \frac{\vert \beta_{XY} \widetilde Y \vert}{\vert \widetilde X \vert} = \vert \beta_{XY} \vert \\
-\cos \varphi &= \frac{\vert \beta_{YX} \widetilde X \vert}{\vert \widetilde Y \vert} = \vert \beta_{YX} \vert \\
-\cos^2 \varphi &= \vert \beta_{XY} \beta_{YX} \vert \stackrel{sign(\beta_{XY}) = sign(\beta_{YX})}{=} \beta_{XY} \beta_{YX}
-\end{split}
-\end{equation}
-
-Recall that the angle $\varphi$ can be interpreted as the partial correlation
-between $X$ and $Y$ holding $Z$ fixed, so it follows form equaitions~\eqref{eq:part_cor_cos}
-
-\[
-\pCorr^2(X,Y; Z) = \cos^2 \varphi = \beta_{XY} \beta_{YX}
-\]
-
-\subsection{Partial correlation as correlaiton between residuals}
-
-\marginnote{
-???
-}
-
-\begin{theorem}
-Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
-correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
-\[
-X = \alpha_1 Y + \alpha_2 Z + u
-\]
-and the residuals $\hat v$ in the model
-\[
-Y = \beta_1 Z + \beta_2 Z + v
-\]
-\end{theorem}
-
-\begin{proof}
-The first step is to find the residuals in the mentioned regresisions.
-For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
-which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
-and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
-in Figure~\ref{fig:pcorr_t_x}.
-
-Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
-\label{fig:pcorr_t_x}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
-\label{fig:pcorr_t_y}}
-\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
-\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
-\end{center}
-\end{figure}
-
-After all we put these figure together and the goal is to find the angle between
-the red and blue lines which is the same as the angle between the dashed
-red and blue lines. However, before that we need to apply translating to them to
-get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
-complementary angles, angles in triangles and parallelograms
-we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
-is eqaul to $\varphi$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
-\label{fig:pcorr_t_xy}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
-\label{fig:pcorr_t_lin}}
-\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
-\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-\end{proof}
-
 \section{Probability distributions}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/ttest.pdf}
+\includegraphics[width=0.45\linewidth]{figures/ttest.pdf}
 \label{fig:ttest_3d}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/ttest_lin.pdf}
+\includegraphics[width=0.45\linewidth]{figures/ttest_lin.pdf}
 \label{fig:ttest_lin}}
@@
 \centering
-\includegraphics[width=0.55\linewidth]{images/ftest.pdf}
+\includegraphics[width=0.55\linewidth]{figures/ftest.pdf}
 \caption{F-statistic as the cotangens squared of $\varphi$}
@@
 \end{figure}
-
-\end{document}

### 2018-02-02 2018/figures/pic3.tex
restructured

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}
\tdplotsetmaincoords{100}{110}
\begin{tikzpicture}[scale=5,tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (un) at (0.75,0.75,0);
\coordinate (bx) at (-1.75,0.5,0);
\coordinate (unl) at (1.5,1.5,0);

\tdplotsetcoord{Y}{1.75}{30}{90}

\draw[->] (O) -- (bx) node[anchor=south west]{$x$};
\draw[->] (O) -- (Y) node[anchor=south west]{$y$};
\draw[->] (O) -- (un) node[anchor=south east]{$\bar{y}\cdot\mathbf{1}$};
\draw[dashed,->] (O) -- (unl) node[anchor=south west]{$\mathbf{1}$};

\draw[->] (O) -- (Yxy) node[anchor=south west]{$\hat{y}$};
\draw[dashed, color=blue] (Y) -- (Yxy) node[midway,color=black,anchor=west]{\tiny$\sqrt{RSS}$};
\draw[dashed, color=blue] (Y) -- (un) node[midway,color=black,anchor=east]{\tiny$\sqrt{TSS}$};
\draw[dashed, color=blue] (un) -- (Yxy) node[midway,color=black,anchor=north east]{\tiny$\sqrt{ESS}$};

\tkzMarkRightAngle[size=0.03](Y,Yxy,un);

\tdplotsetthetaplanecoords{110}
\tdplotdrawarc[tdplot_rotated_coords]{(un)}{0.03}{12}{110}{anchor=south west}{\footnotesize$\varphi$}

\end{tikzpicture}

\end{document}

### 2018-02-02 2018/paper.tex
restructured

--- 2018-01-11 2018/paper.tex
@@
 
-%\section{Literature review}
+\input{chapters/1_random_variables.tex}
 
-\section{Regression}
+\input{chapters/2_regression.tex}
 
-\subsection{Sample correlation}
+\input{chapters/3_partial_correlation.tex}
 
-\marginnote{
-\begin{align*}
-\sCorr(x,y) &= \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} \\
-&= \frac{\frac{1}{n-1}\sum_{i=1}^n (x_i - \bar x)(y_i - \bar y)}{\sqrt{\frac{1}{n-1} \sum_{i=1}^n (y_i - \bar y)\frac{1}{n-1} \sum_{i=1}^n (\hat y_i - \bar{\hat y})}}
-\end{align*}
-}
-
-The most crucial part in defining correlation geometrically is definig the dot product as it enables to compute the length of a vecotr:
-\[
-|\vec{a}| = \sqrt{\langle \vec{a},  \vec{a}\rangle}
-\]
-and the angle between any two vectors:
-\[
-\cos(\vec{a}, \vec{b}) = \frac{\langle \vec{a},  \vec{a}\rangle}{|\vec{a}| |\vec{b}|}
-\]
-Now we define scalar product of two vectors
-$x =
-\begin{pmatrix}
-x_1 \\
-\vdots \\
-x_n
-\end{pmatrix}$
-and
-$y =
-\begin{pmatrix}
-y_1 \\
-\vdots \\
-y_n
-\end{pmatrix}$
-as a sample covariation between them:
-\[
-\langle x, y \rangle = \sCov(x, y)
-\]
-The main characteristics of a vector are its length and direction.
-So, we introduce the length
-\[
-\sqrt{\sCov(x,x)} = \sqrt{\sVar(x)} = \sigma_x
-\]
-and the angle between two random vectors
-\[
-\cos(x,y) = \frac{\sCov(x,y)}{\sqrt{\sVar(x)\sVar(y)}} = \sCorr(x,y)
-\]
-Note that from the definition of the angle it follows that the sample correlation can range from $-1$ to $1$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_def.pdf}
-\caption{Vector $x$ of length $\sigma_x$ and random vector $y$ of length $\sigma_y$, $\cos \varphi$ is the angle between $x$ and $y$.}
-\label{fig:corr_def}
-\end{center}
-\end{figure}
-
-Another important geometrical tool is projection.
-Recall that for any two vectors the scalar product $\langle \vec{a}, \vec{b} \rangle$ can be interpreted as the length of projected $\vec{b}$ multuplied by the length of $\vec{a}$.
-The projection itself is $cos(\vec{a}, \vec{b}) \vec{b}$.
-Same holds for the vectors sampled from some distribution.
-The projection of such a vector $y$ onto $\{cx| c \in \mathbb{R}\}$ is $\hat y = \sCorr(x,y) \cdot y$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.4\linewidth]{images/corr_proj.pdf}
-\caption{Vector $y$ projected onto vector $x$.}
-\label{fig:corr_proj}
-\end{center}
-\end{figure}
-
-Looking at Figure~\ref{fig:corr_proj}, we can interpret the square of sample correlation coefficient.
-Using the fact that $cos^2 \varphi$ is the squared ratio of the leg adjacent to $\varphi$ to hypotenuse, we can conclude that
-\[
-\sCorr^2(x,y) = \frac{\sVar(\hat y)}{\sVar(y)}
-\]
-as the variance of a vector is associated with the square of its length.
-Thus, the sample correlation coefficient squared shows the fraction of variance in $y$ which can be explained
-with the most similar vector proportional to $x$.
-
-\subsection{Sample correlation when a constatnt vector added}
-
-\marginnote{
-\begin{align*}
-\sCorr(x + \alpha \mathbf{1}, y) &= \frac{\sCov(x + \alpha \mathbf{1}, y)}{\sVar(x + \alpha \mathbf{1}) \sVar(y)} \\
-&= \frac{\sCov(x,y)\sCov(\alpha \mathbf{1},y)}{\sVar(x)\sVar(y)} \\
-&= \frac{\sCov(x,y)}{\sVar(x)\sVar(y)} \\
-&= \sCorr(x,y)
-\end{align*}
-}
-
-\begin{theorem}
-Adding a vector of constants does not affect the sample correlation coefficient:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-where $\alpha \in \mathbb{R}$.
-\end{theorem}
-
-\begin{proof}
-Firstly, we project vectors $x$ and $y$ onto $Lin^{\perp}(\mathbf{1})$ in order to get $x^c = x - \bar x$ and $y^c = y - \bar y$
-(`c' stands for `centred'). It can be shown that the matrix corresponding to projecting onto the line spanned by a vector of all ones has the following form
-\[
-\frac{\mathbf{1}^T \mathbf{1}}{\mathbf{1} \mathbf{1}^T} = \frac{\begin{pmatrix} 1 & \ldots & 1 \end{pmatrix} \begin{pmatrix} 1 \\ \vdots \\ 1 \end{pmatrix}}{\sum_{i=1}^n 1} = \begin{pmatrix} \frac{1}{n} & \ldots & \frac{1}{n} \\ \vdots & \ddots & \vdots \\ \frac{1}{n} & \ldots & \frac{1}{n} \end{pmatrix}
-\]
-Thus, projecting onto an orthogonal subspace is equivalent to substracting the projected vector, i.e., the vector of averages, from the original one.
-
-Also note that the angle $\varphi$ between original and centred vectors remains the same.
-The results of this step is shown in Figure~\ref{fig:corr_xyc}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\linewidth]{images/corr_const_cent.pdf}
-  \caption{Centred vectors $x^c$ and $y^c$}
-  \label{fig:corr_xyc}
-\end{figure}
-
-Then we need to derive a new vector $\tilde x$ with constants added to each component.
-Geometrically adding a vector of costants means adding a vector of all ones
-scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
-Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
-$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
-After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
-By the properties of projection it is of no difference whether to project
-the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
-and $\beta x$ — the result is the same.
-So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
-onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
-in Figure~\ref{fig:corr_tildex}.
-Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_br_up.pdf}
-  \caption{Decomposition and projection of $\tilde x$}
-  \label{fig:corr_tildex}
-\end{figure}
-
-Finally, putting everything together we finish the proof:
-\[
-\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)
-\]
-
-\begin{figure}
-  \centering
-  \includegraphics{images/corr_const_final.pdf}
-  \caption{$\sCorr(x + \alpha \mathbf{1}, y) = \sCorr(x,y)$ as the corresponding angles are equal.}
-  \label{fig:corr_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Sample correlation coefficient in simple linear regression}
-
-\marginnote{
-Assuming the underlying relationship between $x$ and $y$ to be
-\[
-y_i = \beta_1 + \beta_2 x_i + \varepsilon_i \quad i=1,\ldots,n
-\]
-where $\varepsilon_i$ is an error term the following holds
-\begin{align*}
-\sCorr(y, \hat y) &= \frac{\sCov(y)\sCov(\hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_1 + \hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_1 + \hat \beta_2 x)}} \\
-&= \frac{\sCov(y)\sCov(\hat \beta_2 x)}{\sqrt{\sVar(y)\sVar(\hat \beta_2 x)}} \\
-&= \frac{\hat \beta_2 \sCov(y)\sCov(x)}{|\hat \beta_2| \sqrt{\sVar(y)\sVar(x)}} \\
-&= sign(\hat \beta_2) \frac{\sCov(y)\sCov(x)}{\sqrt{\sVar(y)\sVar(x)}}
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-has the property
-\[
-\sCorr(y, \hat y) = sign(\hat \beta_2) \sCorr(y, x)
-\]
-\end{theorem}
-
-\begin{proof}
-Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
-picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
-It has been shown earlier that the correlation coefficient represents the angle
-betweem two random vectors.
-So in order to complete the proof we need to find the appropriate angles and compare them.
-
-\begin{figure}[h!]
-\center{\includegraphics[width=0.35\linewidth]{images/corr_eq_positive.pdf}}
-\caption{Vectors $x$, $y$ and $\mathbf{1}$.}
-\label{fig:corr_pos_base}
-\end{figure}
-
-However, it seems to be difficult to compare the angles in the three dimensional space.
-That is why we start with projecting both $x$ and $y$ onto the space perpendicular to the vector of all ones $\mathbf{1}$ as shown in Figure~\ref{fig:corr_pos_xyc}.
-We denote this space as $Lin^{\perp}(\mathbf{1})$. The resulting vectors are $x - \bar x \cdot \mathbf{1}$  and $y - \bar y \cdot \mathbf{1}$ respectively
-since projection of any vector $\vec{a}$ on the line given by a vector of all ones yields the vector of averages $\vec{\bar a}$.
-
-In order to get the angle between $y$ and $\hat y$ we should start with regressing $y$ on $Lin(x, \mathbf{1})$.
-Then the only thing thing left is to project $\hat y$ onto $Lin^{\perp}\mathbf{1}$ since the $y$ vector has already been projected.
-Note that the projected $\hat y$ falls onto tha span of vector $x - \bar x \cdot \mathbf{1}$ as it can be decomposed into a sum $a x + b \mathbf{1}$ where $a, b \in \mathbb{R}$.
-$a x$ is projected in the same way as $x$ and $b \mathbf{1}$ yields zero when projected onto the orthogonal space.
-The result of this step is shown in Figure~\ref{fig:corr_pos_yhatc}.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_xy_centered.pdf}
-\label{fig:corr_pos_xyc}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.6\linewidth]{images/corr_eq_positive_yhat.pdf}
-\label{fig:corr_pos_yhatc}}
-\caption{\subref{fig:corr_pos_xyc}: `Centred' $x$ and $y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$; \subref{fig:corr_pos_yhatc}: `Centred' $\hat y$, i.e., projected onto $Lin^{\perp}(\mathbf{1})$.}
-\end{center}
-\end{figure}
-
-Since the projection of $\hat y$ lies exactly on the span of vector $x - \bar x \cdot \mathbf{1}$, we can conclude that $cos \varphi = \cos \varphi '$ and to put it another way $\sCorr(x,y) = \sCorr(y, \hat y)$.
-
-Now consider the case when $\hat \beta_2 < 0$.
-Note that the sign of $\beta_1$ does not influence the correlation coefficient sign.
-The only difference is that now $\hat y$ is projected onto the span of  $x - \bar x \cdot \mathbf{1}$ and not on this vector itself while the projections of $x$ and $y$ remain the same.
-Looking at Figure~\ref{fig:corr_negative} we deduce that the angle betwween $y$ and $\hat y$ is compelement to the angle between $x$ and $y$.
-Using trigonometric properties, we simplify $\cos(180^{\circ} - \varphi) = -\cos\varphi$ which in turn implies $\sCorr(x,y) = -\sCorr(y,\hat y)$.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics{images/corr_eq_negative.pdf}
-\label{fig:corr_negative}
-\caption{Case of $\beta_2 < 0$.}
-\end{center}
-\end{figure}
-\end{proof}
-
-\subsection{RSS + ESS = TSS}
-
-\marginnote{
-Consider a regresion model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-The OLS estimator for the vector of coefficients $\beta$ is
-\[
-\hat \beta = (X^T X)^{-1} X^T y
-\]
-and the residual vector is
-\begin{align*}
-\hat e &= y - \hat y \\
-&= y - X \hat \beta \\
-&= y - X (X^T X)^{-1} X^T y
-\end{align*}
-Then we define residual sum of squares (RSS), explained sum of squares (ESS) and total sum of squares (TSS) as follows:
-\begin{align*}
-RSS &= \lVert y - \hat y \rVert^2_2 \\
-ESS &= \lVert \hat y - \bar y \rVert^2_2 \\
-TSS &= \lVert y - \bar y \rVert \\
-\end{align*}
-Disclosing parentheses and using the fact that $\hat{y}^T y = \hat{y}^T \hat{y}$
-\begin{align*}
-\hat{y}^T y &= \beta^T X^T y \\
-&=  y^T X (X^T X)^{-1} X^T y \\
-\hat{y}^T \hat{y} &= \beta^T X^T X \beta \\
-&= y^T X (X^T X)^{-1} X^T X (X^T X)^{-1} X^T y \\
-&= y^T X (X^T X)^{-1} X^T y
-\end{align*}
-we obtain
-\begin{align*}
-RSS &= y^T y -\hat{y}^T \hat{y} \\
-ESS &= \hat{y}^T \hat{y} - \hat{y}^T \bar{y} + \bar{y}^T \bar{y} \\
-TSS &= y^T y - 2 y^T \bar y +  \bar{y}^T \bar{y}
-\end{align*}
-When putting everything together all the terms cancel out which proves
-\[
-ESS + RSS = TSS
-\]
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-RSS + ESS = TSS
-\]
-where $RSS = \lVert y - \hat y \rVert^2_2$, $ESS = \lVert \hat y - \bar y \rVert^2_2$, $TSS = \lVert y - \bar y \rVert^2_2$.
-\end{theorem}
-
-\begin{proof}
-The proof will be presented for the case of two regressor $x$ and $\mathbf{1}$ in order for the picture to be clear.
-However, the same logic applies for the case of $k$ regressors.
-
-We start with depicting the vectors $y \in \mathbb{R}^{n-2}$ and $x, \mathbf{1} \in \mathbb{R}^2$.
-Then we project $y$ onto $Lin(x, \mathbf{1})$ and obtain $\hat y$ which is shown in Figure~\ref{fig:tss_base}.
-
-From this picture we can immediately derive $\sqrt{RSS}$ as by definition this is the squared difference between $y$ and $\hat y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_base.pdf}
-\label{fig:tss_base} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_rss.pdf}
-\label{fig:tss_rss}}
-\caption{\subref{fig:tss_base}: Vectors $y \in \mathbb{R}^{n-2}$ and $\hat y \in Lin(x, \mathbf{1})$; \subref{fig:tss_rss}: Residual sum of squares.}
-\end{center}
-\end{figure}
-
-So as to visualize $ESS$ and $TSS$ we first need to visualize vector of averages $\bar y$.
-Geometrically this means projecting a vector onto a line spanned by vector $\mathbf{1}$.
-
-Now we both project $y$ and $\hat y$ onto $\mathbf{1}$ and following the definition obtain $\sqrt{TSS}$ as the difference vector $y - \bar y$ and $\sqrt{ESS}$ as the vector $\hat y - \bar y$.
-
-\begin{figure}[h!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_tss.pdf}
-\label{fig:tss_tss}}
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/tss_sum_ess.pdf}
-\label{fig:tss_ess}}
-\caption{\subref{fig:tss_tss}: Total sum of squares; \subref{fig:tss_ess}: Explained sum of squares.}
-\end{center}
-\end{figure}
-
-The final step is to put everything together.
-Note that since $y - \hat y$ is perpendicular to $Lin(x, \mathbf{1})$ it is also perpendicular to $\hat y - \bar y$ and $\mathbf{1}$ as these vectoros are in $Lin(x, \mathbf{1})$.
-Then, applying the theorem of three perpendiculars we conclude that the foot of vector $y - \bar y$ is the same point as the foot of the vector $\hat y - \bar y$.
-Thus, we obtain a right angle triangle and can apply the Pythagorean theorem for the catheti $\sqrt{RSS}$ and $\sqrt{ESS}$ and the
-hypotenuse $\sqrt{TSS}$:
-\[
-(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2
-\]
-
-\begin{figure}[h!]
-\center{\includegraphics{images/tss_sum_final.pdf}}
-\caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
-\label{fig:tss_final}
-\end{figure}
-
-\end{proof}
-
-\subsection{Determination coefficient}
-
-\marginnote{
-\begin{align*}
-\sCorr^2(y,\hat y) &= \left(\frac{\sCov(y, \hat y)}{\sqrt{\sVar(y)\sVar(\hat y)}}\right)^2 \\
-&= \frac{\sCov(y, \hat y) \sCov(y, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sCov(\hat y + e, \hat y) \sCov(\hat y + e, \hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{(\sCov(\hat y, \hat y) + \sCov(e, \hat y))(\sCov(\hat y, \hat y) + \sCov(e, \hat y))}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y) \sVar(\hat y)}{\sVar(y)\sVar(\hat y)} \\
-&= \frac{\sVar(\hat y)}{\sVar(y)} \\
-&= \frac{ESS}{TSS} \\
-&= R^2
-\end{align*}
-}
-
-\begin{theorem}
-A linear regression model with $n$ observations and $k$ explanatory variables including a constant unit vector
-\[
-y = X \beta + \varepsilon
-\]
-has the following property
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{theorem}
-
-\begin{proof}
-Proving this theorem geometrically means showing that the determination coefficient can be interpreted as some squared angle
-which happens to be eqaul to the squared angle betwen $y$ and $\hat y$.
-
-Consider Figure~\ref{fig:tss_final} from the previous proof.
-It was shown there that the vectors $y - \bar y$, $y - \hat y$ and $\hat y - \bar y$ form a right triangle.
-Having defined the determination coefficient as
-\[
-R^2 = \frac{ESS}{TSS}
-\]
-we conclude that its geometric interpretaion is
-\[
-\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
-\]
-as shown in Figure~\ref{fig:r_sq_angle}.
-
-\begin{figure}
-  \center{\includegraphics{images/r_sq_ss.pdf}}
-  \caption{Determination coefficient as squared $\cos \varphi$}
-  \label{fig:r_sq_angle}
-\end{figure}
-
-Recall that the sample correlation coefficient two vectors was defined earlier as the angle between these two vectors.
-Thus, we conclude that $\sCorr(y, \hat y)$ is the angle between $y$ and $\hat y$ which is also eqaul to $\cos \varphi$.
-Finally, squaring both sides, we obtain
-\[
-R^2 = \sCorr^2(y, \hat y)
-\]
-\end{proof}
-
-\subsection{Regression line and point of averages}
-
-\marginnote{
-If the regression contains the intercept, the following equation holds:
-\begin{align*}
-\hat y  &= X \hat \beta \\
-&= X (X^T X)^{-1} X^T y \\
-&= X (X^T X)^{-1} X^T X \beta + X (X^T X)^{-1} X^T \varepsilon
-\end{align*}
-Premultiplying both sides by $X^T$, we obtain:
-\begin{align*}
-X^T \hat y &=  X^T X (X^T X)^{-1} X^T X \beta \\
-&+ X^T X (X^T X)^{-1} X^T \varepsilon \\
-&= X^T X \beta + X^T \varepsilon
-\end{align*}
-This is a system of equations. The first row of $X^T$ is $\mathbf{1}$ vector, so we can write out the first equation:
-\[
-\sum_{i=1}^n \hat y_i = \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-From the first equation in the system
-\[
-X^T \hat y = X^T y
-\]
-we obtain
-\[
-\sum_{i=1}^{n} \hat y_i = \sum_{i=1}^{n} y
-\]
-And this finishes the proof:
-\[
-\frac{1}{n} \sum_{i=1}^{n} y = \frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{k} x_{ij} \beta_{j}
-\]
-}
-
-\begin{theorem}
-In a linear regression model with one explanatory variable and constant term
-\[
-y = \beta_1 + \beta_2 x + \varepsilon
-\]
-the point of averages lies on the estimated regression line.
-\end{theorem}
-
-\begin{proof}
-For the geometrical proof it suffices to show that $\hat y$ is a linear combination of the regressors, which is true by construction,
-and that $\frac{1}{n} \sum_{i=1}^{n} \hat y_i = \frac{1}{n} \sum_{i=1}^{n} y$. In order for the pictures to be more clear the proof will be presented for the case of two regressors.
-
-The first step is regressing $y$ on $Lin(\mathbf{1}, x)$. As shown in Figure~\ref{fig:averages_lin}, we obtain $\hat y$ as a linear combination of $\mathbf{1}$ and $x$.
-The next step is to regress both $y$ and $\hat y$ on $\mathbf{1}$ which results in $\bar y$ and $\bar \hat y$ correspondingly.
-By the theorem of three perpendiculars, $\bar y = \bar \hat y$ which is shown in Figure~\ref{fig:averages_bars}.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_1.pdf}
-\label{fig:averages_lin} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/averages_2.pdf}
-\label{fig:averages_bars}}
-\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
-\end{center}
-\end{figure}
-
-\end{proof}
-
-\subsection{Frisch–Waugh–Lovell theorem}
-
-\marginnote{From regresison~\ref{eq:fwl_2} we get the following estimator:
-\begin{align*}
-\hat\beta_2 &= ((M_1 X_2)^T M_1 X_2)^{-1}(M_1 X_2)^T M_1 y \\
-&= (X_2^T M_1^T M_1 X_2)^{-1}  X_2^T M_1^T M_1 y \\
-&= (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\end{align*}
-As for regresison~\ref{eq:fwl_1}, let us note that due to $y = \hat y + \hat u$ $y$ can be decomposed as follows:
-\[
-y = Py + My = X_1 \hat \beta_1 + X_2 \hat \beta_2 + My
-\]
-Premultiplying both sides by $X_2^T M_1$, we obtain:
-\begin{align*}
-X_2^T M_1 y &= X_2^T M_1 X_1 \hat\beta_1 + X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&=  X_2^T M_1 X_2 \hat\beta_2 + X_2^T M_1 M y \\
-&= X_2^T M_1 X_2 \hat\beta_2
-\end{align*}
-On the last step we used the fact that
-\begin{align*}
-(X_2^T M_1 M y)^T = y^T M^T M_1^T X_2 \\
-= y^T M M_1 X_2 = y^T M X_2 = 0^T
-\end{align*}
-Assuming $X_2^T M_1 X_2$ is invertible, we get the same estimator
-\[
-\hat\beta_2 = (X_2^T M_1 X_2)^{-1}  X_2^T M_1 y
-\]
-}
-
-\begin{theorem}
-Consider regression
-\begin{equation} \label{eq:fwl_1}
-y = X_1 \beta_1 + X_2 \beta_2 + u
-\end{equation}
-where $X_{n \times k} = [X_1 X_2]$, i.e. $X_1$ consists of first $k_1$ columns of $X$ and $X_2$ consists of remaining $k_2$ columns of $X$,
-$\beta_1$ and $\beta_2$ are comfortable, i.e. $k_1 \times 1$ and $k_2 \times 1$ vectors.
-Consider another regresison
-\begin{equation}  \label{eq:fwl_2}
-M_1 y = M_1 X_2 \beta_2 + M_1 u
-\end{equation}
-where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
-Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
-\end{theorem}
-
-\begin{proof}
-Geometrical proof will be presented for the following model:
-\begin{equation} \label{eq:fwl_proof}
-y_i = \beta_1 x_i + \beta_2 z_i + u_i
-\end{equation}
-
-We start with regression `all-at-once' and will distinct its coefficients with index $(1)$.
-The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
-which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_1_regression_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_1_regression_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_1_regression_lin}}
-\caption{\subref{fig:fwl_1_regression_3d}: Regression of $y$ on $Lin(x,z)$; \subref{fig:fwl_1_regression_lin}: $Lin(x, z)$.}
-\end{center}
-\end{figure}
-
-As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.
-
-\begin{equation}\label{eq:fwl_2_y_clean}
-\begin{aligned}
-y &= \alpha z + \varepsilon \\
-\hat\alpha &= \frac{y^T z}{z^T z} \\
-\tilde{y} &= \hat\varepsilon = y - \frac{y^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.
-
-\begin{equation}\label{eq:fwl_2_x_clean}
-\begin{aligned}
-x &= \gamma z + \nu \\
-\hat\gamma &= \frac{x^T z}{z^T z} \\
-\tilde{x} &= \hat\nu = x - \frac{x^T z}{z^T z}z
-\end{aligned}
-\end{equation}
-
-Geometric results of these two steps are presented in~\ref{fig:fwl_2_regression_first}.
-
-Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
-So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_2_regression_first} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_2_regression_translation.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_2_regression_trans}}
-\caption{\subref{fig:fwl_2_regression_first}: Regression of $y$ on $z$ and of $x$ on $z$;
-\subref{fig:fwl_2_regression_trans}: Translation of $\tilde{x}$.}
-\end{center}
-\end{figure}
-
-Now, let us picture all the results in one figure and mark some main points.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_3d.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
-\label{fig:fwl_3_3d} }
-\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.4\linewidth]{images/fwl_3_lin.pdf}
-%\caption{$Lin(x, z)$}
-\label{fig:fwl_3_lin}}
-\caption{\subref{fig:fwl_3_3d}: Point A stands for the origin, B — $\hat\gamma z$,
-C — $x$, D — $\hat\alpha z$, E — intersection of vector $x$ and line parallel to $\tilde x$,
-F — $\hat\beta_1^{(1)} x$, G — $\hat\beta_1^{(2)} \tilde{x}$; \subref{fig:fwl_3_lin}: $Lin(x,z)$.}
-\end{center}
-\end{figure}
-
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
-and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
-Having two congruent angles, triangles ABC and FHC are simillar.
-Then, it follows:
-\[
-\frac{AF}{AC} = \frac{BH}{BC} \Leftrightarrow \frac{\hat\beta_1^{(1)}x}{x} = \frac{\hat\beta_1^{(2)}\tilde x}{\tilde x} \Leftrightarrow \hat\beta_1^{(1)} = \hat\beta_1^{(2)}
-\]
-\end{proof}
-
-%\section{Geometric properties of distributions}
-
-\section{Partial correlation}
-
-\subsection{Definition of partial correlation}
-
-\marginnote{
-Partial correlation is the measure of degree of dependence between two random
-variables while controlling for the effect of other random variables:
-\[
-\pCorr(X,Y; Z) = \frac{\pCov(X,Y;Z)}{\sqrt{\pVar(X;Z)\pVar(Y;Z)}}
-\]
-where $\pVar(X;Z) = \Var(X - \alpha Z)$, $\alpha$ is such a constant that $\Cov(X - \alpha Z, Z) = 0$,
-and $\pCov(X,Y; Z) = \Cov(X - \alpha Z, Y - \beta Z)$, $\alpha$, $\beta$ are such
-constants that $\Cov(X - \alpha Z, Z) = \Cov(Y - \beta Z, Z) = 0$
-}
-
-Partial correlation can be defined into two ways.
-We will provide both definitions and show their equivalence.
-
-\begin{definition}
-Partial correlation between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the correlation coefficient between the residuals in regression of $X$ onto
-$Z$ and the residuals in regression of $Y$ onto $Z$.
-\end{definition}
-
-Firstly, we project random variable $X$ onto $Z$, which yields $\E(X)$.
-The residuals in this regression are $X - \E(X)$ — a vector in $Lin^{\perp}(Z)$.
-We will call this variable `centered' and label it as $\widetilde X$.
-Repeating this step for $Y$ yields `centerd' variable $\tilde Y = Y - \E(Y) \in Lin^{\perp}(Z)$.
-The angle between $\widetilde X$ and $\widetilde Y$ ($\varphi$ in Figure~\ref{fig:pcorr_def1})
-is the correaletion coefficient between these `centred' random variables and
-the partial correlaiton between the original ones.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.55\linewidth]{images/partial_corr.pdf}
-  \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
-  \label{fig:pcorr_def1}
-\end{figure}
-
-\begin{definition}
-Partial correlaiton between random variables $X$ and $Y$ holding random variable $Z$
-fixed is the geometric mean between the coefficient $\hat \beta_{XY}$ in regression
-\[
-\hat X = \hat \beta_{XY} Y + \hat \beta_{XZ} Z
-\]
-and the coefficinent $\hat \beta_{YX}$ in regression
-\[
-\hat Y = \hat \beta_{YX} X + \hat \beta_{YZ} Z
-\]
-which has the same sign as the coefficients $\hat \beta_{XY}$ and $\hat \beta_{YX}$.
-\end{definition}
-
-Following the definition, we need to start with regressing variable $X$ onto
-$Y$ and $Z$. Then, the vector we obtained $\hat X$ can be broken up into the sum of
-$\hat \beta_{XY} Y$ and $\hat \beta_{XZ} Z$.
-Projecting $\hat \beta_{XY} Y$ onto $Lin^{\perp}(Z)$ we results in a vector
-$\alpha \widetilde Y$ where $\widetilde Y = Y - \E(Y)$ is the projection of $Y$ onto $Lin^{\perp}(Z)$.
-
-By the properties of similar triangles
-\[
-\frac{\beta_{XY} Y}{Y} = \frac{\alpha \widetilde Y}{\widetilde Y} \Leftrightarrow \beta_{XY} = \alpha
-\]
-
-In the same way we perform a regression of $Y$ onto $X$ and $Z$ and repeat the
-same steps as for $X$. Finally, we get the whole picture:
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.55\linewidth]{images/part_corr_2.pdf}
-\label{fig:part_corr_alt}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.35\linewidth]{images/part_corr_2_lin.pdf}
-\label{fig:part_corr_alt_lin}}
-\caption{\subref{fig:part_corr_alt}: Alternative definition of the partial correlation;
-\subref{fig:part_corr_alt_lin}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-Having plotted $Lin^{\perp}(Z)$ now we can express $\cos \varphi$ in terms of $\beta_{XY}$
-and $\beta_{YX}$
-
-\begin{equation}\label{eq:part_cor_cos}
-\begin{split}
-\cos \varphi &= \frac{\vert \beta_{XY} \widetilde Y \vert}{\vert \widetilde X \vert} = \vert \beta_{XY} \vert \\
-\cos \varphi &= \frac{\vert \beta_{YX} \widetilde X \vert}{\vert \widetilde Y \vert} = \vert \beta_{YX} \vert \\
-\cos^2 \varphi &= \vert \beta_{XY} \beta_{YX} \vert \stackrel{sign(\beta_{XY}) = sign(\beta_{YX})}{=} \beta_{XY} \beta_{YX}
-\end{split}
-\end{equation}
-
-Recall that the angle $\varphi$ can be interpreted as the partial correlation
-between $X$ and $Y$ holding $Z$ fixed, so it follows form equaitions~\eqref{eq:part_cor_cos}
-
-\[
-\pCorr^2(X,Y; Z) = \cos^2 \varphi = \beta_{XY} \beta_{YX}
-\]
-
-\subsection{Partial correlation as correlaiton between residuals}
-
-\marginnote{
-???
-}
-
-\begin{theorem}
-Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
-correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
-\[
-X = \alpha_1 Y + \alpha_2 Z + u
-\]
-and the residuals $\hat v$ in the model
-\[
-Y = \beta_1 Z + \beta_2 Z + v
-\]
-\end{theorem}
-
-\begin{proof}
-The first step is to find the residuals in the mentioned regresisions.
-For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
-which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
-and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
-in Figure~\ref{fig:pcorr_t_x}.
-
-Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
-\label{fig:pcorr_t_x}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
-\label{fig:pcorr_t_y}}
-\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
-\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
-\end{center}
-\end{figure}
-
-After all we put these figure together and the goal is to find the angle between
-the red and blue lines which is the same as the angle between the dashed
-red and blue lines. However, before that we need to apply translating to them to
-get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
-complementary angles, angles in triangles and parallelograms
-we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
-is eqaul to $\varphi$.
-
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
-\label{fig:pcorr_t_xy}}
-%\hspace{4ex}
-\subfigure[]{
-\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
-\label{fig:pcorr_t_lin}}
-\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
-\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
-\end{center}
-\end{figure}
-
-\end{proof}
+\input{chapters/4_distributions.tex}
 

### 2018-02-03 2018/chapters/2_regression.tex
restructuring finished, fwl v2

@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/tss_sum_base.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_yhat.pdf}
 \label{fig:tss_base} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/tss_sum_rss.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_rss.pdf}
 \label{fig:tss_rss}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/tss_sum_tss.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_tss.pdf}
 \label{fig:tss_tss}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/tss_sum_ess.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_rss_ess_tss_sqr_ess.pdf}
 \label{fig:tss_ess}}
@@
 \begin{figure}[h!]
-\center{\includegraphics{figures/tss_sum_final.pdf}}
+\center{\includegraphics{figures/02_rss_ess_tss_final.pdf}}
 \caption{$(\sqrt{RSS})^2 + (\sqrt{ESS})^2 = (\sqrt{TSS})^2$}
@@
 \[
-\sqrt{R^2} = \frac{\sqrt{ESS}}{\sqrt{TSS}} = \cos \varphi
+R^2 = \frac{ESS}{TSS} = \cos^2 \varphi
 \]
@@
 \begin{figure}
-  \center{\includegraphics{figures/r_sq_ss.pdf}}
+  \center{\includegraphics{figures/02_determination_coefficient.pdf}}
   \caption{Determination coefficient as squared $\cos \varphi$}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/averages_1.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_averages_yhat_decomposed.pdf}
 \label{fig:averages_lin} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/averages_2.pdf}
+\includegraphics[width=0.4\linewidth]{figures/02_averages_final.pdf}
 \label{fig:averages_bars}}
-\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbb{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbb{1}$.}
+\caption{\subref{fig:averages_lin}: Regression of $y$ on $Lin(\mathbf{1},x)$; \subref{fig:averages_bars}: Regression of $y$ and $\hat y$ on $\mathbf{1}$.}
 \end{center}
@@
 \end{equation}
-where $M_1 = I - P_1$ projects onto the orthogonal complement of the column space of $X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the projection onto the column space of $X_1$.
-Then the estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the same as the estimate from regression~\ref{eq:fwl_2}.
+where $M_1 = I - P_1$ projects onto the~orthogonal complement of the~column space~of
+$X_1$ and $P_1 = X_1(X_1^TX_1)^{-1}X_1^T$ is the~projection onto the~column space of~$X_1$.
+Then the~estimate of $\beta_2$ from regression~\ref{eq:fwl_1} will be the~same
+as the~estimate from regression~\ref{eq:fwl_2}.
 \end{theorem}
 
+There are two ways to~visualize the~proof of~the~Frisch-Waugh-Lovell theorem
+using geometric concepts. Both are presented below.
+
 \begin{proof}
-Geometrical proof will be presented for the following model:
+1.~Consider the following model:
 \begin{equation} \label{eq:fwl_proof}
@@
 
-We start with regression `all-at-once' and will distinct its coefficients with index $(1)$.
-The only step in obtaining $\beta_1^{(1)}$ is regressing $y$ on $Lin(x,z)$ and then expanding $\hat y$ as a linear combination of basis vectors $x$ and $z$,
-which is shown in Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin} depicts $Lin(x, z)$.
+We start with regression `all-at-once' and will distinct its coefficients with~index $(1)$.
+The~only step in~obtaining $\beta_1^{(1)}$ is regressing $y$ on~$Lin(x,z)$ and
+then expanding $\hat y$ as a~linear combination of~basis vectors $x$ and $z$,
+which is shown in~Figure~\ref{fig:fwl_1_regression_3d}. Figure~\ref{fig:fwl_1_regression_lin}
+depicts $Lin(x, z)$.
 
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_yhat_decomposed.pdf}
 \label{fig:fwl_1_regression_3d} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_1_regression_lin.pdf}
-%\caption{$Lin(x, z)$}
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_yhat_decomposed_lin.pdf}
 \label{fig:fwl_1_regression_lin}}
@@
 
-As for the model~\ref{eq:fwl_2}, where several regressions are performed consecutively, we start with regressing $y$ on $z$, resulting in $\tilde{y}$, which we will refer to as ``cleansed'' $y$.
+As for the~model~\ref{eq:fwl_2}, where several regressions are performed consecutively,
+we start with regressing $y$~on~$z$, resulting in $\tilde{y}$,
+which we will refer~to as~`cleansed' $y$.
 
@@
 
-Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — ``cleansed'' $x$.
+Following that, $x$ is regressed on $z$, resulting in $\tilde{x}$ — `cleansed' $x$.
 
@@
 
-Finally, `cleansed' $y$ must be regressed on `cleansed' $x$. However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
-So at first, we fix this problem by translation and after taht obtain $\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
+Finally, `cleansed' $y$ must be regressed on `cleansed' $x$.
+However, it cannot be performed immediately as $\tilde{y}$ and $\tilde{x}$ are skew lines.
+So at first, we fix this problem by translation and after taht obtain
+$\hat\beta_1^{(2)}\tilde x$ (see Figure~\ref{fig:fwl_2_regression_trans}).
 
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_cleansed_variables.pdf}
 \label{fig:fwl_2_regression_first} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_2_regression_translation.pdf}
-%\caption{$Lin(x, z)$}
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_translation.pdf}
 \label{fig:fwl_2_regression_trans}}
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_3_3d.pdf}
-%\caption{Regression of $y$ on $Lin(x,z)$} %% подпись к рисунку
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_final.pdf}
 \label{fig:fwl_3_3d} }
@@
 \subfigure[]{
-\includegraphics[width=0.4\linewidth]{figures/fwl_3_lin.pdf}
-%\caption{$Lin(x, z)$}
+\includegraphics[width=0.4\linewidth]{figures/02_fwl_v1_final_lin.pdf}
 \label{fig:fwl_3_lin}}
@@
 
-In Figure~\ref{fig:fwl_3_lin} segments AF and BH = DG stand for $\hat\beta_1^{(1)}x$
+In Figure~\ref{fig:fwl_3_lin} segments $AF$ and $BH = DG$ stand for $\hat\beta_1^{(1)}x$
 and $\hat\beta_1^{(2)}\tilde x$ respectively, while segments AC and BC represent $x$ and $\tilde{x}$.
@@
 \]
+
+2. Alternatively, we could implement a~concept close to the~partial correlation.
+In the~same model~\ref{eq:fwl_proof} we wiil treat $z$ vector fixed and again
+consecutively cleanse the $x$ and $y$ variables by projecting them onto
+the~space orthogonal to $z$, i.e., $Lin^\perp(z)$ as demonstrated in Figure~\ref{fig:fwl_v2_rcleansed}.
+Then we perform a~regression of the~`cleansed' $\tilde y$ on the~`cleansed' $\tilde x$
+(see Figure\ref{fig:fwl_v2_regression_cleansed}).
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_cleansed_variables.pdf}
+\label{fig:fwl_v2_rcleansed}}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_cleansed_regression.pdf}
+\label{fig:fwl_v2_regression_cleansed}}
+\caption{\subref{fig:fwl_2_regression_first}: `Cleansed' variables $\tilde x$ and $\tilde y$;
+\subref{fig:fwl_2_regression_trans}: `Cleansed' $\tilde y$ regressed on `cleansed' $\tilde{x}$.}
+\end{center}
+\end{figure}
+
+Now we show that the~latter regression produces $\hat \beta_1$ coefficient which
+is exatly the coefficient from the~`one-step' regression of~original $y$ onto
+original $x$ and $z$. Recall that the~vector $y$ can be split~up into a~sum of
+some multiple of $x$ and some multiple of $z$. Since the second term is
+the~orthogonal component its projection yields zero. The~multiple of $x$
+is equal to $\hat \beta_1$ by construction.
+
+Assume that the coefficient at $\tilde x$ is some unknown variable $\hat \delta$.
+Then consider the similar triangles in the $Lin(x,z)$. From the~proportions
+we obtain:
+\[
+\frac{CE}{CA} = \frac{CD}{CB} \Leftrightarrow \frac{\hat \beta_1 x}{x} = \frac{\hat \delta \tilde x}{\tilde x} \Rightarrow \hat \beta_1 = \hat \delta
+\]
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_similar_triangles.pdf}
+\label{fig:fwl_v2_triangles}}
+\subfigure[]{
+\includegraphics[width=0.45\linewidth]{figures/02_fwl_v2_final.pdf}
+\label{fig:fwl_v2_final}}
+\caption{\subref{fig:fwl_2_regression_first}: Similar triangles: $\bigtriangleup ABC \sim \bigtriangleup EDC$;
+\subref{fig:fwl_2_regression_trans}: Alternative proof for the Frisch-Waugh-Lovell theorem.}
+\end{center}
+\end{figure}
+
 \end{proof}

### 2018-02-03 2018/chapters/3_partial_correlation.tex
restructuring finished, fwl v2

@@
   \centering
-  \includegraphics[width=0.55\linewidth]{figures/partial_corr.pdf}
+  \includegraphics[width=0.55\linewidth]{figures/03_partial_correlation_definition.pdf}
   \caption{Partial correlation between $X$ and $Y$ while $Z$ is fixed.}
@@
 \subfigure[]{
-\includegraphics[width=0.55\linewidth]{figures/part_corr_2.pdf}
+\includegraphics[width=0.55\linewidth]{figures/03_partial_correlation_regression_definition.pdf}
 \label{fig:part_corr_alt}}
@@
 \subfigure[]{
-\includegraphics[width=0.35\linewidth]{figures/part_corr_2_lin.pdf}
+\includegraphics[width=0.35\linewidth]{figures/03_partial_correlation_regression_definition_lin.pdf}
 \label{fig:part_corr_alt_lin}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_x.pdf}
+\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_x.pdf}
 \label{fig:pcorr_t_x}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/pcorr_theorem_y.pdf}
+\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_y.pdf}
 \label{fig:pcorr_t_y}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/pcorr_t_xy.pdf}
+\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_xy.pdf}
 \label{fig:pcorr_t_xy}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/pcorr_t_lin.pdf}
+\includegraphics[width=0.45\linewidth]{figures/03_partial_correlation_residuals_xy_lin.pdf}
 \label{fig:pcorr_t_lin}}

### 2018-02-04 2018/chapters/2_regression.tex
chi-sq def

@@
 Now we show that the~latter regression produces $\hat \beta_1$ coefficient which
-is exatly the coefficient from the~`one-step' regression of~original $y$ onto
+is exatly the~coefficient from the~`one-step' regression of~original $y$ onto
 original $x$ and $z$. Recall that the~vector $y$ can be split~up into a~sum of
@@
 
-Assume that the coefficient at $\tilde x$ is some unknown variable $\hat \delta$.
-Then consider the similar triangles in the $Lin(x,z)$. From the~proportions
+Assume that the~coefficient at~$\tilde x$ is some unknown variable $\hat \delta$.
+Then consider the~similar triangles in the $Lin(x,z)$. From the~proportions
 we obtain:

### 2018-02-04 2018/chapters/4_distributions.tex
restructuring finished, fwl v2
chi-sq def

@@
 
+\marginnote{
+Let $z_i \stackrel{iid}{\sim} \mathcal{N}(0,1)$.
+Then $Q$ follows the chi-squared distribution with $k$ degrees of freedom
+if it can be written as
+\[
+Q = z_1^2 + z_2^2 + \ldots + z_k^2.
+\]
+
+This definition is a particular case of the geometric one.
+Consider projecting a vector $z = (z_1, z_2, \ldots, z_n)$ form $\mathbb{R}^n$
+onto the~$k$-dimensional subspace $S$ of vectors which first $k$ coordinates
+are arbitrary and all the rest are zeros. As a result we would get
+\[
+\hat z = (z_1, z_2, \ldots, z_k, 0, \ldots, 0).
+\]
+Squaring the length of the projection, we obtain
+\[
+Q = \lVert \hat z \rVert = z_1^2 + z_2^2 + \ldots + z_k^2.
+\]
+}
+
+\begin{theorem}
 Consider a random vector $z \in \mathbb{R}^n$ which components are independent
 and follow standard normal distribution, $z_i \sim \mathcal{N}(0,1)$.
-Consider also a fixed $k$-dimensional subspace in $\mathbb{R}^n$ $L$.
+Consider also a fixed $k$-dimensional subspace $L$ in $\mathbb{R}^n$.
 Let the projection of vector $z$ onto the subspace $L$ be $\hat z$ and
@@
 Then $Q$ follows the chi-squared distribution with $k$ degrees of freedom.
+\end{theorem}
+
+\begin{proof}
+First, it can be shown that the~projected vector $\hat z$ is the~original vector $z$
+multiplied by the~projection matrix $H = X(X^T X)^{-1}X^T$ where the columns of $X$
+are fixed linearly independent vectors $x_1, \ldots, x_k$ in $L$
+or equivalently $colX = Lin(x_1, \ldots, x_k)$.
+This matrix is  also often referred to as `hat-matrix'.
+Then the statement in~the~theorem can be rewritten as follows:
+\[
+\hat z^T \hat z = (Hz)^T Hz = z^T H^T H z = z^T H^2 z = z^T H z,
+\]
+applying the idempotence property in the last step.
+
+Another nice property of~the~hat-mattix is symmetry.
+Thus, it can be decomposed as
+\[
+H = P D P^T,
+\]
+where we choose the vectors of matrix $P$ to be unit and orthogonal,
+and $D = diag{\lambda}$ where $\lambda_i$ is an eigenvalue of $H$.
+
+\begin{marginfigure}
+  \includegraphics[width=\linewidth]{figures/04_chi_squared_example.pdf}
+  \caption{Consider a $3$-dimensional example, $colX = Lin(x_1, x_2)$ and $col^{\perp}X = Lin(x_3)$.
+  $H x_1 = x_1$ and $H x_2 = x_2$ since they are in $colX$. However, $H x_3 = 0$ as $x_3 \perp colX$.
+  Projecting an arbitrary vector onto $colX$ yileds $Hy = \hat y \in Lin(x_1, x_2)$
+  while projecting onto $col^{\perp}X$ results in $(I-H)y = \hat{\hat y} \in Lin(x_3)$.}
+\end{marginfigure}
+
+Since $H^2 = H$ the eigenvalues are either $0$ or $1$.
+Recall that $H$ projects a vector onto $colX$.
+Then for any $x_i$, $i = 1, \ldots, k$, $H x_i = x_i \cdot 1$ since
+any $x_i$ is already in $colX$. This implies that $\lambda_1 = \ldots = \lambda_k = 1$.
+There are also $n-k$ vectors in the subspace orthogonal to $colX$.
+So for any $x_i$, $i= k+1, \ldots, n$, the orthogonal projection yields zero.
+We conclude that $\lambda_{k+1} = \ldots = \lambda_n = 0$.
+
+Rewritting the theroem statement further, we obtain
+\[
+z^T H z = z^T P D P^T z = (P^T z)^T D (P^T z) = \tilde z^T D \tilde z = \tilde z_1^2 + \ldots + \tilde z_k^2.
+\]
+Now we explore $\tilde z$ given $z \sim \mathcal{N}(0, I)$:
+\begin{align*}
+&\tilde z = P^T z \\
+&\E(\tilde z) = \E(P^T z) = P^T \E(z) = 0 \\
+&\Var(\tilde z) = \Var(P^T z) = P^T \Var(z) (P^T)^T = P^T P = I
+\end{align*}
+So we conclude that $\tilde z_1^2 + \ldots + \tilde z_k^2 \sim \chi^2_k$.
+
+\end{proof}
 
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/ttest.pdf}
+\includegraphics[width=0.45\linewidth]{figures/04_ttest.pdf}
 \label{fig:ttest_3d}}
@@
 \subfigure[]{
-\includegraphics[width=0.45\linewidth]{figures/ttest_lin.pdf}
+\includegraphics[width=0.45\linewidth]{figures/04_ttest_lin.pdf}
 \label{fig:ttest_lin}}
@@
 \centering
-\includegraphics[width=0.55\linewidth]{figures/ftest.pdf}
+\includegraphics[width=0.55\linewidth]{figures/04_ftest.pdf}
 \caption{F-statistic as the cotangens squared of $\varphi$}

### 2018-02-12 2018/chapters/4_distributions.tex
normal, t, f

@@
 \subsection{Normal}
+
+By contrast with substantial majority of books, the univariate normal distribution
+can be derived form the multivariate normal distribution.
+In this section we show how to obtain the univariate normal from the bivariate.
+
+\begin{marginfigure}
+  \includegraphics[width=\linewidth]{figures/04_normal.pdf}
+  \caption{Black dots represent the gas molecules.
+  The red dot stands for the one we catch.
+  Its speed along the horizontal axis is $v_1$, i.e., the first component of
+  the velocity vector, and its speed along the vertical axis is $v_2$.}
+\end{marginfigure}
+
+Assume there are gas molecules moving chaotically on a plane
+and we can measure the velocity vector of one of them.
+We denote this vector as $V = \begin{pmatrix} X \\ Y \end{pmatrix}$
+where $X$ and $Y$ satnd for the horizontal and vertical components respectively.
+Assume additionally that
+\begin{enumerate}
+  \item the joint distribution finction $f(x,y)$ does not depend on
+  the vector $(X,Y)^T$ direction but depends on its length only,
+  \item the orthogonal components of the velocity vector are independent.
+\end{enumerate}
+
+\marginnote{
+It is more common to introduce a standard normal distribution in terms of its PDF
+\begin{definition}
+A continuous random variable $\xi$ has a standard normal distribution
+if its PMF is given by
+\[
+f_{\xi} (x) = \frac{1}{\sqrt{2\pi}}\exp\left(-\frac{x^2}{2}\right).
+\]
+\end{definition}
+After that multivariate normal is defined.
+\begin{definition}
+Let $\xi_i \stackrel{iid}{\sim} \mathcal{N}(0, 1)$ then $\xi \sim \mathcal{N}(\vec{0}, I)$
+where $\xi = \begin{pmatrix} \xi_1 \\ \vdots \\ \xi_n \end{pmatrix}$,
+$I$ is $n \times n$ identity matrix, and its PMF is
+\[
+f_{\xi} (x_1, \ldots, x_n) = \frac{1}{(\sqrt{2\pi})^n}\exp\left(-\frac{x_1^2+\ldots+x_n^2}{2}\right).
+\]
+\end{definition}
+And finally, location-scale transformations are applied.
+}
+
+\begin{theorem}
+The only way the assumptions (1) and (2) are satisfied is when
+$X \sim \mathcal{N}(0, \sigma^2)$, $Y \sim \mathcal{N}(0, \sigma^2)$
+and $X$, $Y$ are independent.
+\end{theorem}
+
+\begin{proof}
+First of all, consider a vector $V' = \begin{pmatrix} -Y \\ X \end{pmatrix}$, i.e.,
+the original vector $V$ rotated $90^{\circ}$ counterclock-wise.
+By the assumption (1), this operation did not change the distribution of $V$.
+Hence, $V' \sim V$ which implies $-Y \sim X$ and $X \sim Y$.
+It follows that
+\[
+\begin{cases}
+\E(-Y) = \E(X) \\
+\E(X) = \E(Y)
+\end{cases}
+\]
+which holds for $\E(X) = \E(Y) = 0$ only.
+
+Likewise, $\Var(X) = \Var(Y)$ and here we additionally assume
+without loss of generality that $\Var(X) = \Var(Y) = 1$.
+
+Next, we introduce the angle between $V$ and the horizontal axis $U$ and
+the length of the velocity vector $R = \sqrt{X^2 + Y^2}$.
+Obviously, $X = R \cos U$ and $Y = R \sin U$.
+Note that since the joint distribution of $X$ and $Y$ depends only on
+the length of vector $V$ the distribution function of $U$ can only be constant,
+thus $U \sim Unif(0, 2\pi)$.
+
+Applying assumption (1) again, we conclude that the joint distribution function
+can be written as a function of the length of the velocity vector, or equivalently
+of the length squared:
+\[
+f(x,y) = h(x^2 + y^2).
+\]
+By the assumption (2), orthogonal components of $V$ are independent.
+Hence, the joint PMF can be decomposed to the product of marginal ones:
+\[
+f(x,y) = f(x)f(y) = g(x^2)g(y^2)
+\]
+where the latter equality was written for convenience.
+Putting everything together, we obtain
+\[
+h(x^2 + y^2) = g(x^2)g(y^2).
+\]
+Next, we take the derivative of both sides with respect to $y^2$ and
+then substitute $y^2 = 0$ to get a constant $k$:
+\begin{align*}
+h'(x^2 + y^2) &= g(x^2)g'(y^2) \\
+h'(x^2) &= g(x^2)g'(0) \\
+h'(x^2) &= k \cdot g(x^2)
+\end{align*}
+
+\marginnote{
+In order to obtain $k$ we computed
+\begin{align*}
+\E(X^2) &= \int_{-\infty}^{\infty} x^2 \sqrt{c} e^{kx^2} dx \\
+&= \left. x \cdot \sqrt{c} e^{kx^2} \cdot \frac{k}{2} \right|_{-\infty}^{\infty} - \int_{-\infty}^{\infty} \sqrt{c} e^{kx^2} \cdot \frac{1}{2k} dx \\
+&= - \frac{1}{2k} \int_{-\infty}^{\infty} \sqrt{c} e^{kx^2} \\
+&= - \frac{1}{2k} \cdot 1 \\
+&= 1.
+\end{align*}
+}
+
+Solving the differential equation, we obtain
+\[
+h(x^2) = c e^{kx^2}, \quad c \in \mathbb{R}.
+\]
+So the joint PMF can be written as follows:
+\[
+f(x,y) = h(x^2 + y^2) = c e^{k(x^2+y^2}
+\]
+and due to independece of $X$ and $Y$ the PMF of $X$ is
+\[
+f(x) = \sqrt{c} e^{kx^2}.
+\]
+
+\marginnote{
+In order to obtain $c$ we computed:
+\begin{align*}
+\int_{-\infty}^{\infty} \int_{-\infty}^{\infty} e^{-\frac{x^2 + y^2}{2}} dx dy &= \int_{0}^{2\pi} \int_{0}^{\infty} e^{-\frac{r^2}{2}}r dr d\theta \\
+&= \int_{0}^{2\pi} \left(\int_{0}^{\infty} e^{-u} du \right) d\theta \\
+&= \int_{0}^{2\pi} 1 d\theta \\
+&= 2 \pi.
+\end{align*}
+}
+
+In oreder to find the constant $k$, we need to solve $\E(X^2) = 1$.
+Computing the integral we obtain $k=-\frac{1}{2}$.
+
+Finally, we need to normalize $f(x) = c e^{-\frac{x^2 + y^2}{2}}$ so as to
+obtain $c$. Again, computing another integral,
+we conclude that $c=(2\pi)^{-1}$ which finishes the proof.
+\end{proof}
+
+Notice, that any other $\mathcal{N}(\mu, \sigma^2)$ can be obtained by applying
+location-scale transformations.
+
+The theorem can be generalized to the n-dimensional case.
+\begin{theorem}\label{th:mvn}
+The vector $z = \begin{pmatrix} z_1 \\ \vdots \\ z_n \end{pmatrix}$
+folllows the standard multivariate normal distribution and its components
+are independent if and only if
+\begin{enumerate}
+  \item f(z) depends on $\vert z \vert$ only,
+  \item the projections of vector $z$ onto the orthogonal subspaces $A$ and $B$
+  in $\mathbb{R}^n$ are independent.
+\end{enumerate}
+\end{theorem}
 
@@
 
-\begin{theorem}
+\begin{theorem}\label{th:chi}
 Consider a random vector $z \in \mathbb{R}^n$ which components are independent
@@
 where we choose the vectors of matrix $P$ to be unit and orthogonal,
-and $D = diag{\lambda}$ where $\lambda_i$ is an eigenvalue of $H$.
+and $D = diag{(\lambda_1, \ldots, \lambda_n)}$ where $\lambda_i$ is an eigenvalue of $H$.
 
@@
 \subsection{Student's}
+
+\marginnote{
+A continuous random variable $T$ has Student's distribution with $k$ degrees
+of freedom if it can be expressed as
+\[
+T = \frac{Z}{\gamma_k/k},
+\]
+where $Z \sim \mathcal{N}(0,1)$, $\gamma_k \sim \chi^2_{k}$ and
+$Z$, $\gamma_k$ are independent.
+}
+
+\begin{definition}
+Let  $z = \begin{pmatrix} z_0 \\ z_1 \\ \vdots \\ z_k \end{pmatrix}$
+where $z_i \sim \mathcal{N}(0, \sigma^2), i=0, \ldots, k$ and are independent.
+Let $L_1$ be 1-dimensional subspace in $\mathbb{R}^{k+1}$, $L_2$ an orthogonal to $L_1$ subspace.
+Then
+\[
+T = \frac{\lVert H_1 z \rVert}{\lVert H_2 z \rVert / \sqrt{dim L_2}}
+\]
+where $\lVert H_1 z \rVert$ is the length of projection onto $L_1$ of the vector $z$,
+$\lVert H_2 z \rVert$ — the length of the projection onto $L_2$,
+follows Student's distribution with $dim L_2$ degrees of freedom.
+\end{definition}
+
+Let us choose two orthogonal subspaces: one-dimensional $L_1$ and
+$k$-dimensional $L_2$.
+
+Previously we showed that, the squared length of projection follows
+the chi-squared distribution with the degrees of freedom equal to the dimension
+onto which the vector was projected. Thus, $\lVert H_1 z \rVert^2 \sim \chi^2_1$
+and $\lVert H_2 z \rVert^2 \sim \chi^2_{k}$.
+Now we can express $T^2$ as a ratio of the per-dimension lengths squared:
+\[
+T^2 = \frac{\lVert H_1 z \rVert^2}{\lVert H_2 z \rVert^2 \ dim L_2}
+\]
+Taking the square root of both sides, we obtain:
+\[
+T = \frac{\lVert H_1 z \rVert}{\lVert H_2 z \rVert  \sqrt{dim L_2}} = \tg \varphi
+\]
+The latter equlity can be illustarted with a $3$-dimensional example (see Figure~\ref{fig:f_dist}).
 
@@
 
+\marginnote{
+Generally, the following definition is given.
+\begin{definition}
+Let $\gamma_1 \sim \chi^2_{k_1}$, $\gamma_2 \sim \chi^2_{k_2}$,
+$\gamma_1$, $\gamma_2$ independent.
+Then
+\[
+\frac{\gamma_1/k_1}{\gamma_2/k_2} \sim F_{k_1, k_2}.
+\]
+\end{definition}
+}
+
+\begin{definition}\label{def:f}
+Let $z = \begin{pmatrix} z_1 \\ \vdots \\ z_n \end{pmatrix}$
+where $z_i \sim \mathcal{N}(0, \sigma^2)$ and are independent.
+Let $L_1$, $L_2$ be orthogonal subspaces in $\mathbb{R}^n$.
+then
+\[
+F = \frac{\lVert H_1 z \rVert^2 / dim L_1}{\lVert H_2 z \rVert^2 / dim L_2} \sim F_{dim L_1, dim L_2},
+\]
+where $\lVert H_1 z \rVert^2$, $\lVert H_2 z \rVert^2$ are the squared lengths
+of $z$ projected onto $L_1$ and $L_2$ respectively.
+\end{definition}
+
+\marginnote{
+numerical example?
+}
+
+Recall that by Theorem~\ref{th:mvn} the projections of a standard noraml vector
+onto orthogonal subspaces are independent.
+Thus, in terms of the Definition~\ref{def:f} $H_1 z$ and $H_2 z$ are independent.
+Next, from the Theorem~\ref{th:chi} where we defined the chi-squared distribution
+it follows that the squared lengths of these projections follow
+the chi-squared distribution with the number of degrees of freedom
+equal to the dimension of the subspace onto which the vector was projected.
+In other words, $\lVert H_1 z \rVert^2 \sim \chi^2_{dim L_1}$,
+$\lVert H_2 z \rVert^2 \sim \chi^2_{dim L_2}$.
+
+Taking the ratio of these length squared, we get the interpretaion
+of the angle between the original vector $z$ and its projection onto $L_1$:
+\[
+tg^2 \varphi = \frac{\lVert H_1 z \rVert^2}{\lVert H_2 z \rVert^2}.
+\]
+Adjusting this ratio to the degrees of freedom, we get the desired definition.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.55\linewidth]{figures/04_f_dist_example.pdf}
+\caption{F-distribution as the ratio of the projection lengths squared
+adjusted to the dimensions of the subspaces.}
+\label{fig:f_dist}
+\end{figure}
+
 \subsection{F-test}
